### 2014-05-12 report/draft.tex
End of night commit. Report transitioned into Latex Some graphs updated.

\documentclass[12pt]{article}

\title{\vspace{-2.0cm}Stuff}
\title{ST 599 Project 1 Report}
\author{Wanli Zhang \& Matt Edwards}
\date{21st April 2014}

\usepackage{fullpage}
\usepackage{floatrow}
\usepackage{fourier}
\usepackage{parskip}% http://ctan.org/pkg/parskip
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\begin{document}
\begingroup  
  \centering
  \LARGE Master's Report - Draft\\[1em]
  \large Matt Edwards\\
  11 May 2014\par
\endgroup

\section{Introduction}
Transects, both point and line, are a method of population density estimation often used in the ecological sciences. They are useful when populations are too small to be seen from the air (birds, rodents) or may be masked by the canopy (ground plants.) Recent studies that use distance sampling include Norwegian bird populations (Pedersen et al. 2012), Komodo dragon prey estimates (Ariefiandy et al. 2013), Serengeti carnivore abundance (Durant et al. 2011), and estimates for butterfly populations (Isaac et al. 2011). A study by Cox et al. extends point transect sampling for use with underwater acoustics for estimates of krill populations (2011), and a study by Rowcliffe et al. combines distance sampling with camera traps to estimate populations (2011).

\subsection{History}
\texttt{TO DO}

\subsection{Micronesian Bird Data}
The Micronesian Forest Bird Survey of 1982 (Ramsey, et al.) is a work using Variable Circular Plots (VCPs) placed using a random-systematic transect design to monitor the density of several bird species in the Micronesian islands.

Each island was divided into regions. Initial transects were given random positions and angles, with subsequent transects placed parallel, 2km apart. Stations were placed every 150m along the transects. Two observers, standing 20 m apart, would wait several minutes for the wildlife to quiet down after their arrival, and then spend 8 minutes noting any birds they could observe either visually or aurally. Distance and species were noted, as well as other covariates such as foliage density, weather conditions, etc.

Figure 1 illustrates the distance data for the Collared Kingfisher. This data represents the aggregated data from two islands, Tinian and Rota, and all four observers. We can observe several things in this plot:

First, there are very low counts of birds immediately around the station, a peak at around 100 m, a fairly level section from 100--200 m, and a drop off after 200 m. This would seem to indicate that the presence of the observers caused either movement away from the station (the crest around 100m) or a suppression of vocalizations and/or movement directly around the station.

Secondly, note that the detection distances are high between 100--200 m, and recall that the stations were place 150 m apart. Since the stations on a transect were (generally) visited in sequence, this means it is possible a bird was observed from more than one station.

Lastly, the further away from the station, the distance observations were heaped into categories. With smaller bin widths, this is also evident in the 100--200 m range as well.

\begin{figure}
\caption{Aggregated Collared Kingfisher Observation Distances 1982, Rota \& Tinian}
\includegraphics[width=\textwidth]{../images/histogram_distance.pdf}
\end{figure}

\texttt{I don't like the flow here.}
\section{Application}
With Line Transects, a (straight) line is randomly placed in the study area, then walked by the observer, with any objects of interest noted and the distance perpendicular to the transect recorded. With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground. This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for (moving) wildlife and watch where they are walking. (Ramsey, Scott 1979) VCP have the added benefit of allowing the wildlife to settle. Animals may react to an observer moving through the area by hiding or moving away. If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal, animal behavior may be more natural and less influenced by the action of the observer.

Observations of objects can take the form of visual sightings or detection of auditory cues, or one confirmed by the other, as long as an accurate distance can be measured.

\subsection{Notation}
This notation is borrowed from Quang (1993) following Buckland (1987)

\begin{tabular}{r l}
$t$:	&	Number of plots surveyed\\
$D$:	& Common density of animals in all plots\\
$m_j$:	& Number of detected animals in the $j$th plot\\
$\bar m$:	& Average number detected per plot\\
$\mu$:	& Expected number of animals in a plot, $\mu = E(m_j)$\\
$n$:	& Aggregate numbe of animals detected in all plots, $n=\sum_{j=1}^t m_j$\\
$\lambda$:	&	Expected total number of detected animals in all plots, $\lambda=E(n)=t\mu$\\
$R_i$:	& Distance from station to the $i$th detected animal\\
$R_1,\cdots,R_n$: & Detected distances pooled from all plots\\
$s$:	& Sample standard deviation of $R_1,\cdots,R_n$\\
$f(r)$: & Common probability density function (pdf) of the detected distances, $R_1,\cdots,R_n$\\
$h$: 	& Bandwidth used in kernel estimation
\end{tabular}\\

For a typical individual plot:

\begin{tabular}{r l}
$N$:	&	True number of animals in the plot (unknown)\\
$m$:	&	Number of detected animals in the plot\\
$g(r)$:	&	Radial detection function, $g(r)$=Pr(Animal is detected | Animal is at distance $r$)\\
$w$:	&	Distance beyond which no animals can be detected (= radius of the plot)\\
$p$:	& Mean detection probability within the plot\\
$B$:	&   $B = 2/(pw^2)$\\

\end{tabular}

\subsection{Assumptions}
Buckland (2004) lists three primary assumptions:
\begin{enumerate}
\item[i] Objects at the line or point are detected with certainty $g(0)=1$
\item[ii] Objects are detected at their initial location
\item[iii] Measurements are exact
\end{enumerate}

There are additional assumptions listed by both Buckland and Ramsey and Scott (1979) and summarized by Quang (1993):
\begin{enumerate}
\item[iv] Sightings of animals are independent events
\item[v] at least one animal is detected $n > 0$
\item[vi] detectability is the same for all animals at distance r and in all directions.
\item[vii] no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
\item[viii] the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
\item[ix] the stations are chosen at random in the study area
\end{enumerate}

Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE

\subsection{Broad Strokes}
In general, an estimate of population density is a count of objects observed divided by the area in which they were observed times the probability of observing the object. (Ramsey, Scott 1979)

Equation 1: $D = E[n]/(A_r)(P_r)$

In practice, the expected value of n is taken to be n, the number of objects observed. The estimation of the denominator is where things get tricky.

If we combine $(A_r)(P_r)$ into an effective area we get:

Equation 2: $A~_R = 2(pi) int?0?R g(y)ydy$

$G(r)$ is the detectability function, and represents the probability of detection at distance $r$. It is often scaled so that $g(0)=1$. (In terms of area of a circle, it is obvious that the $integral{g(y)ydy}$ represents $r^2$ in the $2pi r^2$ equation.)

Much of the uncertainty in VCP estimation comes from how the effective area is estimated. $g(y)$ is often modeled as a parametric class (exponential, half-normal, hazard function) which is then scaled so that $g(0)=1$. In addition to guessing which function to use, we must also estimate the parameters.

It is plausible to think that as objects are further away from the observer, they are more difficult to detect, and therefore, we would have a greater probability of missing observations that were further away. So, in calculating $g(y)$, do we assume perfect detection out to the furthest point of observation? Or do we truncate our observations at some distance closer to the observer? Ramsey \& Scott (1979) refer to these as the effective radius ($\rho$) and the basal radius ($r$) respectively.

The problem lies in that in practice, we don't know where the basal radius is, and it must be estimated from our data.

\subsection{Properties}
\subsubsection{Bias \& Consistency}
Buckland et al. (2001) and Ramsey \& Scott (1970) both stated that if our detectability estimate $g(z)=$1 out to $r$, then $D$ is unbiased.

Barry and Walsh (2001) compared design-based and  model-based estimates of density and variance. (Their focus was line transects, but they state it is analogous to point transects and the conclusions should carry across.) They concluded that model-based estimates were unbiased if and only if the assumptions of objects being independent and uniformly located held. They noted that in practice, this is probably not a feasible assumption. For design-based inference, they caution about both the density and variance estimates.

Roeder, Dennis, \& Garton (1987) use simulations to demonstrate that no method is completely applicable in 100\% of situations, and will go sideways at some point, depending on the specific conditions.

The overall conclusion is that, in theory, density estimates using VCPs are unbiased and (reasonably) consistent, but that the assumptions necessary for the theory to hold are untenable.

As stated in the section on choosing $g(r)$, it will vary depending on the specific situation, and it is up to the scientist and statistician to understand the data and choose the best approach.

\subsubsection{Independence}
Assumption \texttt{ix}, that the VCPs are placed randomly through the study area, would seem to implicitly state that their observational ranges might by chance overlap, meaning a single object of interest could be observed from more than one station. Thompson (p240) likewise indicates random placement of line transects which would seem to allow for a potential overlap of observation zones.

Buckland et al (2001) also discuss the random placement of both line and point transects, without addressing the potential for sightings to overlap. Reynolds et al. (1980) state that the possibility of observing the same bird from two stations should be avoided. Buckland (2006) says simulations showed that analysis was robust to violations of independence (meaning overlap, or multiple observations on same object).

On page 233, Buckland et al. (2001) discuss that each line or point must be ``randomly and independently located,'' with ``an equal probability of selection'' for all portions of the study area.

On page 235, Buckland et al. states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet unsampled lines.''

They do state that for random-systematic transects, the methodology used by Ramesy et al (1982), that each transect line should be treated and analyzed as a cluster.

Barry and Welsh (2001) in discussing line transects, treat it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection. This would seem to (but does not explicitly) endorse the possibility of overlapping VCPs.

One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously throughout the space [reference].  (If the data is clustered we have other problems, not addressed here.) If that is the case, then it would make sense thought-level idea that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station. This echoes Buckland’s statement a few paragraphs previous.

Thomson mentions (page 244) that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.

\section{Simulation}
As seen in the previous section, the issue of independence of the point transects does not have a consensus in the literature. If independence is required for proper estimation, then the random-systematic placement method should not be used.

Through simulation, we can explore what happens under three different VCP placement schemes, and compare the resulting density estimates to their known values to evaluate performance.

\subsection{VCP Arrangement}
Simulation will address three options for VCP arrangement:
\begin{itemize}
\item Structured: VCPs placed so that observation distances do not overlap
\item Random: VCPs placed completely at random
\item Random-Systematic: Two transects of 18 stations each, placed according to Ramsey et al's design: 150 m between stations, 2 km distance between parallel transects.
\end{itemize}

Figures 2 and 3 illustrate the first two arrangements. Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota. The surveyed area of the Palie region totaled 9.41 km$^2$ during the original survey, and contained two transects of 16 and 17 stations each, 33 stations total (Ramsey et al 1982).

As seen in Figure 1, there is very little detection happening beyond 500 m. There are 15 observations beyond 500 m, which is 0.9\% of the data.

Treating 500 m as our truncation distance $w$, the study area in the simulation has been raised to 36 km$^2$ to allow for stations where the observation area did not overlap. An area of 49 km$^2$ is generated with the given density, then truncated to 36 km$^2$ to prevent any edge effect from the random generation of points. The stations are placed a minimum of 500 km (0.5 on the graph) away from the edges to prevent border truncation.

For the random-systematic, an angle was chosen between $0$ and $\pi$. A transect of 18 stations 150 m apart was constructed, and a second transect of 18 stations constructed running parallel. The transects were then randomly placed with in the graph, no closer to the edge than 500 km (0.5 on graph.) Figures 4 and 5 are two examples.

\begin{figure}
	\centering
	\caption{VCP Layout options. On graph, 1 unit = 1 km. Circles represent 200 m mark.}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{../images/layout_structured.pdf}
		\caption{Structured Layout}
		\label{fig:structured}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{../images/layout_random.pdf}
		\caption{Totally Random Layout}
		\label{fig:structured}
	\end{subfigure}

	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{../images/layout_rand-sys-4.pdf}
		\caption{A Random-Systematic Layout}
		\label{fig:structured}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\textwidth}
		\includegraphics[width=\textwidth]{../images/layout_rand-sys-6.pdf}
		\caption{Another Random-Systematic Layout}
		\label{fig:structured}
	\end{subfigure}
\end{figure}

\end{document}

### 2014-05-12 fiddling/DataExploration.R

@@ -8,7 +8,7 @@ library(dplyr)
 library(ggplot2)
 library(lubridate) # date/time manipulation
 
-raw.data <- tbl_df(read.csv("COKI.csv", stringsAsFactors=FALSE))
+raw.data <- tbl_df(read.csv("data/COKI.csv", stringsAsFactors=FALSE))
 
 ## ===== CODING ====
 
@@ -34,7 +34,13 @@ Palie.82 <- filter(raw.data.2, Island==2, Year==82, Transect %in% c(5,11))
 write.csv(Palie.82, file="Palie_82.csv", row.names=FALSE)
 
 ## ====== Graphs =======
-ggplot(coki.82, aes(x=Distance))+geom_histogram(binwidth=10)
+ggplot(coki.82, aes(x=Distance))+geom_histogram(binwidth=20, fill="white", colour="red")
+ggplot(coki.82, aes(x=Distance))+geom_histogram(aes(fill = ..count..), binwidth=20)+scale_fill_gradient("Count")+theme_bw(18)+scale_x_continuous(breaks=seq(0,700,100))+xlab("Observation Distance (m)")+ylab("# of Birds")
+
+ggsave("images/histogram_distance.pdf", width=8, height=4)
+
+
+sum(coki.82$Distance>500)/1673
 
 
 ggplot(coki.82, aes(x=Distance, fill=factor(Type)))+geom_dotplot(stackgroups=TRUE, binwidth=1, method="histodot")+ylim(0, 0.4)+xlim(0,200)
@@ -111,3 +117,9 @@ coki.82 %.%
 
 
 filter(by.Date.2, Date== as.POSIXct("1982-03-29", tz="UTC"))
+
+
+==============================
+  
+pts <- seq(0, 700, by=20)
+hist(coki.82$Distance, breaks=pts)

### 2014-05-13 report/draft.tex
Completed: Half normal & Emperical detection functions. Completed: Detected Functions Working On: Density Estimates.

@@
 \end{enumerate}
+
+Ramsey 1979
+Uniform distribution, random process distribution, independent distribution,
 
@@
 
+\section{Simulation Setup}
+\subsection{$g(0)$}
+\subsubsection{Half Normal}
+For a maximum possible observation distance of 0.500 (500 m), the parameter for the half-normal detection function is $\theta = 8.77$, with a scale parameter $\delta = 0.179$.
+
+The half-normal parameter $\theta$ is related to the standard deviation of a standard normal distribution by the relationship:
+
+$\theta = \frac{\sqrt{\pi /2}}{\sigma}$
+
+If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where P(X > 3.5) < 0.001) then theta is:
+
+$\sigma = w/3.5 = 0.1429$
+
+$\theta = \frac{\sqrt{\pi /2}}{0.1429}=8.7732$
+
+For a half-normal with parameter $\theta=8.7732$, $f(0)=5.5852$ so to scale the value to 1:
+
+$\delta_{HN} = 1/5.5852 = 0.1790$
+
+\subsubsection{Empirical Observations}
+\begin{figure}
+
+	\includegraphics[width=\textwidth]{../images/loess.pdf}
+	\caption{Empirical Observation Distances from Original data in meters, with LOESS regression lines. Bin width=5 m\label{fig:by5}}
+
+\end{figure}
+
+We can see in Figure \ref{fig:by5} that above 100 m, observations were commonly heaped into 10 or 25 m distances, with the latter being more heavily used.
+Above 200 m, 50 m heaps were common.
+To estimate a density function, I used a loess regression with a span of 0.2 on the counts from 5 meter bins of the empirical data.
+This is the blue line in Figure \ref{fig:by5}.
+
+Using the \texttt{hist()} function in R, the 1982 observation distances were broken into counts by 5 m groups.
+Using the bin midpoints, the density estimate was multiplied by 5 to get the density value for that bin.
+A LOESS regression was run with the midpoint as $x$ and $y=density*5$, using the \texttt{loess()} package in R.
+
+This gives us an object that can be used with \texttt{predict()} to get density estimates for specific distances.
+The values generated by feeding our midpoints into our prediction function are represented by the red dashed line in Figure \ref{fig:by5}.
+
+Following the scaling of the highest point to 1, as was done with the half normal curve, where the highest point was $g(0)$, we also find $\delta$ for this empirical detection probability.
+A sequence of $x$ values was generated covering the range from 0--500 m.
+This was plugged into the \texttt{predict()} function using our LOESS object.
+The highest density value was then scaled to be 1:
+
+$\delta_{EMP}=\frac{1}{max(predicted)}=40.5780$
+
 \end{document}

### 2014-05-13 code/ExampleGraphs.R

@@ -4,6 +4,7 @@
 # requires get.Coords from VCP_Utilities.R
 
 library(dplyr)
+library(ggplot2)
 
 # get coords, using a density of 20 birds/km2
 x.y <- get.Coords( 20)
@@ -22,7 +23,7 @@ for(i in 1:36) {
 
 # generate and save plot
 # you will get warnings when running these, because it is trimming out points to get to the 0,6 bounds
-ggplot()+geom_point(data=x.y, aes(X,Y))+geom_point(aes(x=stat.X, y=stat.Y), color="#D55E00", inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour="#D55E00", size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+ggplot()+geom_point(data=x.y, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color="#D55E00", inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour="#D55E00", size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
 
 ggsave("images/layout_structured.pdf", width=8, height=8)
 
@@ -147,3 +148,33 @@ ggplot()+geom_point(data=x.y, aes(X,Y))+geom_point(data=stat.df, aes(x=stat.X, y
 
 
 ggsave("images/layout_rand-sys-6.pdf", width=8, height=8)
+
+
+## ------------------------  LOess REgression ----------------------------------
+
+hist.info2<- hist(coki.82$Distance, breaks=seq(0,700,5), freq=TRUE)
+by5 <- data.frame(x=hist.info2$mids, y=hist.info2$density*5)
+by5.lo <- loess(y~x, data=by5, span=0.2)
+predict(by5.lo,100)
+th.x <- seq(0,700,5)
+th.y <- predict(by5.lo,th.x)
+th.y[1] <- th.y[141] <- 0
+
+
[…]

### 2014-05-13 code/VCPLocationGenerators.R

@@ -1,0 +1,111 @@
+## Station Generation Code
+
+
+## Generates a structured layout
+#   hardcoded (yes, bad) for my partiular set up.
+#   returns a data frame with columns x and y
+VCP.structuredLayout <- function() {
+  stat.X <- rep(seq(0.5, 5.5,1),6)
+  stat.Y <- sort(rep(seq(0.5, 5.5,1),6))
+  stations <- data.frame(x=stat.X, y=stat.Y)
+  return(stations)
+}
+
+
+## n = quantity to generate, min= minimum x and y value, max=max x and y value
+#   assumes square layout.
+#   returns a data frame with columns x and y
+VCP.randomLayout <- function(n=36, min=0.5, max=5.5){
+  stat.X <- runif(n, min, max)
+  stat.Y <- runif(n, min, max)
+  stations <- data.frame(x=stat.X, y=stat.Y)
+  return(stations)
+}
+
+
+## Generates a random-systematic layout with 2 transects and a 0.15 (150 m) distance between stations, and 2km distance between transects
+#   returns a data frame with columns x, y and t_id for transect ID.
+#   hardcoded (still bad) for my partiular set up.
+VCP.randSystematicLayout <- function() {
+  
+  #generate random angle theta from 0 to pi (Quadrants I, & II)
+  theta <- runif(1,0,pi)
+  # generate 1 transect with 18 stations
+  start.x <- 0
+  start.y <- 0
+  trans_id <- 1
+  stat.temp <- data.frame(x=start.x, y=start.y, t_id=trans_id)
+  
+  last.x <- start.x
+  last.y <- start.y
+  for(i in 2:18){
+    next.x <- last.x+0.15*cos(theta)
+    next.y <- last.y+0.15*sin(theta)
+    stat.temp <- rbind(stat.temp, c(next.x, next.y, trans_id))
+    last.x <- next.x
+    last.y <- next.y
+  }
+  
+  # get delta = angle perpendicular to line on angle theta
+  delta <-0
+  if(theta < (pi/2)){ 
+    # if we're in quadrant I, add 90 degrees
+    delta <- pi/2 + theta
+  } else if( theta > (pi/2)){ 
+    # if we're in quadrant II, subtract 90 degrees
+    delta <- theta - pi/2
[…]

### 2014-05-13 fiddling/DataExploration.R

@@ -35,9 +35,10 @@ write.csv(Palie.82, file="Palie_82.csv", row.names=FALSE)
 
 ## ====== Graphs =======
 ggplot(coki.82, aes(x=Distance))+geom_histogram(binwidth=20, fill="white", colour="red")
-ggplot(coki.82, aes(x=Distance))+geom_histogram(aes(fill = ..count..), binwidth=20)+scale_fill_gradient("Count")+theme_bw(18)+scale_x_continuous(breaks=seq(0,700,100))+xlab("Observation Distance (m)")+ylab("# of Birds")
 
-ggsave("images/histogram_distance.pdf", width=8, height=4)
+ggplot(coki.82, aes(x=Distance))+geom_histogram(aes(fill = ..count..), binwidth=5)+scale_fill_gradient("Count")+theme_bw(18)+scale_x_continuous(breaks=seq(0,700,100))+xlab("Observation Distance (m)")+ylab("# of Birds")
+
+ggsave("images/histogram_distance5.pdf", width=8, height=4)
 
 
 sum(coki.82$Distance>500)/1673

### 2014-05-13 fiddling/SimTesting.R

@@ -1,0 +1,80 @@
+### test simulation
+library(dplyr)
+library(fdrtool) # for halfnormal which you need for project_functions
+
+# set up environment
+D.known <- 20
+max.r <- 0.5
+nsim <- 50
+
+# half normal parameters for g(r)
+hn.params <- VCP.defineHalfnorm(max.r)
+
+# for emperical g(r), pull original distances, scale to our "1=1 km" scale
+raw.data <- tbl_df(read.csv("data/COKI.csv", stringsAsFactors=FALSE))
+coki.82 <- filter(raw.data, Year==82)
+distances <- coki.82$Distance/1000 
+breakpoints <- seq(0,.7,.005)
+emp.params <- VCP.defineLoess(distances, breakpoints)
+
+# VCP points
+xy.structured <- VCP.structuredLayout() # won't change
+
+
+D.hat.df <- data.frame()
+
+## IN loop =======================================
+for(i in 1:nsim){
+  # generate map
+  x.y = get.Coords(D.known)
+  
+  # get random VCP coords
+  xy.random <- VCP.randomLayout()
+  xy.transect <- VCP.randSystematicLayout()
+  
+  # for each VCP layout
+  #   for each detection function
+  #     get the density estimate.
+  
+  D.pass <- c() # group variables
+  D.hat.df <- rbind(D.hat.df, D.pass) # add to data frame
+}
+
+
+## g.type can be "hnorm" or "emp"; it defaults to "hnorm" if an unrecognized value entered
+
+VCP.dHat <- function(xy.objects, xy.vcp, params, g.type="hnorm", transects=FALSE){
+  n.vcp <- nrow(xy.vcp)
+  
+  # for each point (transect)
+  for(i in 1:n.vcp){
+    # get next x, y 
+    curr.x <- xy.vcp[i,"x"]
+    curr.y <- xy.vcp[i,"y"]
+    # consider objects within x+/- w and y+/- w
+    #   (not considering "noise" for this sim)
+    
+    xmin = curr.x-w
+    xmax = curr.x+w
+    ymin = curr.y-w
+    ymax = curr.y+w
+    
+    candidate.xy <- filter(xy.objects, (xmin <= x & x <= xmax) & (ymin <= y & y <= ymax))
+    
+    # calculate straight-line distance from observer to object, R.j
+    
+    candidate.xy <- mutate(candidate.xy, x.dist = x-curr.x, y.dist=x-curr.y, R.j = sqrt(x.dist^2+y.dist^2))
+    
[…]

### 2014-05-13 fiddling/VCP_Utilities.R

@@ -24,7 +24,7 @@ get.Coords <- function(dens){
   cmin <- -0.5
   X <- runif(km2*dens, min=cmin, max=cmax)
   Y <- runif(km2*dens, min=cmin, max=cmax)
-  ret <- data.frame(X,Y)
+  ret <- data.frame(x=X,y=Y)
  # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
   return (ret)
 }

### 2014-05-13 fiddling/detection.R

@@ -1,0 +1,237 @@
+### 5/13 - Detection functions
+
+library(dplyr)
+library(fdrtool) # for halfnormal which you need for project_functions
+
+## test data
+x.y <- get.Coords(20)
+w <- .500 # how will the function get this?
+
+stations <- VCP.structuredLayout()
+i<-1
+
+### Candidate POints
+# INPUT: need data frame, w, 
+
+
+# for each point (transect)
+
+# get next x, y 
+curr.x <- stations[i,"x"]
+curr.y <- stations[i,"y"]
+# consider objects within x+/- w and y+/- w
+#   (not considering "noise" for this sim)
+
+xmin = curr.x-w
+xmax = curr.x+w
+ymin = curr.y-w
+ymax = curr.y+w
+
+candidate.xy <- filter(x.y, (xmin <= x & x <= xmax) & (ymin <= y & y <= ymax))
+
+# calculate straight-line distance from observer to object, R.j
+
+candidate.xy <- mutate(candidate.xy, x.dist = x-curr.x, y.dist=x-curr.y, R.j = sqrt(x.dist^2+y.dist^2))
+# feed R.j into detection function
+
+
+# store list of detected R.j in data frame, with station # and transect # if applicable
+
+
+### Detection Function g(r)
+
+# for each point, using specified g(r)
+
+# get probability of detection
+
+# get binary detected
+
+# return list of 0/1 detected.
+
+# For half normal, need theta and the scaling factor
+
+# for exponential, need --- how did I get exponential into my head?
+
+
+th=seq(0,1, length=100)
+plot(th, dexp(th,14.09))
+plot(th, exp(1-2.0408*th))
+
+dexp(0.4, 14.09)
+dexp(0,14.09)
+
+plot(th, pexp(th,14.09))
+
+paras <- VCP.defineHalfnorm(0.5)
+
+plot(th, dhalfnorm(th, paras$theta))
+dhalfnorm(0, paras$theta)
+## not sure how I'm going ot scale the exponential.
+
+
+
+## Emperical
+
+# get coki.82 from DataExploration
+
+dist.82 <- coki.82$Distance
+names(dist.82) <- "R.j"
+my_breaks<- c(seq(0,100,10),700)
+hist.info<- hist(coki.82$Distance, breaks=my_breaks)
+
+hist.info$counts
+
+nrow(filter(coki.82, Distance <100 & Distance >= 90))
+nrow(filter(coki.82, Distance <=100 & Distance > 90)) ## This is teh scheme hist uses.
+
[…]

### 2014-05-16 report/draft.tex
Worked with the simulation data, I think it's all set, but need to test it (line test) and then run a couple simulations.
Transect Analysis - not sure whether to analyze seperately and combine, or treat as one transect as they do with line transects.

@@
 \endgroup
-
+\texttt{NOTE TO FIX: I've been using POint Transect and VCP interchangably, Buckland makes a distinction (2004, pg 6) that Point Transect you only survey the area immediately around the point, where as VCP, you are censusing $\pi w^2$.
+However, in the 2006 paper, he refers to them interchangably.}
 \section{Introduction}
@@
 
+\texttt{I know I read about how to deal with the random-systematic method, but now, of course, can't find it.
+With Line transect sampling, they essentially treat the data as if it came from one long transect, so that's what I'm going to do until I can find another reference.}
+
 \end{document}

### 2014-05-16 code/VCPLocationGenerators.R

@@ -4,10 +4,12 @@
 ## Generates a structured layout
 #   hardcoded (yes, bad) for my partiular set up.
 #   returns a data frame with columns x and y
-VCP.structuredLayout <- function() {
+VCP.structuredLayout <- function(n=36) {
   stat.X <- rep(seq(0.5, 5.5,1),6)
   stat.Y <- sort(rep(seq(0.5, 5.5,1),6))
-  stations <- data.frame(x=stat.X, y=stat.Y)
+  s_id <- seq(1,n)
+  t_id <- rep(1,n)
+  stations <- data.frame(x=stat.X, y=stat.Y, s_id=s_id, t_id=t_id)
   return(stations)
 }
 
@@ -18,7 +20,9 @@ VCP.structuredLayout <- function() {
 VCP.randomLayout <- function(n=36, min=0.5, max=5.5){
   stat.X <- runif(n, min, max)
   stat.Y <- runif(n, min, max)
-  stations <- data.frame(x=stat.X, y=stat.Y)
+  s_id <- seq(1,n)
+  t_id <- rep(1,n)
+  stations <- data.frame(x=stat.X, y=stat.Y, s_id=s_id, t_id=t_id)
   return(stations)
 }
 
@@ -26,7 +30,7 @@ VCP.randomLayout <- function(n=36, min=0.5, max=5.5){
 ## Generates a random-systematic layout with 2 transects and a 0.15 (150 m) distance between stations, and 2km distance between transects
 #   returns a data frame with columns x, y and t_id for transect ID.
 #   hardcoded (still bad) for my partiular set up.
-VCP.randSystematicLayout <- function() {
+VCP.randSystematicLayout <- function(n=36) {
   
   #generate random angle theta from 0 to pi (Quadrants I, & II)
   theta <- runif(1,0,pi)
@@ -101,9 +105,11 @@ VCP.randSystematicLayout <- function() {
   x.tem = stat.temp$x + x.shift
   y.tem = stat.temp$y + y.shift
   
+  s_id = seq(1,n)
   stat.df <- data.frame(
     x = stat.temp$x + x.shift,
     y = stat.temp$y + y.shift,
+    s_id = s_id,
     t_id = stat.temp$t_id
   )
   

### 2014-05-16 fiddling/SimTesting.R

@@ -45,6 +45,8 @@ for(i in 1:nsim){
 
 VCP.dHat <- function(xy.objects, xy.vcp, params, g.type="hnorm", transects=FALSE){
   n.vcp <- nrow(xy.vcp)
+  m <- rep(0, n.vcp)
+  R.j <- data.frame()
   
   # for each point (transect)
   for(i in 1:n.vcp){
@@ -74,7 +76,26 @@ VCP.dHat <- function(xy.objects, xy.vcp, params, g.type="hnorm", transects=FALSE
     }
     
     # store list of detected R.j in data frame, with station # and transect # if applicable
-    
+    m[i] <- sum(detected)
+    R.j <- rbind(R.j, filter(candidate.xy, detected==1)$R.j)
   }
   
+}
+
+
+## needs a vector m, that is the length of the # of VCPS, and a vector R.j that is all the detected distances.
+
+
+VCP.Dhat.hnormKernel <- function(m, R.j) {
+  #VCP density from Quang, 1993
+  m.bar <- mean(m) # mean # of objects per plot.
+  n.hat <- sum(m) #sum of objects surveyed
+  h <- sd(R.j)*n.hat^(-1/5)
+  t <- length(m)
+  #D.hat.2 <- (1/(sqrt(2*pi)*pi*(h^3)*t)) * sum(R.j*exp(-((R.j)^2)/(2*h^2)    ))
+  # Quang provides two methods, the previous line should return same value as following sequence.
+  K <- sqrt(2 * pi)^(-1) * (exp((-0.5) * (R.j/h)^2) * ((-0.5) * (2 * (R.j/h))))
+  B.hat = (-2/(n.hat*h^2))*sum(K)
+  D.hat = (m.bar*B.hat)/(2*pi)
+  return(D.hat)  
 }

### 2014-05-19 code/ExampleGraphs.R
Simulations run effectively.  tested sim loop for both hnorm and emp detections. worked out kinks. ran 100 sims.

@@ -178,3 +178,7 @@ m.500 <- predict(by5.lo, 500)
 delta.RS*m.500
 
 
+hnorm.p <- VCP.defineHalfnorm(.5)
+th.x=seq(0,.6,length=100)
+th.y=dhalfnorm(th.x, hnorm.p$theta)*hnorm.p$delta
+plot(x=th.x, y=th.y, type="l")

### 2014-05-19 code/VCPFunctions.R

@@ -1,0 +1,104 @@
+#### Matt's VCP Utilitles Package
+
+library(dplyr)
+
+get.Coords <- function(dens){
+  # hard coding for 36km^2, with a 0.5 buffer on either side, then trimming.
+  #  km2 <- ha_to_km2(ha)
+  #  cmax <- sqrt(km2)
+  km2 <- 49
+  cmax <- 6.5
+  cmin <- -0.5
+  X <- runif(km2*dens, min=cmin, max=cmax)
+  Y <- runif(km2*dens, min=cmin, max=cmax)
+  ret <- data.frame(x=X,y=Y)
+  # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
+  return (ret)
+}
+
+
+# INPUT: A vector of distance values, and a vector of equally spaced breakpoints
+# RETURN: delta for Loess, and the Loess object for use with prediction
+# Try to only run this once, so that we're not repeatedly doing the loess regression.
+VCP.defineLoess <- function(distances, breakpoints){
+  hist.info <- hist(distances, breaks=breakpoints, freq=FALSE)
+  bin.w <- breakpoints[2]-breakpoints[1]
+  hist.df <- data.frame(x=hist.info$mids, y=hist.info$density*bin.w) # density should now sum to 1
+  hist.lo <- loess(y~x, data=hist.df, span=0.2) # span is hard coded for my particular application
+  
+  # find delta
+  th.x <- seq(min(breakpoints), max(breakpoints),bin.w)
+  th.y <- predict(hist.lo,th.x)
+  max.y <- which.max(th.y)
+  delta.RS <- 1/th.y[max.y]
+  
+  params <- list(delta= delta.RS, dist.lo = hist.lo)
+  
+  return(params)
+}
+
+
+# INPUT: a vector of distance values; a params object with a scaling factor named "delta" and a loess regression object named "dist.lo". Unpredictable results if distances and the loess object are on different scales.
+# RETURN: a vector of yes/no if the point was detected at that distance.
+VCP.loessProb <- function(distances, params ){
+  pi.raw <- predict(params$dist.lo, distances)*params$delta
+  detected <- VCP.detected(pi.raw)
+  return(detected)  
+}
+
+
+# INPUT: a vector of distance values; 
+#   a params object with a scaling factor named "delta" and a parameter "theta" that is the parameter to a half-normal distribution
+# RETURN: a vector of yes/no if the point was detected at that distance.
+VCP.hnormProb <- function(distances, params){
[…]

### 2014-05-19 fiddling/SimTesting.R

@@ -5,7 +5,7 @@ library(fdrtool) # for halfnormal which you need for project_functions
 # set up environment
 D.known <- 20
 max.r <- 0.5
-nsim <- 50
+nsim <- 1000
 
 # half normal parameters for g(r)
 hn.params <- VCP.defineHalfnorm(max.r)
@@ -23,6 +23,9 @@ xy.structured <- VCP.structuredLayout() # won't change
 
 D.hat.df <- data.frame()
 
+# http://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/
+pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+
 ## IN loop =======================================
 for(i in 1:nsim){
   # generate map
@@ -36,14 +39,83 @@ for(i in 1:nsim){
   #   for each detection function
   #     get the density estimate.
   
-  D.pass <- c() # group variables
+  D.hnorm.S <- VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm")
+  D.hnorm.R <- VCP.dHat(x.y, xy.random, hn.params, w=max.r, g.type="hnorm")
+  D.hnorm.T <- VCP.dHat(x.y, xy.transect, hn.params, w=max.r, g.type="hnorm")
+  
+  #D.pass <- cbind(D.hnorm.S, D.hnorm.R, D.hnorm.T) # group variables
+  
+  D.emp.S <- VCP.dHat(x.y, xy.structured, emp.params, w=max.r, g.type="emp")
+  D.emp.R <- VCP.dHat(x.y, xy.random, emp.params, w=max.r, g.type="emp")
+  D.emp.T <- VCP.dHat(x.y, xy.transect, emp.params, w=max.r, g.type="emp")
+  
+  D.pass <- cbind(D.hnorm.S, D.hnorm.R, D.hnorm.T, D.emp.S, D.emp.R, D.emp.T)
+  #D.pass <- cbind(D.emp.S)
+  
   D.hat.df <- rbind(D.hat.df, D.pass) # add to data frame
+  
+  setWinProgressBar(pb, i, title=paste(round(i/nsim*100,0),"% done"))
 }
 
 
+write.csv(D.hat.df, "data/Sim1_05-18.csv", row.names=FALSE)
+
+close(pb)
+
[…]

### 2014-05-19 fiddling/VCP_Utilities.R

@@ -1,6 +1,21 @@
 ## Utility
 library(dplyr)
 
+get.Coords <- function(dens){
+  # hard coding for 36km^2, with a 0.5 buffer on either side, then trimming.
+  #  km2 <- ha_to_km2(ha)
+  #  cmax <- sqrt(km2)
+  km2 <- 49
+  cmax <- 6.5
+  cmin <- -0.5
+  X <- runif(km2*dens, min=cmin, max=cmax)
+  Y <- runif(km2*dens, min=cmin, max=cmax)
+  ret <- data.frame(x=X,y=Y)
+  # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
+  return (ret)
+}
+
+
 ha_to_km2 <- function(ha){
   return (ha/100)
 }
@@ -15,19 +30,6 @@ n.toSim <- function(ha, dens) {
   return (km2*dens)  # if we have 9 square kilometers, we want approx "density" # of birds per km2
 }
 
-get.Coords <- function(dens){
-  # hard coding for 36km^2, with a 0.5 buffer on either side, then trimming.
-#  km2 <- ha_to_km2(ha)
-#  cmax <- sqrt(km2)
-  km2 <- 49
-  cmax <- 6.5
-  cmin <- -0.5
-  X <- runif(km2*dens, min=cmin, max=cmax)
-  Y <- runif(km2*dens, min=cmin, max=cmax)
-  ret <- data.frame(x=X,y=Y)
- # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
-  return (ret)
-}
 
 is.in.bounds <- function(x,y, xmin=0.5, xmax=5.5, ymin=0.5, ymax=5.5) {
   ok <- TRUE

### 2014-05-19 fiddling/detection.R

@@ -175,6 +175,15 @@ VCP.defineLoess <- function(distances, breakpoints){
 # RETURN: a vector of yes/no if the point was detected at that distance.
 VCP.loessProb <- function(distances, params ){
   pi.raw <- predict(params$dist.lo, distances)*params$delta
+  
+  # the Loess prediction returns NA for observations beyond the scope of the original data
+  # turn them into 0 probability, so the detected function doesn't choke.
+  
+  na.idx <- which(pi.raw %in% NA)
+  for(i in 1:length(na.idx)){
+    pi.raw[na.idx[i]] <- 0
+  }
+  
   detected <- VCP.detected(pi.raw)
   return(detected)  
 }

### 2014-05-19 report/draft.tex
Some simulation analysis.

--- 2014-05-13 report/draft.tex
@@
 \endgroup
-
+\texttt{NOTE TO (maybe) FIX: I've been using POint Transect and VCP interchangeably, Buckland makes a distinction (2004, pg 6) that Point Transect you only survey the area immediately around the point, where as VCP, you are censusing $\pi w^2$.
+However, in the 2006 paper, he refers to them interchangeably.}
 \section{Introduction}
@@
 
+\texttt{I know I read about how to deal with the random-systematic method, but now, of course, can't find it.
+With Line transect sampling, they essentially treat the data as if it came from one long transect, so that's what I'm going to do until I can find another reference.}
+
+\section{Results}
+\subsection{Half-normal vs. Empirical $g(x)$}
+\begin{figure}
+
+	\includegraphics[width=\textwidth]{../images/Emp_Vs_Hnorm.pdf}
+	\caption{1000 Simulations, Empirical vs Half-normal Detection function, for three different layouts. \label{fig:sim1}}
+
+\end{figure}
+Figure \ref{fig:sim1} shows the results of 1000 simulations, each analyzed with Empirical and Half-normal detection function, $g(x)$, and using each of the three layouts, Structured, Random, and Transect.
+All $\hat{D}$  were done using the kernel method with a normal kernel (Quang 1993).
+
+For each simulation, a single layout of ``objects'' is generated, and that map is analyzed by each of the 6 combinations of VCP layout and $g(x)$.
+
+As we would expect, the density estimates using $g(x)$ based on the Empirical distance observations is biased low.
+We also see bias in the estimates using the half-normal function for $g(x)$, but it is much less severe than in the Empirical data.
+
+Addressing our primary question of interest, does the overlap of the VCP affect the population density estimate due to violation of independence, the answer would appear to be no.
+Looking at the simulations with the Half-normal detection function, the histograms are fairly similar for the three layouts.
+
+\begin{table}
+	\caption{Empirical vs. Half-normal detection function, across all three layouts. 1000 simulations.}
+	\begin{tabular}{ r r r r r r r r r}
+	   $g(x)$	 & Layout		& Mean	& Std Dev	& LB 95 & UB 95	& 25th	& Median	&  75th \\ \hline \hline
+	   Empirical & Random		& 14.52 & 3.35 		& 7.97	& 21.08	& 12.17	& 14.46	& 16.74\\
+	   Empirical & Structured	& 14.50 & 3.22 		& 8.19 	& 20.81 & 12.38 & 14.45 & 16.62\\
+	   Empirical & Transect		& 14.46 & 3.53 		& 7.55 	& 21.37 & 11.85 & 14.36 & 16.88\\ \hline
+	 Half-normal & Random 		& 18.17 & 4.50 		& 9.34 	& 27.00 & 15.24 & 17.94 & 21.00\\
+	 Half-normal & Structured	& 17.94 & 4.29 		& 9.54 	& 26.35 & 15.19 & 17.67 & 20.78\\
+	 Half-normal & Transect		& 18.04 & 4.49 		& 9.23 	& 26.84 & 14.90 & 17.75 & 20.74\\
+
+	\end{tabular}
+	\label{table:sim1}
+\end{table}
+
+\texttt{TO DO: along with d.hat, get the variance estimate for d.hat, and return T/F if the true value was contained in the interval.}
+
 \end{document}

### 2014-05-19 code/SimulationAnalysis_05-18.R

@@ -1,0 +1,34 @@
+library(ggplot2)
+
+raw.data <- read.csv("data/Sim1_05-18.csv", stringsAsFactors=F, header=T)
+
+d.hats <- c(raw.data$D.hnorm.S, raw.data$D.hnorm.R, raw.data$D.hnorm.T,
+            raw.data$D.emp.S, raw.data$D.emp.R, raw.data$D.emp.T)
+
+g.X <- c(rep("Half-normal", 3000), rep("Empirical",3000))
+layout <- c(
+    rep("Structured",1000),
+    rep("Random", 1000),
+    rep("Transect", 1000),
+    rep("Structured",1000),
+    rep("Random", 1000),
+    rep("Transect", 1000)
+  )
+
+
+dhat.df <- data.frame(d.hats, g.X, layout)
+
+
+ggplot(dhat.df, aes(x=d.hats))+geom_histogram(aes(y=..density..), binwidth=1, colour="black", fill="white")+facet_grid(layout~g.X)+theme_bw(18)+geom_vline(xintercept=20, colour="red", linetype="longdash")+xlab("Density Estimates")+ylab("Frequency")
+
+
+ggsave("images/Emp_Vs_Hnorm.pdf", width=6, height=6)
+
+
+dhat.df %.% group_by(g.X, layout) %.% 
+  summarise(avg=mean(d.hats), sigma=sd(d.hats), med=median(d.hats),
+            lb.95 = avg-1.96*sigma, ub.95=avg+1.96*sigma,
+            lb.iqr = quantile(d.hats,0.25), ub.iqr=quantile(d.hats,0.75)
+            )
+
+

### 2014-05-21 Vectors.R
brainstorming about Vectors.

@@ -1,0 +1,17 @@
+## mainly notes to me about vector.
+
+From C and x change can get angle.
+
+From angle and new C can get x and y changes.
+
+Only necessary for the "compounded" version
+
+
+need to keep Original version somewhere & "current" x.y
+
+
+chance to move would be dnorm() based on original R.j
+then randomly pick distance. Weighted?
+
+for the "compounded" version, base next change off the updated x.y
+so, will need to have a "movement" variable: no, temp, compounded

### 2014-05-22 code/VCPFunctions.R
Fixed a couple things, tried to figure out the bias, but the Bias correctections in Quang only made it worse.

@@ -42,6 +42,13 @@ VCP.defineLoess <- function(distances, breakpoints){
 # RETURN: a vector of yes/no if the point was detected at that distance.
 VCP.loessProb <- function(distances, params ){
   pi.raw <- predict(params$dist.lo, distances)*params$delta
+  
+  # using the Empirical data, distances beyond .69 will have a probability of "NA". Need to set these to 0.
+  idx <- which(pi.raw %in% NA)
+  for(i in 1:length(idx)){
+    pi.raw[idx[i]] <- 0
+  }
+  
   detected <- VCP.detected(pi.raw)
   return(detected)  
 }

### 2014-05-22 fiddling/PerfectlyHomogenous.R

@@ -1,0 +1,61 @@
+library(ggplot2)
+library(fdrtool)
+
+xvals <- seq(.1,6, by =.2)
+yvals <-seq(.25/4, 6, by=.25)
+
+30*24
+6*6*20
+
+
+xy.perfect <- data.frame()
+ny <- length(yvals)
+for(i in 1:length(xvals)){
+ x <- rep(xvals[i],ny)
+ new <- cbind(x, y=yvals)
+ 
+ xy.perfect <- rbind(xy.perfect, new)
+  
+}
+
+
+
+stat.X <- rep(seq(0.5, 5.5,1),6)
+stat.Y <- sort(rep(seq(0.5, 5.5,1),6))
+
+angle <- seq(-pi, pi, length=50)
+ell <- data.frame()
+for(i in 1:36) {
+  df <- data.frame(gp=i, a = sin(angle)*0.2+stat.X[i], b = cos(angle)*0.2+stat.Y[i])
+  ell <- rbind(ell, df)
+}
+
+# generate and save plot
+# you will get warnings when running these, because it is trimming out points to get to the 0,6 bounds
+ggplot()+geom_point(data=xy.perfect, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color="#D55E00", inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour="#D55E00", size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+
+#ggsave("images/layout_structured.pdf", width=8, height=8)
+
+xy.structured <- VCP.structuredLayout()
+xy.random <- VCP.randomLayout()
+xy.transect <- VCP.randSystematicLayout()
+hn.params <- VCP.defineHalfnorm(.5)
+
+my.data <- data.frame()
+for(i in 1:50){
+  xy.random <- VCP.randomLayout()
+  xy.transect <- VCP.randSystematicLayout()
+  
+  d.s <- VCP.dHat(xy.perfect, xy.structured, hn.params, w=0.5, g.type="hnorm")
+  d.r <- VCP.dHat(xy.perfect, xy.random, hn.params, w=0.5, g.type="hnorm")
+  d.t <- VCP.dHat(xy.perfect, xy.transect, hn.params, w=0.5, g.type="hnorm")
+  new <- data.frame(D.hat = c(d.s, d.r, d.t), Layout=c("S", "R", "T"))
+  my.data <- rbind(my.data, new)
+}
+
+
[…]

### 2014-05-22 fiddling/QuangHalfNorm.R

@@ -1,0 +1,17 @@
+# Quang's Half Normal
+
+# this was used in Quang's simulations
+r=seq(0,2, length=100)
+tau=0.6
+y <- exp(-(r^2)/(2*tau^2))
+# that is just the kernel of hte halfnormal distribution. It defaults to 1 at 0
+# he's using Tau to control the spread, but I think my method gives me more control.
+plot(r, y)
+
+sqrt(0.5)
+
+hn.params
+
+
+y2 <- hn.params$delta*dhalfnorm(r, hn.params$theta)
+plot(r, y2)

### 2014-05-22 fiddling/SimTesting.R

@@ -3,9 +3,9 @@ library(dplyr)
 library(fdrtool) # for halfnormal which you need for project_functions
 
 # set up environment
-D.known <- 20
+D.known <- 50
 max.r <- 0.5
-nsim <- 1000
+nsim <- 100
 
 # half normal parameters for g(r)
 hn.params <- VCP.defineHalfnorm(max.r)
@@ -58,7 +58,7 @@ for(i in 1:nsim){
 }
 
 
-write.csv(D.hat.df, "data/Sim1_05-18.csv", row.names=FALSE)
+#write.csv(D.hat.df, "data/Sim1_05-18.csv", row.names=FALSE)
 
 close(pb)
 
@@ -107,15 +107,15 @@ hist(D.hat.df$D.emp.S)
 ### set up for manual looping----
 xy.objects <- x.y
 xy.vcp <- xy.structured
-params <- emp.params
-g.type= "emp"
+params <- hn.params
+g.type= "hnorm"
 i <- 1
 w <- max.r
 ####
 
 ## g.type can be "hnorm" or "emp"; it defaults to "hnorm" if an unrecognized value entered
 
-VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transects=FALSE){
+VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transects=FALSE, true.D){
   n.vcp <- nrow(xy.vcp)
   m <- rep(0, n.vcp)
   R.j <- data.frame()
@@ -187,16 +187,73 @@ ggplot()+geom_point(data=xy.objects, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y
 ## needs a vector m, that is the length of the # of VCPS, and a vector R.j that is all the detected distances.
 
 
-VCP.Dhat.hnormKernel <- function(m, R.j) {
+## ------ Testing for Bias --------
+# this uses data from a density = 50 simulation
+R.i <- R.j$R.j
+
+VCP.Dhat.hnormKernel(m, R.j$R.j)  #52.14975
+
+## - buckland p 160
+sigma.hat.2 <- sum( (R.i^2)/(2*sum(m)))
+h.hat <- 1/sigma.hat.2
+
[…]

### 2014-05-22 fiddling/detection.R

@@ -149,6 +149,7 @@ sum(th.y2, na.rm=T) # 0.992455
 
 
 ### ------------------- FUNCTIONS -------------------------
+# Already in VCP Functions
 
 # INPUT: A vector of distance values, and a vector of equally spaced breakpoints
 # RETURN: delta for Loess, and the Loess object for use with prediction

### 2014-05-22 Vectors.R
Temporary movement working Compounded movement coming along.

@@ -15,3 +15,98 @@ then randomly pick distance. Weighted?
 
 for the "compounded" version, base next change off the updated x.y
 so, will need to have a "movement" variable: no, temp, compounded
+
+
+## Temp Move
+
+# filter to candidate x.y, calculate R.j
+# update C to new C before passing to detected. need to record that updated R.j as the detection distance.
+
+## !! Candidate.xy is a filter of the (orginal.xy), so we can update it's R.j with impunity, it will only effect that VCP
+
+## perm Move
+# as above, only update the x.y not the MASTER but the list passed for that pass.
+
+
+if(move.type="temp"){
+  new.Rj = VCP.movement(R.j)
+  
+}
+
+distances <- runif(20, 0, .2)
+
+VCP.movement <- function(distances, threshold=0.1){
+  hnorm.p <- VCP.defineHalfnorm(threshold)
+  # determine probability of movement
+  pi.raw <- hnorm.p$delta*dhalfnorm(distances, hnorm.p$theta)
+  #determine if it moved
+  moved <- VCP.detected(pi.raw)
+  
+  # This is arbitrary. Assuming an average movement of 50 m (.05) at 0 distance, 
+  #   reducing linearly to 0 movement at 110 m (.110). I felt there would be more movement
+  #   closer to teh station, and less further away.
+  m<- -.05/.110 # slope downwards (less movement, further away)
+  b <- .05  # 50 meters of movement at 0
+  mvmt.base <- m*distances+b
+
+  # add a little noise to the movement
+  mvmt.noise <- rnorm(length(pi.raw), sd=(threshold/10))
+  
+  mvmt <- moved*(mvmt.base+mvmt.noise)
+  return(mvmt)
+}
+
+
+my.df <- data.frame(moved, pi.raw, distances)
+my.df[order(distances), ]
+
+origin.xy <- xy.structured[12,]
+
+## assumes candidate.xy contains fields "x.dist" which is the change in X, and R.j, which is the hypotenuse
+## origin.xy is a dataframe with at least variables x and y - NOT NEEDED
+VCP.newXY <- function(candidate.xy){
+  thetas <- acos(candidate.xy$x.dist/candidate.xy$R.j)*sign(candidate.xy$y.dist)
[…]

### 2014-05-22 fiddling/SimTesting.R

@@ -104,12 +104,12 @@ median(D.hat.df$D.emp.T)
 
 hist(D.hat.df$D.emp.S)
 
-### set up for manual looping----
+### set up for manual looping------------------
 xy.objects <- x.y
 xy.vcp <- xy.structured
 params <- hn.params
 g.type= "hnorm"
-i <- 1
+i <- 12
 w <- max.r
 ####
 

### 2014-05-23 fiddling/VCPvsLine.R
Brainstorming

@@ -1,0 +1,4 @@
+
+# E(n) = D*Area*Probability of observing something in that area. 
+
+# I suspect that with the line transect, we'll see the nice shoulder curve that we are told to expect, but that with the circular plot, we'll see something that's lower near 0, simply due to the nature of the area of circles

### 2014-05-23 code/VCPFunctions.R
Stuff dealing with circles and squares.

@@ -64,6 +64,10 @@ VCP.hnormProb <- function(distances, params){
   return(detected)  
 }
 
+VCP.dhnorm <- function(r, params){
+  return(dhalfnorm(r, params$theta)*params$delta)
+}
+
 # returns the parameters for a half normal probability
 ## INPUT:
 #  w: the radius of detectability, beyond which detectability is essentially 0

### 2014-05-23 fiddling/VCPvsLine.R

@@ -2,3 +2,54 @@
 # E(n) = D*Area*Probability of observing something in that area. 
 
 # I suspect that with the line transect, we'll see the nice shoulder curve that we are told to expect, but that with the circular plot, we'll see something that's lower near 0, simply due to the nature of the area of circles
+
+
+D <- 20
+hn.paras <- VCP.defineHalfnorm(.5)
+L <- 36*.15 # length of single transect of 36 VCP separated by .15 km
+
+pi*(.5^2) #- area of circle with radius= 0.5 = 0.7853982
+L <- (pi*(.5^2))/(2*.5)
+L <- 1
+A.lt <- 2*r*L
+A.vcp <- pi*(r^2)
+
+N.lt <- D*(2*r*L)*VCP.dhnorm(r)
+N.vcp <- D*(pi*(r^2))*VCP.dhnorm(r)
+
+
+r <- seq(0,.5, by=0.01)
+A.lt.diff <- c(0,diff(A.lt))
+A.vcp.diff <- c(0,diff(A.vcp))
+
+hist(A.lt.diff)
+hist(A.vcp.diff)
+
+plot(r, A.lt)
+plot(r, A.vcp)
+
+library(ggplot2)
+ggplot()+geom_line(aes(x=r, y=A.lt), size=2)+geom_line(aes(x=r, y=A.vcp), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("Area covered")
+
+## does this need a legend? probably. This is for a square and a circle with the same area.
+
+
+# for each "band" further from the observer, in 0.01 km increments, get the expected # by multiplying the Density*Area*Probability of being seen there.
+
+#For Probability in band, will use halfway point in interval to estimate. 
+
+d.pts <- seq(0.005, .5, by=0.01)
+p.dpts <- VCP.dhnorm(d.pts, hn.paras)
+
+
+# Density * Area in Band * probability in band 
+
+LT <- D * A.lt.diff[2:51]*p.dpts
+VCP <- D * A.vcp.diff[2:51] * p.dpts
+
+ggplot()+geom_line(aes(x=r[2:51], y=LT), size=2)+geom_line(aes(x=r[2:51], y=VCP), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("E[N]")
+
+
[…]

### 2014-05-24 report/draft.tex
Some more with movement and adding content to draft

@@
 
+\texttt{TO DO: look at 3 moved categories for the 3 layouts}
+
+\section{The Problem with Circles}
+To test the effectiveness of my movement algorithm, I compared the detection distances from the data where the objects moved, to the detection distances from the same data, assuming they stayed still.
+The movement data, (left panel, Figure \#) is squashed more towards 0.1 Km than the still data (right panel).
+But there's a problem.
+The still data doesn't show the nice shoulder at 0 that the literature instructs us to expect.
+
+\texttt{TO DO: Images}
+
+Figure \ref{fig:simDist} shows the distances from 36 randomly placed VCPs to the objects within the max observation distance.
+As we can see, there are simply fewer objects close to the observer, which increases as the distance moves away.
+(Rephrase) This is an artifact of the VCP being a circle.
+
+\texttt{TO DO: Same graph for Line transect}
+
+\begin{figure}
+	\center
+	\includegraphics[width=0.5\textwidth]{../images/simulated_bird_distances.pdf}
+	\caption{Simulated bird distances from center point of 36 randomly placed VCPs.\label{fig:simDist}}
+
+\end{figure}
+
+With a line transect, as we move away from the line, the distance "surveyed" increases linearly. With a point transect, as we move away from the point, the distance surveyed increases exponentially.
+
+\texttt{TO DO: Equations showing what happens when you compare distance at r vs distance at r+x}
+
+\begin{figure}
+	\centering
+	\caption{Squares (Line Transect, solid) vs. Circles (Point Transect, dashed)}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{../images/rect-circ-area.pdf}
+		\caption{Area of rectangle vs circle at same distance from point/line}
+		\label{fig:structured}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{../images/rect-circ-detection.pdf}
+		\caption{Estimated detection numbers for rectangle (solid) vs circle (dashed)}
+		\label{fig:structured}
+	\end{subfigure}
+
+\end{figure}
+
+Thought - is this why we use h(0) for the point transects? need to verify, h(r) = f'(r)??  budkland 1987 paper has h(r) = f(r)/r, need to check teh book.
+
 \end{document}

### 2014-05-24 Vectors.R

@@ -170,9 +170,9 @@ for(i in 1:nsim){
 
 ddist.mov <- Rj.move$Rj
 ddist.still <- Rj.still$Rj
-
-hist(ddist.mov)
-hist(ddist.still)
+brks <- seq(0,.62, by=.02)
+hist(ddist.mov, breaks=brks)
+hist(ddist.still, breaks=brks)
 
 # the hell? the "still" distances also have a drop close to the point
 
@@ -203,6 +203,10 @@ for(j in 1:36){
 
 dist.raw <- raw.Rj$R.j
 hist(dist.raw)
+
+ggplot()+geom_histogram(aes(x=dist.raw), fill="white", colour="black")+theme_bw(18)+xlab("Distance from Observer")+ylab("Number of Birds")
+
+ggsave("images/simulated_bird_distances.pdf", width=8, height=8)
 
 # so, at least with this map, we're dealing with most of the points are not within .1 to .2 of the thing
 # both with structured

### 2014-05-24 fiddling/VCPvsLine.R

@@ -7,18 +7,19 @@
 D <- 20
 hn.paras <- VCP.defineHalfnorm(.5)
 L <- 36*.15 # length of single transect of 36 VCP separated by .15 km
+r <- seq(0,.5, by=0.01)
 
 pi*(.5^2) #- area of circle with radius= 0.5 = 0.7853982
 L <- (pi*(.5^2))/(2*.5)
-L <- 1
+#L <- 1
 A.lt <- 2*r*L
 A.vcp <- pi*(r^2)
 
-N.lt <- D*(2*r*L)*VCP.dhnorm(r)
-N.vcp <- D*(pi*(r^2))*VCP.dhnorm(r)
+#N.lt <- D*(2*r*L)*VCP.dhnorm(r)
+#N.vcp <- D*(pi*(r^2))*VCP.dhnorm(r)
 
 
-r <- seq(0,.5, by=0.01)
+
 A.lt.diff <- c(0,diff(A.lt))
 A.vcp.diff <- c(0,diff(A.vcp))
 
@@ -30,6 +31,8 @@ plot(r, A.vcp)
 
 library(ggplot2)
 ggplot()+geom_line(aes(x=r, y=A.lt), size=2)+geom_line(aes(x=r, y=A.vcp), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("Area covered")
+
+ggsave("images/rect-circ-area.pdf", width=6, height=6)
 
 ## does this need a legend? probably. This is for a square and a circle with the same area.
 
@@ -48,6 +51,8 @@ LT <- D * A.lt.diff[2:51]*p.dpts
 VCP <- D * A.vcp.diff[2:51] * p.dpts
 
 ggplot()+geom_line(aes(x=r[2:51], y=LT), size=2)+geom_line(aes(x=r[2:51], y=VCP), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("E[N]")
+
+ggsave("images/rect-circ-detection.pdf", width=6, height=6)
 
 
 # what this says to me is that even with the best case scenario, original observation distance, accurate, detection exactly follows the half-normal curve , you will not see teh "shoulder" in the emperical 

### 2014-05-24 code/DensityFunctions.R
Code cleanup Starting on the movement sim

@@ -1,0 +1,67 @@
+#### Density Estimate functions
+#     Originally in the SimTesting file
+
+
+
+VCP.Dhat.hnormKernel <- function(m, R.i) {
+  #VCP density from Quang, 1993
+  m.bar <- mean(m) # mean # of objects per plot.
+  n.hat <- sum(m) #sum of objects surveyed
+  h <- sd(R.i)*n.hat^(-1/5)
+  t <- length(m)
+  #D.hat.2 <- (1/(sqrt(2*pi)*pi*(h^3)*t)) * sum(R.i*exp(-((R.i)^2)/(2*h^2)    ))
+  # Quang provides two methods, the previous line should return same value as following sequence.
+  K.prime <- sqrt(2 * pi)^(-1) * (exp((-0.5) * (R.i/h)^2) * ((-0.5) * (2 * (R.i/h))))
+  B.hat = (-2/(n.hat*h^2))*sum(K.prime)
+  D.hat = (m.bar*B.hat)/(2*pi)
+  return(D.hat)  
+}
+
+
+
+## g.type can be "hnorm" or "emp"; it defaults to "hnorm" if an unrecognized value entered
+
+VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transects=FALSE, true.D){
+  n.vcp <- nrow(xy.vcp)
+  m <- rep(0, n.vcp)
+  R.j <- data.frame()
+  
+  # for each point (transect)
+  for(i in 1:n.vcp){
+    # get next x, y 
+    curr.x <- xy.vcp[i,"x"]
+    curr.y <- xy.vcp[i,"y"]
+    # consider objects within x+/- w and y+/- w
+    #   (not considering "noise" for this sim)
+    
+    xmin = curr.x-w
+    xmax = curr.x+w
+    ymin = curr.y-w
+    ymax = curr.y+w
+    
+    candidate.xy <- filter(xy.objects, (xmin <= x & x <= xmax) & (ymin <= y & y <= ymax))
+    
+    # calculate straight-line distance from observer to object, R.j
+    
+    candidate.xy <- mutate(candidate.xy, x.dist = x-curr.x, y.dist=y-curr.y, R.j = sqrt(x.dist^2+y.dist^2))
+    
+    # feed R.j into detection function
+    detected=NULL
+    if(tolower(g.type)=="emp"){
+      detected <- VCP.loessProb(candidate.xy$R.j, params)
+      #print("emp")
[…]

### 2014-05-24 code/MovementFunctions.R

@@ -1,0 +1,45 @@
+### Movement functions
+
+# VCP.movement - temporary
+# VCP.newXY - compounded
+
+
+
+VCP.movement <- function(distances, threshold=0.1){
+  hnorm.p <- VCP.defineHalfnorm(threshold)
+  # determine probability of movement
+  pi.raw <- hnorm.p$delta*dhalfnorm(distances, hnorm.p$theta)
+  #determine if it moved
+  moved <- VCP.detected(pi.raw)
+  
+  # This is arbitrary. Assuming an average movement of 50 m (.05) at 0 distance, 
+  #   reducing linearly to 0 movement at 110 m (.110). I felt there would be more movement
+  #   closer to teh station, and less further away.
+  m<- -.05/.110 # slope downwards (less movement, further away)
+  b <- .05  # 50 meters of movement at 0
+  mvmt.base <- m*distances+b
+  
+  # add a little noise to the movement
+  mvmt.noise <- rnorm(length(pi.raw), sd=(threshold/10))
+  
+  mvmt <- moved*(mvmt.base+mvmt.noise)
+  return(mvmt)
+}
+
+
+
+## assumes candidate.xy contains fields "x.dist" which is the change in X, and R.j, which is the hypotenuse
+## origin.xy is a dataframe with at least variables x and y - NOT NEEDED
+VCP.newXY <- function(candidate.xy){
+  thetas <- acos(candidate.xy$x.dist/candidate.xy$R.j)*sign(candidate.xy$y.dist)
+  movement <- VCP.movement(candidate.xy$R.j)
+  
+  new.Rj <- candidate.xy$R.j + movement
+  
+  #these are changes in xy, not
+  x.chng <- movement*cos(thetas)
+  y.chng <- movement*sin(thetas)
+  
+  update.xy <- candidate.xy %.% mutate( new.x = x+x.chng, new.y=y+y.chng, R.j2=new.Rj)
+  return(update.xy)
+}

### 2014-05-24 code/SimulationDriver.R

@@ -1,0 +1,69 @@
+## Simulation Code
+
+## ============ Needed Libraries ============ 
+library(dplyr)
+library(fdrtool) # for halfnormal which you need for project_functions
+
+# ============ set up environment ============ 
+D.known <- 20
+max.r <- 0.5
+nsim <- 100
+
+# === half normal parameters for g(r) ===
+hn.params <- VCP.defineHalfnorm(max.r)
+
+# for emperical g(r), pull original distances, scale to our "1=1 km" scale
+raw.data <- tbl_df(read.csv("data/COKI.csv", stringsAsFactors=FALSE))
+coki.82 <- filter(raw.data, Year==82)
+distances <- coki.82$Distance/1000 
+breakpoints <- seq(0,.7,.005)
+emp.params <- VCP.defineLoess(distances, breakpoints)
+
+# === Structured VCP points (Doesn't Change) ===
+xy.structured <- VCP.structuredLayout() 
+
+
+## ====================================================================================
+#   Simulation for Half-normal Vs. Emperical, no movement
+## ====================================================================================
+
+# === Empty Data Frame for Results ===
+D.hat.df <- data.frame()
+
+# http://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/
+pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+
+## IN loop -----------------------------------------------------------------
+for(i in 1:nsim){
+  # generate map
+  x.y = get.Coords(D.known)
+  
+  # get random VCP coords
+  xy.random <- VCP.randomLayout()
+  xy.transect <- VCP.transectLayout()
+  
+  # for each VCP layout
+  #   for each detection function
+  #     get the density estimate.
+  
+  D.hnorm.S <- VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm")
+  D.hnorm.R <- VCP.dHat(x.y, xy.random, hn.params, w=max.r, g.type="hnorm")
+  D.hnorm.T <- VCP.dHat(x.y, xy.transect, hn.params, w=max.r, g.type="hnorm")
+  
+  #D.pass <- cbind(D.hnorm.S, D.hnorm.R, D.hnorm.T) # group variables
+  
[…]

### 2014-05-24 code/VCPFunctions.R

@@ -3,20 +3,6 @@
 library(dplyr)
 library(fdrtool) # for half normal distribution
 
-get.Coords <- function(dens){
-  # hard coding for 36km^2, with a 0.5 buffer on either side, then trimming.
-  #  km2 <- ha_to_km2(ha)
-  #  cmax <- sqrt(km2)
-  km2 <- 49
-  cmax <- 6.5
-  cmin <- -0.5
-  X <- runif(km2*dens, min=cmin, max=cmax)
-  Y <- runif(km2*dens, min=cmin, max=cmax)
-  ret <- data.frame(x=X,y=Y)
-  # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
-  return (ret)
-}
-
 
 # INPUT: A vector of distance values, and a vector of equally spaced breakpoints
 # RETURN: delta for Loess, and the Loess object for use with prediction

### 2014-05-24 code/VCPLocationGenerators.R

@@ -1,6 +1,19 @@
 ## Station Generation Code
 
 
+get.Coords <- function(dens){
+  # hard coding for 36km^2, with a 0.5 buffer on either side, then trimming.
+  km2 <- 49
+  cmax <- 6.5
+  cmin <- -0.5
+  X <- runif(km2*dens, min=cmin, max=cmax)
+  Y <- runif(km2*dens, min=cmin, max=cmax)
+  idx <- seq(1:length(X))
+  ret <- data.frame(x=X,y=Y,o_id=idx)
+  # ret2 <- filter(ret, !((X < 0 | X > 6) | (Y < 0 | Y > 6)))
+  return (ret)
+}
+
 ## Generates a structured layout
 #   hardcoded (yes, bad) for my partiular set up.
 #   returns a data frame with columns x and y

### 2014-05-24 code/VCPvsLine.R

@@ -57,4 +57,4 @@ ggsave("images/rect-circ-detection.pdf", width=6, height=6)
 
 # what this says to me is that even with the best case scenario, original observation distance, accurate, detection exactly follows the half-normal curve , you will not see teh "shoulder" in the emperical 
 
-# because of the way the area of a circle works, if we did see a shoulder in emperical data from VCP
+# because of the way the area of a circle works, if we did see a shoulder in emperical data from VCP, something would be wrong

### 2014-05-24 fiddling/Vectors.R

@@ -1,5 +1,9 @@
 ## mainly notes to me about vector.
 
+## 5/24 - functions copied into Code folder
+# may be some useful things here confirming that the movement code works.
+
+
 From C and x change can get angle.
 
 From angle and new C can get x and y changes.

### 2014-05-24 fiddling/detection.R

@@ -3,6 +3,8 @@
 library(dplyr)
 library(fdrtool) # for halfnormal which you need for project_functions
 
+## -- FUNCTIONS ARE IN VCPFUNCTIONS
+
 ## test data
 x.y <- get.Coords(20)
 w <- .500 # how will the function get this?

### 2014-05-24 fiddling/moveSimTest.R

@@ -1,0 +1,29 @@
+
+xy.vcp <- VCP.structuredLayout()
+xy.objects <- get.Coords(20)
+params <- VCP.defineHalfnorm(.5)
+w <- 0.5
+g.type <- "hnorm"
+m.type <- "temp"
+true.D <- 20
+
+VCP.dhat.movement <- function ((xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", m.type="temp", true.D=20)){
+  n.vcp <- nrow(xy.vcp)
+  m <- rep(0, n.vcp)
+  observed <- data.frame()
+  
+  # for each VCP
+  
+  # filter points to current
+  
+  # initiate movement
+  
+  # get observed based on new positions
+  
+  # if mtype="perm"
+    # update original database.
+  
+  # once done w/ all VCP, get D.hat based on movement positions.
+  
+  
+}

### 2014-05-26 code/functions/DensityFunctions.R
FInal simulations, with movement.

@@ -3,7 +3,7 @@
 
 
 
-VCP.Dhat.hnormKernel <- function(m, R.i) {
+VCP.Dhat.hnormKernel <- function(m, R.i, true.D=NULL) {
   #VCP density from Quang, 1993
   m.bar <- mean(m) # mean # of objects per plot.
   n.hat <- sum(m) #sum of objects surveyed
@@ -14,7 +14,16 @@ VCP.Dhat.hnormKernel <- function(m, R.i) {
   K.prime <- sqrt(2 * pi)^(-1) * (exp((-0.5) * (R.i/h)^2) * ((-0.5) * (2 * (R.i/h))))
   B.hat = (-2/(n.hat*h^2))*sum(K.prime)
   D.hat = (m.bar*B.hat)/(2*pi)
-  return(D.hat)  
+  
+  ret <- NULL
+  if(is.null(true.D))
+    ret <- D.hat
+  else{
+    inCI <- VCP.inCI(D.hat, h, true.D, t)
+    ret <- list(Dhat = D.hat, inCI = inCI)
+  }
+  
+  return(ret)  
 }
 
 
@@ -62,6 +71,66 @@ VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transect
     R.j <- rbind(R.j, candidate.xy %.% filter(detected==1) %.% select(x, y, R.j) %.% mutate(s_id = i, t_id=xy.vcp$t_id[i]))
   }
   
-  dhat <- VCP.Dhat.hnormKernel(m, R.j$R.j)
+  dhat <- VCP.Dhat.hnormKernel(m, R.j$R.j, true.D)
+  return(dhat)
+}
+
+
+VCP.dhat.movement <- function (xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", m.type="temp", true.D=20){
+  n.vcp <- nrow(xy.vcp)
+  m <- rep(0, n.vcp)
+  observed <- data.frame()
+  
+  # for each VCP
+  for(i in 1:n.vcp){
+    
+    # filter points to current
+    curr.x <- xy.vcp[i,"x"]
+    curr.y <- xy.vcp[i,"y"]
+    
+    # consider objects within x+/- w and y+/- w
+    #   (not considering "noise" for this sim)
+    xmin = curr.x-w
+    xmax = curr.x+w
+    ymin = curr.y-w
+    ymax = curr.y+w
+    
[…]

### 2014-05-26 fiddling/Vectors.R

@@ -40,7 +40,7 @@ if(move.type="temp"){
 distances <- runif(20, 0, .2)
 
 VCP.movement <- function(distances, threshold=0.1){
-  hnorm.p <- VCP.defineHalfnorm(threshold)
+  hnorm.p <- VCP.defineHalfnorm(threshold*2, .75)
   # determine probability of movement
   pi.raw <- hnorm.p$delta*dhalfnorm(distances, hnorm.p$theta)
   #determine if it moved
@@ -49,12 +49,12 @@ VCP.movement <- function(distances, threshold=0.1){
   # This is arbitrary. Assuming an average movement of 50 m (.05) at 0 distance, 
   #   reducing linearly to 0 movement at 110 m (.110). I felt there would be more movement
   #   closer to teh station, and less further away.
-  m<- -.05/.110 # slope downwards (less movement, further away)
+  m<- -.05/.150 # slope downwards (less movement, further away)
   b <- .05  # 50 meters of movement at 0
   mvmt.base <- m*distances+b
 
   # add a little noise to the movement
-  mvmt.noise <- rnorm(length(pi.raw), sd=(threshold/10))
+  mvmt.noise <- abs(rnorm(length(pi.raw), sd=(threshold/10))) # try this with only positive noise
   
   mvmt <- moved*(mvmt.base+mvmt.noise)
   return(mvmt)
@@ -175,8 +175,9 @@ for(i in 1:nsim){
 ddist.mov <- Rj.move$Rj
 ddist.still <- Rj.still$Rj
 brks <- seq(0,.62, by=.02)
-hist(ddist.mov, breaks=brks)
+hist(ddist.mov, breaks=brks, freq=FALSE)
 hist(ddist.still, breaks=brks)
+abline(a=.05, b=-.05/.150)
 
 # the hell? the "still" distances also have a drop close to the point
 

### 2014-05-26 fiddling/moveSimTest.R

@@ -1,4 +1,126 @@
 
+## ============ Needed Libraries ============ 
+library(dplyr)
+library(fdrtool) # for halfnormal which you need for project_functions
+library(ggplot2)
+
+# ============ set up environment ============ 
+D.known <- 20
+max.r <- 0.5
+nsim <- 1000
+OSUOrange <- rgb(243/255,115/255,33/255) #http://oregonstate.edu/brand/color-palette
+OSURed <- rgb(192/255,49/225,26/255)
+OSUBlue1 <- rgb(93/255, 135/255, 161/255)
+OSUBlue2 <- rgb(156/255, 197/255, 202/255)
+
+# === half normal parameters for g(r) ===
+hn.params <- VCP.defineHalfnorm(max.r)
+
+# not doing emperical for this
+# for emperical g(r), pull original distances, scale to our "1=1 km" scale
+#raw.data <- tbl_df(read.csv("data/COKI.csv", stringsAsFactors=FALSE))
+#coki.82 <- filter(raw.data, Year==82)
+#distances <- coki.82$Distance/1000 
+#breakpoints <- seq(0,.7,.005)
+#emp.params <- VCP.defineLoess(distances, breakpoints)
+
+# === Structured VCP points (Doesn't Change) ===
+xy.structured <- VCP.structuredLayout() 
+
+
+## ====================================================================================
+#   Simulation for Half-normal Vs. Emperical, no movement
+## ====================================================================================
+
+# === Empty Data Frame for Results ===
+D.hat.df <- data.frame()
+
+# http://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/
+pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+
+## IN loop -----------------------------------------------------------------
+for(i in 1:nsim){
+  # generate map
+  x.y = get.Coords(D.known)
+  
+  # get random VCP coords
+  xy.random <- VCP.randomLayout()
+  xy.transect <- VCP.transectLayout()
+  
+  # for each VCP layout
+  #   for each detection function
+  #     get the density estimate.
+  
+  # no movement
+  
+  D.S.nomove <- c(VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm",
+                           true.D=D.known),
[…]

### 2014-05-26 report/draft.tex
Report Staging

@@
 \endgroup
-\texttt{NOTE TO (maybe) FIX: I've been using POint Transect and VCP interchangeably, Buckland makes a distinction (2004, pg 6) that Point Transect you only survey the area immediately around the point, where as VCP, you are censusing $\pi w^2$.
+\texttt{NOTE TO (maybe) FIX: I've been using Point Transect and VCP interchangeably, Buckland makes a distinction (2004, pg 6) that Point Transects you only survey the area immediately around the point, where as VCP, you are censusing $\pi w^2$.
 However, in the 2006 paper, he refers to them interchangeably.}

### 2014-05-26 fiddling/moveSimTest.R

@@ -112,6 +112,8 @@ write.csv(D.hat.df, "data/moveSim_05-25.csv", row.names=F)
 
 ggplot(D.hat.df, aes(x=Dhat))+geom_histogram(aes(y=..density..), binwidth=1, colour="white", fill=OSUOrange)+facet_grid(Layout~MoveType)+theme_bw(18)+geom_vline(xintercept=20, colour=OSURed, linetype="longdash", size=1)+xlab("Density Estimates")+ylab("Frequency")
 
+ggsave("images/MovementSim2.pdf", width=10, height=6)
+
 #hmm. not really the drastic differences I was expecting.
 #after tweaking movement, it looks better
 

### 2014-05-27 report/draft.tex
Reworked document, with bibliography

--- 2014-05-24 report/draft.tex
@@
 \documentclass[12pt]{article}
-
-\title{\vspace{-2.0cm}Stuff}
-\title{ST 599 Project 1 Report}
-\author{Wanli Zhang \& Matt Edwards}
-\date{21st April 2014}
 
@@
 \usepackage{subcaption}
+%\usepackage[round,authoryear]{natbib}
+\usepackage[utf8]{inputenc}
+\usepackage[backend=biber,citestyle=authoryear,bibstyle=authoryear,maxbibnames=10]{biblatex}
+\addbibresource{VCPbibliography.bib}
 
 \begin{document}
+cite \cite{buckland2001} \cite{komodo2013} \\
+cite* \cite*{buckland2001} \cite*{komodo2013}\\
+parencite \parencite{buckland2001} \parencite{komodo2013}\\
+parencite* \parencite*{buckland2001} \parencite*{komodo2013}\\
+textcite \textcite{buckland2001} \textcite{komodo2013}\\
+smartcite - does the super script/foot note thing.
+\newpage
 \begingroup  
   \centering
-  \LARGE Master's Report - Draft\\[1em]
+  \LARGE Master's Report - Variable Circular Plots: Station Placement and the Independence Assumption\\[1em]
   \large Matt Edwards\\
-  11 May 2014\par
+  26 May 2014\par
 \endgroup
-\texttt{NOTE TO (maybe) FIX: I've been using POint Transect and VCP interchangeably, Buckland makes a distinction (2004, pg 6) that Point Transect you only survey the area immediately around the point, where as VCP, you are censusing $\pi w^2$.
-However, in the 2006 paper, he refers to them interchangeably.}
+
 \section{Introduction}
-Transects, both point and line, are a method of population density estimation often used in the ecological sciences.
-They are useful when populations are too small to be seen from the air (birds, rodents) or may be masked by the canopy (ground plants.) Recent studies that use distance sampling include Norwegian bird populations (Pedersen et al.
-2012), Komodo dragon prey estimates (Ariefiandy et al.
-2013), Serengeti carnivore abundance (Durant et al.
-2011), and estimates for butterfly populations (Isaac et al.
-2011).
-A study by Cox et al.
-extends point transect sampling for use with underwater acoustics for estimates of krill populations (2011), and a study by Rowcliffe et al.
-combines distance sampling with camera traps to estimate populations (2011).
+Estimating population density of plants or animals is a persistent issue in the ecological sciences.
+Distance sampling, in the form of line and point transects (alternatively, Variable Circular Plots (VCP)), is a common estimation method.
+They are useful when populations are too small to be seen from the air (birds, rodents) or may be masked by the canopy (ground plants.)
 
-\subsection{History}
-\texttt{TO DO}
+Distance sampling has been around since the early 1900's.
+In the 1970's, more rigorous scientific methods were applied, including papers by Ramsey and Scott \parencite*{ramsey1979,ramsey1981} which augmented work done by Emlen in 1971.
+\cite{burnham1980} is a monograph on the theory and application of line transect sampling.
+Non-parametric methods have been examined by \textcite{quang1993,mack1998}.
+Distance sampling continues to be used and modified, combined with mark-recapture methods \parencite{laake2011}, extended for use with underwater acoustics for estimates of krill populations \parencite{krill2011}, and a study by \textcite{camera2011} combines distance sampling with camera traps to estimate populations.
+
+\textcite{buckland2001} has a more in-depth discussion of the history of distance sampling.
+
+Recent studies that use distance sampling include Norwegian bird populations \parencite{pedersen2012}, Komodo dragon prey estimates \parencite{komodo2013}, Serengeti carnivore abundance \parencite{serengeti2011}, and estimates for butterfly populations \parencite{butterfly2011}.
+
+With Line Transects, a (straight) line is randomly placed in the study area, then walked by the observer, with any objects of interest noted and the distance perpendicular to the transect recorded.
+With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground.
+This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for (moving) wildlife and watch where they are walking \cite{ramsey1979}.
+VCP have the added benefit of allowing the wildlife to settle.
+Animals may react to an observer moving through the area by hiding or moving away.
+If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal, animal behavior may be more natural and less influenced by the presence of the observer.
+
+Observations of objects can take the form of visual sightings or detection of auditory cues, or one confirmed by the other, as long as an accurate distance can be measured.
+
+\textcite[6]{buckland2001} draws a distinction between point transects, where only the area immediately surrounding the point is surveyed (area of estimated perfect detectability), and Variable Circular Plots, where you census the area $\pi w^2$, where $w$ is the furthest distance at which you can observe the object of interest.
+However, in \textcite{buckland2006, quang1993} the terms are used interchangeably.
+They will be used interchangeably here also, with the understanding this paper addresses the second type, where an area of $pi*w^2$ is censused.
 
 \subsection{Micronesian Bird Data}
-The Micronesian Forest Bird Survey of 1982 (Ramsey, et al.) is a work using Variable Circular Plots (VCPs) placed using a random-systematic transect design to monitor the density of several bird species in the Micronesian islands.
+\begin{figure}
+	\caption{Aggregated Collared Kingfisher Observation Distances, Rota \& Tinian, 20 m bins: \cite{micronesian}}
+	\includegraphics[width=\textwidth]{../images/histogram_dist_20m.pdf}
+	\label{fig:82dist}
+\end{figure}
+The Micronesian Forest Bird Survey \cite{micronesian} is a work using Variable Circular Plots (VCPs) placed using a random-systematic transect design to monitor the density of several bird species in the Micronesian islands.
 
-Each island was divided into regions.
-Initial transects were given random positions and angles, with subsequent transects placed parallel, 2km apart.
+Each of the five islands was divided into regions.
+Initial transects were given random positions and angles, with subsequent transects placed parallel, 2km apart (the ``random-systematic'' method).
 Stations were placed every 150m along the transects.
 Two observers, standing 20 m apart, would wait several minutes for the wildlife to quiet down after their arrival, and then spend 8 minutes noting any birds they could observe either visually or aurally.
-Distance and species were noted, as well as other covariates such as foliage density, weather conditions, etc.
+Distance and species were noted, as well as other covariates such as foliage density, and weather conditions.
 
-Figure 1 illustrates the distance data for the Collared Kingfisher.
+Figure~\ref{fig:82dist} illustrates the distance data for the Collared Kingfisher.
 This data represents the aggregated data from two islands, Tinian and Rota, and all four observers.
@@
 First, there are very low counts of birds immediately around the station, a peak at around 100 m, a fairly level section from 100--200 m, and a drop off after 200 m.
-This would seem to indicate that the presence of the observers caused either movement away from the station (the crest around 100m) or a suppression of vocalizations and/or movement directly around the station.
+This would seem to indicate that the presence of the observers caused either movement away from the station (the crest around 100m) or alternatively a suppression of vocalizations and/or movement directly around the station.
 
-Secondly, note that the detection distances are high between 100--200 m, and recall that the stations were place 150 m apart.
-Since the stations on a transect were (generally) visited in sequence, this means it is possible a bird was observed from more than one station.
+Secondly, the further away from the station, the distance observations were heaped into convenient distances. With smaller bin widths, this is also evident in the 100--200 m range as well.
 
-Lastly, the further away from the station, the distance observations were heaped into categories. With smaller bin widths, this is also evident in the 100--200 m range as well.
+Lastly, note that the detection distances are high between 100--200 m, and recall that the stations were placed 150 m apart.
+Since the stations on a transect were (generally) visited in sequence, this means it is possible a bird could have been observed from more than one station.
 
-\begin{figure}
-\caption{Aggregated Collared Kingfisher Observation Distances 1982, Rota \& Tinian}
-\includegraphics[width=\textwidth]{../images/histogram_distance.pdf}
-\end{figure}
+\subsubsection{Independence}
+Both \textcite{ramsey1979,buckland1987} discuss the assumption that the VCPs are placed randomly through the study area.
+This would seem to implicitly state that their observational ranges might by chance overlap, meaning a single object of interest could be observed from more than one station.
+\textcite[240]{thompson2012} likewise indicates random placement of line transects which would seem to allow for a potential overlap of observation zones.
 
-\texttt{I don't like the flow here.}
-\section{Application}
-With Line Transects, a (straight) line is randomly placed in the study area, then walked by the observer, with any objects of interest noted and the distance perpendicular to the transect recorded.
-With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground.
-This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for (moving) wildlife and watch where they are walking.
-(Ramsey, Scott 1979) VCP have the added benefit of allowing the wildlife to settle.
-Animals may react to an observer moving through the area by hiding or moving away.
-If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal, animal behavior may be more natural and less influenced by the action of the observer.
+\textcite{buckland2001} discuss the random placement of both line and point transects, without addressing the potential for sightings to overlap.
+\textcite{reynolds1980} state that the possibility of observing the same bird from two stations should be avoided.
+\textcite{buckland2006} says simulations showed that analysis was robust to violations of independence (meaning overlap, or multiple observations on same object).
 
-Observations of objects can take the form of visual sightings or detection of auditory cues, or one confirmed by the other, as long as an accurate distance can be measured.
+\textcite[233]{buckland2001} discuss that each line or point must be ``randomly and independently located,'' with ``an equal probability of selection'' for all portions of the study area.
 
-\subsection{Notation}
-This notation is borrowed from Quang (1993) following Buckland (1987)
+\textcite[235]{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet unsampled lines.''
 
-\begin{tabular}{r l}
-$t$:	&	Number of plots surveyed\\
-$D$:	& Common density of animals in all plots\\
-$m_j$:	& Number of detected animals in the $j$th plot\\
-$\bar m$:	& Average number detected per plot\\
-$\mu$:	& Expected number of animals in a plot, $\mu = E(m_j)$\\
-$n$:	& Aggregate numbe of animals detected in all plots, $n=\sum_{j=1}^t m_j$\\
-$\lambda$:	&	Expected total number of detected animals in all plots, $\lambda=E(n)=t\mu$\\
-$R_i$:	& Distance from station to the $i$th detected animal\\
-$R_1,\cdots,R_n$: & Detected distances pooled from all plots\\
-$s$:	& Sample standard deviation of $R_1,\cdots,R_n$\\
-$f(r)$: & Common probability density function (pdf) of the detected distances, $R_1,\cdots,R_n$\\
-$h$: 	& Bandwidth used in kernel estimation
-\end{tabular}\\
+They do state that for random-systematic transects, the methodology used by \textcite{micronesian}, that each transect line should be treated and analyzed as a cluster.
 
-For a typical individual plot:
+In discussing line transects, \textcite{barry2001} treat it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
+This would seem to (but does not explicitly) endorse the possibility of overlapping VCPs.
 
-\begin{tabular}{r l}
-$N$:	&	True number of animals in the plot (unknown)\\
-$m$:	&	Number of detected animals in the plot\\
-$g(r)$:	&	Radial detection function, $g(r)$=Pr(Animal is detected | Animal is at distance $r$)\\
-$w$:	&	Distance beyond which no animals can be detected (= radius of the plot)\\
-$p$:	& Mean detection probability within the plot\\
-$B$:	&   $B = 2/(pw^2)$\\
+One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously with some uniform density, $D$, throughout the space \textcite{ramsey1981}.
+(If the data is clustered we have other problems, not addressed here.) If if the uniform distribution is the case, then it would make sense that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
+This echoes Buckland’s statement a few paragraphs previous.
 
-\end{tabular}
+\textcite[244]{thompson2012} mentions (page 244) that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
+
+However, it is clear the literature is divided and unclear on whether or not our observation areas should overlap.
 
 \subsection{Assumptions}
-Buckland (2004) lists three primary assumptions:
+\textcite{buckland2001} lists three primary assumptions:
 \begin{enumerate}
-\item[i] Objects at the line or point are detected with certainty $g(0)=1$
-\item[ii] Objects are detected at their initial location
-\item[iii] Measurements are exact
+\item Objects at the line or point are detected with certainty $g(0)=1$
+\item Objects are detected at their initial location
+\item Measurements are exact
 \end{enumerate}
 
-There are additional assumptions listed by both Buckland and Ramsey and Scott (1979) and summarized by Quang (1993):
+There are additional assumptions listed by both Buckland and \textcite{ramsey1979} and summarized by \textcite{quang1993}:
 \begin{enumerate}
@@
 
-\subsubsection{Independence}
-Assumption \texttt{ix}, that the VCPs are placed randomly through the study area, would seem to implicitly state that their observational ranges might by chance overlap, meaning a single object of interest could be observed from more than one station.
-Thompson (p240) likewise indicates random placement of line transects which would seem to allow for a potential overlap of observation zones.
-
-Buckland et al (2001) also discuss the random placement of both line and point transects, without addressing the potential for sightings to overlap.
-Reynolds et al.
-(1980) state that the possibility of observing the same bird from two stations should be avoided.
-Buckland (2006) says simulations showed that analysis was robust to violations of independence (meaning overlap, or multiple observations on same object).
-
-On page 233, Buckland et al. (2001) discuss that each line or point must be ``randomly and independently located,'' with ``an equal probability of selection'' for all portions of the study area.
-
-On page 235, Buckland et al.
-states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet unsampled lines.''
-
-They do state that for random-systematic transects, the methodology used by Ramesy et al (1982), that each transect line should be treated and analyzed as a cluster.
-
-Barry and Welsh (2001) in discussing line transects, treat it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
-This would seem to (but does not explicitly) endorse the possibility of overlapping VCPs.
-
-One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously throughout the space [reference].
-(If the data is clustered we have other problems, not addressed here.) If that is the case, then it would make sense thought-level idea that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
-This echoes Buckland’s statement a few paragraphs previous.
-
-Thomson mentions (page 244) that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
-
 \section{Simulation}
@@
 		\caption{Totally Random Layout}
-		\label{fig:structured}
+		\label{fig:random}
 	\end{subfigure}
@@
 		\caption{A Random-Systematic Layout}
-		\label{fig:structured}
+		\label{fig:transect1}
 	\end{subfigure}
@@
 		\caption{Another Random-Systematic Layout}
-		\label{fig:structured}
+		\label{fig:transect2}
 	\end{subfigure}
+	\label{fig:layouts}
 \end{figure}
@@
 		\caption{Area of rectangle vs circle at same distance from point/line}
-		\label{fig:structured}
 	\end{subfigure}
@@
 		\caption{Estimated detection numbers for rectangle (solid) vs circle (dashed)}
-		\label{fig:structured}
 	\end{subfigure}
@@
 
+\subsection{Notation}
+This notation is borrowed from Quang (1993) following Buckland (1987)
+
+\begin{tabular}{r l}
+$t$:	&	Number of plots surveyed\\
+$D$:	& Common density of animals in all plots\\
+$m_j$:	& Number of detected animals in the $j$th plot\\
+$\bar m$:	& Average number detected per plot\\
+$\mu$:	& Expected number of animals in a plot, $\mu = E(m_j)$\\
+$n$:	& Aggregate numbe of animals detected in all plots, $n=\sum_{j=1}^t m_j$\\
+$\lambda$:	&	Expected total number of detected animals in all plots, $\lambda=E(n)=t\mu$\\
+$R_i$:	& Distance from station to the $i$th detected animal\\
+$R_1,\cdots,R_n$: & Detected distances pooled from all plots\\
+$s$:	& Sample standard deviation of $R_1,\cdots,R_n$\\
+$f(r)$: & Common probability density function (pdf) of the detected distances, $R_1,\cdots,R_n$\\
+$h$: 	& Bandwidth used in kernel estimation
+\end{tabular}\\
+
+For a typical individual plot:
+
+\begin{tabular}{r l}
+$N$:	&	True number of animals in the plot (unknown)\\
+$m$:	&	Number of detected animals in the plot\\
+$g(r)$:	&	Radial detection function, $g(r)$=Pr(Animal is detected | Animal is at distance $r$)\\
+$w$:	&	Distance beyond which no animals can be detected (= radius of the plot)\\
+$p$:	& Mean detection probability within the plot\\
+$B$:	&   $B = 2/(pw^2)$\\
+
+\end{tabular}
+
+\section{Bibliography}
+
+\printbibliography
+%\bibliographystyle{plain}
+%\bibliography{VCPbibliography}
 \end{document}

### 2014-05-27 code/ExampleGraphs.R

@@ -1,5 +1,9 @@
 # Example Graphs
-
+OSUOrange <- rgb(243/255,115/255,33/255) #http://oregonstate.edu/brand/color-palette
+OSURed <- rgb(192/255,49/225,26/255)
+OSUBlue1 <- rgb(93/255, 135/255, 161/255)
+OSUBlue2 <- rgb(156/255, 197/255, 202/255)
+OSUMedBrn <- rgb(176/255,96/255,16/255)
 # requires dplyr
 # requires get.Coords from VCP_Utilities.R
 

### 2014-05-27 fiddling/moveSimTest.R

@@ -117,10 +117,15 @@ ggsave("images/MovementSim2.pdf", width=10, height=6)
 #hmm. not really the drastic differences I was expecting.
 #after tweaking movement, it looks better
 
-D.hat.df %.% group_by(Layout, MoveType) %.% summarise(avg=mean(Dhat), n.In=sum(inCI))
+move.summ <- D.hat.df %.% group_by(Layout, MoveType) %.% summarise(avg=mean(Dhat), n.In=sum(inCI))
 
 # the lack of difference between PermMove and TempMove makes sense. The majority of the movement (if that's what it was) was within the first 100 m. The stations are 150 m apart. It's not going to affect those quite as much, unless you have REALLY closely spaces VCP.
 
+library(stargazer)
+stargazer(move.summ)
+
+library(xtable)
+xtable(move.summ)
 
 ## ------ Set up for by-hand Method test
 xy.vcp <- VCP.structuredLayout()

### 2014-05-27 report/draft2.tex
continuing to work on report. some modificaitons to graphs, mainly theme_bw(18) changes.

--- 2014-05-27 report/draft.tex
@@
 With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground.
-This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for (moving) wildlife and watch where they are walking \cite{ramsey1979}.
-VCP have the added benefit of allowing the wildlife to settle.
-Animals may react to an observer moving through the area by hiding or moving away.
-If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal, animal behavior may be more natural and less influenced by the presence of the observer.
+This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for potentially moving wildlife at the same time as navigating potentially treacherous terrain.
+\cite{ramsey1979}.
+
+With line transects, animals may react to an observer moving through the area by hiding or moving away.
+If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal (possible with point transects), animal behavior may be more natural and less influenced by the presence of the observer.
 
@@
 
-\subsection{Micronesian Bird Data}
+Objects closer to the observer are typically easier to spot than objects further away.
+We can imagine that there is an area of (nearly) perfect detectability near the observer, that drops off the further away from the observer the object is.
+This detectability function is labeled $g(r)$, with $r$ being the observation distance.
+If no other covariates are tracked, we end up with a count of objects observed, and the distances from the line or point at which they were observed.
+
+It is preferable for the detection function, $g(r)$, to have a "shoulder" at 0, where the probability of detection is 1.
+A half-normal curve, scaled such that $g(0)=1$ is an example \parencite{buckland2001}.
+
+An estimate of population density is a count of objects observed, divided by the area in which they were observed, times the probability of observing the object at that distance.
+
+$\hat{D}=E[n]/(Area)*P(observing object|distance r)$
+
+The difficulty in distance sampling comes from estimating $g(r)$.
+Many methods have been proposed, all with their pros and cons.
+This paper concentrates on the kernel method, as detailed by \textcite{quang1993}.
+This is a non-parametric method, using the kernel of a normal distribution to obtain a kind of ``average observation distance'' to be used in the calculation of the effective area surveyed for our population density estimate.
+
+\section{Micronesian Bird Data}
 \begin{figure}
@@
 Since the stations on a transect were (generally) visited in sequence, this means it is possible a bird could have been observed from more than one station.
+How does this affect the independence of our observations?
+Will it make a difference in our final population density estimates?
 
-\subsubsection{Independence}
+\subsection{Independence}
 Both \textcite{ramsey1979,buckland1987} discuss the assumption that the VCPs are placed randomly through the study area.
@@
 
-However, it is clear the literature is divided and unclear on whether or not our observation areas should overlap.
+It is clear the literature is divided on whether or not our observation areas should overlap. VCP layouts similar to \textcite{micronesian} ``random-systematic'' layout are common in the literature.
 
@@
 \begin{enumerate}
-\item[iv] Sightings of animals are independent events
-\item[v] at least one animal is detected $n > 0$
-\item[vi] detectability is the same for all animals at distance r and in all directions.
-\item[vii] no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
-\item[viii] the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
-\item[ix] the stations are chosen at random in the study area
+	\setcounter{enumi}{3}
+	\item Sightings of animals are independent events
+	\item at least one animal is detected $n > 0$
+	\item detectability is the same for all animals at distance r and in all directions.
+	\item no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
+	\item the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
+	\item the stations are chosen at random in the study area
 \end{enumerate}
 
-Ramsey 1979
-Uniform distribution, random process distribution, independent distribution,
+%Ramsey 1979: Uniform distribution, random process distribution, independent distribution,
 
-Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE
-
-\subsection{Broad Strokes}
-In general, an estimate of population density is a count of objects observed divided by the area in which they were observed times the probability of observing the object. (Ramsey, Scott 1979)
-
-Equation 1: $D = E[n]/(A_r)(P_r)$
-
-In practice, the expected value of n is taken to be n, the number of objects observed. The estimation of the denominator is where things get tricky.
-
-If we combine $(A_r)(P_r)$ into an effective area we get:
-
-Equation 2: $A~_R = 2(pi) int?0?R g(y)ydy$
-
-$G(r)$ is the detectability function, and represents the probability of detection at distance $r$.
-It is often scaled so that $g(0)=1$.
-(In terms of area of a circle, it is obvious that the $integral{g(y)ydy}$ represents $r^2$ in the $2pi r^2$ equation.)
-
-Much of the uncertainty in VCP estimation comes from how the effective area is estimated.
-$g(y)$ is often modeled as a parametric class (exponential, half-normal, hazard function) which is then scaled so that $g(0)=1$.
-In addition to guessing which function to use, we must also estimate the parameters.
-
-It is plausible to think that as objects are further away from the observer, they are more difficult to detect, and therefore, we would have a greater probability of missing observations that were further away.
-So, in calculating $g(y)$, do we assume perfect detection out to the furthest point of observation?
-Or do we truncate our observations at some distance closer to the observer?
-Ramsey \& Scott (1979) refer to these as the effective radius ($\rho$) and the basal radius ($r$) respectively.
-
-The problem lies in that in practice, we don't know where the basal radius is, and it must be estimated from our data.
+% Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE
 
@@
 \subsubsection{Bias \& Consistency}
-Buckland et al. (2001) and Ramsey \& Scott (1970) both stated that if our detectability estimate $g(z)=$1 out to $r$, then $D$ is unbiased.
+\textcite{buckland2001,ramsey1979} Buckland et al. (2001) and Ramsey \& Scott (1970) both stated that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
 
-Barry and Walsh (2001) compared design-based and  model-based estimates of density and variance.
+\textcite{barry2001} Barry and Walsh (2001) compared design-based and  model-based estimates of density and variance.
 (Their focus was line transects, but they state it is analogous to point transects and the conclusions should carry across.) They concluded that model-based estimates were unbiased if and only if the assumptions of objects being independent and uniformly located held.
@@
 
-Roeder, Dennis, \& Garton (1987) use simulations to demonstrate that no method is completely applicable in 100\% of situations, and will go sideways at some point, depending on the specific conditions.
+\textcite{roeder1987} Roeder, Dennis, \& Garton (1987) use simulations to demonstrate that no method is completely applicable in 100\% of situations, and will go sideways at some point, depending on the specific conditions.
 
 The overall conclusion is that, in theory, density estimates using VCPs are unbiased and (reasonably) consistent, but that the assumptions necessary for the theory to hold are untenable.
-
-As stated in the section on choosing $g(r)$, it will vary depending on the specific situation, and it is up to the scientist and statistician to understand the data and choose the best approach.
+The appropriate estimation of $g(r)$ will vary depending on the specific situation, and it is up to the scientist and statistician to understand the data and choose the best approach.
 
@@
 As seen in the previous section, the issue of independence of the point transects does not have a consensus in the literature.
-If independence is required for proper estimation, then the random-systematic placement method should not be used.
+If independence is required for proper estimation, then the random-systematic placement (hereafter, ``transect'') layout should not be used.
 
@@
 \subsection{VCP Arrangement}
-Simulation will address three options for VCP arrangement:
+\label{sec:layout}
+To examine the effect of VCP overlap, we will look at three options for VCP arrangement:
 \begin{itemize}
@@
 	\item Random: VCPs placed completely at random
-\item Random-Systematic: Two transects of 18 stations each, placed according to Ramsey et al's design: 150 m between stations, 2 km distance between parallel transects.
+	\item Transect: Two transects of 18 stations each, placed according to \textcite{micronesian} design: 150 m between stations, 2 km distance between parallel transects.
 \end{itemize}
-
-Figures 2 and 3 illustrate the first two arrangements.
-Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota.
-The surveyed area of the Palie region totaled 9.41 km$^2$ during the original survey, and contained two transects of 16 and 17 stations each, 33 stations total (Ramsey et al 1982).
-
-As seen in Figure 1, there is very little detection happening beyond 500 m. There are 15 observations beyond 500 m, which is 0.9\% of the data.
-
-Treating 500 m as our truncation distance $w$, the study area in the simulation has been raised to 36 km$^2$ to allow for stations where the observation area did not overlap.
-An area of 49 km$^2$ is generated with the given density, then truncated to 36 km$^2$ to prevent any edge effect from the random generation of points.
-The stations are placed a minimum of 500 km (0.5 on the graph) away from the edges to prevent border truncation.
-
-For the random-systematic, an angle was chosen between $0$ and $\pi$.
-A transect of 18 stations 150 m apart was constructed, and a second transect of 18 stations constructed running parallel.
-The transects were then randomly placed with in the graph, no closer to the edge than 500 km (0.5 on graph.) Figures 4 and 5 are two examples.
-
 \begin{figure}
@@
 \end{figure}
+Figures~\ref{fig:structured} and \ref{fig:random} illustrate the first two arrangements. Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota.
+
+As seen in Figure~\ref{fig:by5}, there is very little detection happening beyond 500 m. There are a total of 15 observations beyond 500 m, which is 0.9\% of the data.
+
+The surveyed area of the Palie region totaled 9.41 km$^2$ during the original survey.
+Treating 500 m as our truncation distance $w$, the study area in the simulation has been raised from 9.41 km$^2$ to 36 km$^2$ to allow for stations where the observation area did not overlap.
+An area of 49 km$^2$ is generated with the given density, then truncated to 36 km$^2$ to prevent any edge effect from the random generation of points.
+The stations are placed a minimum of 500 km (0.5 on the graph) away from the edges to prevent border truncation.
+
+The Palie region contained two transects of 16 and 17 stations each, 33 stations total \parencite{micronesian}.
+For the transect layout, an angle was chosen between $0$ and $\pi$.
+A transect of 18 stations 150 m apart was constructed, and a second transect of 18 stations constructed running parallel.
+The transects were then randomly placed with in the graph, no closer to the edge than 500 km (0.5 on graph.) This transect layout is illustrated in Figures~\ref{fig:transect1} and \ref{fig:transect2}.
+The total of 36 VCPs was carried over to the Structured and Random layouts, for an equivalent number of points.
 
 \section{Simulation Setup}
-\subsection{$g(0)$}
-\subsubsection{Half Normal}
+The first simulation compares the three layouts described in section~\ref{sec:layout} with two different options for $g(r)$, a half-normal detection probability function, and a detection probability estimated from the empirical observation distances.
+\subsection{Half Normal}
 For a maximum possible observation distance of 0.500 (500 m), the parameter for the half-normal detection function is $\theta = 8.77$, with a scale parameter $\delta = 0.179$.
@@
 
-If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where P(X > 3.5) < 0.001) then theta is:
+If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where $P(X > 3.5) < 0.001$) then $\theta$ is:
 
@@
 
-\subsubsection{Empirical Observations}
+Half-normal density functions were done in R with the \texttt{fdrtool} package.
+
+\subsection{Empirical Observations}
 \begin{figure}
-
 	\includegraphics[width=\textwidth]{../images/loess.pdf}
 	\caption{Empirical Observation Distances from Original data in meters, with LOESS regression lines. Bin width=5 m\label{fig:by5}}
-
 \end{figure}
 
-We can see in Figure \ref{fig:by5} that above 100 m, observations were commonly heaped into 10 or 25 m distances, with the latter being more heavily used.
-Above 200 m, 50 m heaps were common.
-To estimate a density function, I used a loess regression with a span of 0.2 on the counts from 5 meter bins of the empirical data.
-This is the blue line in Figure \ref{fig:by5}.
+We can see in Figure \ref{fig:by5} that above 100 m, observations were commonly heaped into 10 or 25 m distances, with the latter being more heavily used. Above 200 m, 50 m heaps were common.
 
@@
 Using the bin midpoints, the density estimate was multiplied by 5 to get the density value for that bin.
-A LOESS regression was run with the midpoint as $x$ and $y=density*5$, using the \texttt{loess()} package in R.
+A LOESS regression was run with the midpoint as $x$ and $y=density*5$, using the \texttt{loess()} package in R with a span of 0.2.
+This is the solid line in Figure~\ref{fig:by5}.
 
 This gives us an object that can be used with \texttt{predict()} to get density estimates for specific distances.
-The values generated by feeding our midpoints into our prediction function are represented by the red dashed line in Figure \ref{fig:by5}.
+The values generated by feeding our midpoints into our prediction function are represented by the dashed line in Figure~\ref{fig:by5}.
 
@@
 $\delta_{EMP}=\frac{1}{max(predicted)}=40.5780$
-
-\texttt{I know I read about how to deal with the random-systematic method, but now, of course, can't find it.
-With Line transect sampling, they essentially treat the data as if it came from one long transect, so that's what I'm going to do until I can find another reference.}
 
@@
 Figure \ref{fig:sim1} shows the results of 1000 simulations, each analyzed with Empirical and Half-normal detection function, $g(x)$, and using each of the three layouts, Structured, Random, and Transect.
-All $\hat{D}$  were done using the kernel method with a normal kernel (Quang 1993).
+All $\hat{D}$  were done using the kernel method with a normal kernel \parencite{quang1993}.
 
@@
 \texttt{TO DO: look at 3 moved categories for the 3 layouts}
+
+\subsection{Movement or No Movement}
+\begin{figure}
+
+	\includegraphics[width=\textwidth]{../images/MovementSim2.pdf}
+	\caption{1000 Simulations, Movement, Temporary Movement and Permanent Movement for three different layouts. \label{fig:sim2}}
+
+\end{figure}
+In the Figure~\ref{fig:82dist}, the histogram of empirical observation distances, there appears to be evidence of movement away from the observer.
+To see if the movement would have an effect on the independence of our VCPs, depending on placement, a second simulation was run.
+
+As with the previous simulation, the same three layout types were used. Three movement classes were defined:
+\begin{enumerate}
+\item No Movement: Same as previous simulation, objects stayed at original point
+\item Temporary Movement: Objects within 100 m moved directly away from observer, with more movement if they started closer to observer. They were reset to original location before moving to next VCP
+\item Permanent Movement: As with temporary movement, only not reset before moving on to next VCP.
+\end{enumerate}
+
+Figure~\ref{fig:sim2} and Table~\ref{table:sim2} illustrate the results.
+As with the previous simulation, the results are visually very similar for all three layout types.
+All are biased low, with more bias in the simulations with movement.
+Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
+
+For this simulation, the movement was simulated to happen within the first 100 m from the observer (0.1 on graph), as visually indicated by Figure~\ref{fig:82dist}.
+
+\begin{table}
+	\caption{Half-normal detection function, three layouts, three movement types. 1000 simulations.}
+	\begin{tabular}{ r| r r| r r| r r|}
+
+		& \multicolumn{2}{|c|}{No Movement}	& \multicolumn{2}{|c|}{Compounded Mvmt}	& \multicolumn{2}{|c|}{Temporary Mvmt}\\ 
+ \hline \hline
+
+ Layout		& Mean $\hat{D}$	& \% Capture & Mean $\hat{D}$ & \% Capture & Mean $\hat{D}$ & \% Capture	\\ \hline \hline
+ Random		& 17.86 			& 94.2\% 		& 11.86	& 59.6\%	& 11.76	& 59.3\% \\
+ Structured	& 18.01 			& 94.3\% 		& 11.90 & 59.9\% 	& 11.74 & 59.5\% \\
+ Transect	& 17.90 			& 90.8\% 		& 12.42 & 65.0\% 	& 11.75 & 58.0\% \\ \hline
+
+	\end{tabular}
+	\label{table:sim2}
+\end{table}
+
+no small distances, so overestimating area surveyed compared to # ob
+
+we will be low if top small, bottom large.
 

### 2014-05-27 code/ExampleGraphs.R

@@ -11,7 +11,7 @@ library(dplyr)
 library(ggplot2)
 
 # get coords, using a density of 20 birds/km2
-x.y <- get.Coords( 20)
+x.y <- get.Coords(20)
 
 ## -------------- Structured VCP layout -------------- 
 # get the equi-spaced, independent VCP locations.
@@ -27,7 +27,7 @@ for(i in 1:36) {
 
 # generate and save plot
 # you will get warnings when running these, because it is trimming out points to get to the 0,6 bounds
-ggplot()+geom_point(data=x.y, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color="#D55E00", inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour="#D55E00", size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+ggplot()+geom_point(data=x.y, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color=OSUOrange, inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour=OSUOrange, size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
 
 ggsave("images/layout_structured.pdf", width=8, height=8)
 
@@ -46,7 +46,7 @@ for(i in 1:36) {
 }
 
 # generate and save plot
-ggplot()+geom_point(data=x.y, aes(X,Y))+geom_point(aes(x=stat.X, y=stat.Y), color="#D55E00", inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour="#D55E00", size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+ggplot()+geom_point(data=x.y, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color=OSUOrange, inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour=OSUOrange, size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
 
 ggsave("images/layout_random.pdf", width=8, height=8)
 
@@ -139,19 +139,20 @@ stat.df <- data.frame(
[…]

### 2014-05-27 fiddling/detection.R

@@ -92,7 +92,7 @@ hist.info<- hist(coki.82$Distance, breaks=my_breaks)
 
 # I like how this is looking, but may be aggregating too much
 my_breaks <- c(seq(0,100,20),seq(120,200,20), seq(250,700,50))
-h
+
 
 hist.info2<- hist(coki.82$Distance, breaks=seq(0,700,5), freq=FALSE)
 by5 <- data.frame(x=hist.info2$mids, y=hist.info2$density*5)
@@ -103,7 +103,10 @@ th.y <- predict(by5.lo,th.x)
 th.y[1] <- th.y[141] <- 0
 
 
-ggplot()+geom_histogram(data=by5, aes(x, y), stat="identity")+geom_smooth(data=by5, aes(x, y), method="loess", span=0.2, size=3)+geom_line(inherit.aes=FALSE, aes(x=th.x, y=th.y), color="red", linetype=2, size=2)+theme_bw(18)+xlab("Detection Distance in Meters")+ylab('Detection Density')
+ggplot()+geom_histogram(data=by5, aes(x, y), stat="identity")+geom_smooth(data=by5, aes(x, y), method="loess", span=0.2, size=3, color=OSURed)+geom_line(inherit.aes=FALSE, aes(x=th.x, y=th.y), color=OSUBlue2, linetype=2, size=1.5)+theme_bw(18)+xlab("Detection Distance in Meters")+ylab('Detection Density')
+
+
+ggsave("images/loess.pdf", width=8, height=4)
 
 which.max(th.y) # 19
 th.y[19] # 0.02464387

### 2014-05-28 report/draft2.tex
Finished project draft. still need to fix a couple graphics.

@@
 \usepackage{subcaption}
+\usepackage{textcomp}
 %\usepackage[round,authoryear]{natbib}
@@
 \begin{document}
+Ignore this page, it's so I can have a reference for the different citation formats and will be removed before turning in the paper.
+
 cite \cite{buckland2001} \cite{komodo2013} \\
@@
 Distance sampling, in the form of line and point transects (alternatively, Variable Circular Plots (VCP)), is a common estimation method.
-They are useful when populations are too small to be seen from the air (birds, rodents) or may be masked by the canopy (ground plants.)
+They are useful when individual objects are too small to be seen from the air or may be masked by the canopy (birds, small animals, ground plants.)
 
@@
 \cite{burnham1980} is a monograph on the theory and application of line transect sampling.
-Non-parametric methods have been examined by \textcite{quang1993,mack1998}.
+Non-parametric methods have been examined by \textcite{quang1993,mack1998}, among others.
 Distance sampling continues to be used and modified, combined with mark-recapture methods \parencite{laake2011}, extended for use with underwater acoustics for estimates of krill populations \parencite{krill2011}, and a study by \textcite{camera2011} combines distance sampling with camera traps to estimate populations.
@@
 With Line Transects, a (straight) line is randomly placed in the study area, then walked by the observer, with any objects of interest noted and the distance perpendicular to the transect recorded.
-With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground.
-This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for potentially moving wildlife at the same time as navigating potentially treacherous terrain.
+With point transects, the observer stands in one spot, recording the direct line distance from herself to the object of interest as if it was projected on the ground, within the 360\textdegree.
+This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for moving wildlife at the same time as navigating potentially treacherous terrain.
 \cite{ramsey1979}.
@@
 With line transects, animals may react to an observer moving through the area by hiding or moving away.
-If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal (possible with point transects), animal behavior may be more natural and less influenced by the presence of the observer.
+If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal--possible with point transects--animal behavior may be more natural and less influenced by the presence of the observer.
 
@@
 Objects closer to the observer are typically easier to spot than objects further away.
-We can imagine that there is an area of (nearly) perfect detectability near the observer, that drops off the further away from the observer the object is.
+Considering the probability of detecting an object, we can imagine that there is an area of (nearly) perfect detectability near the observer, that drops off the further away from the observer the object is.
 This detectability function is labeled $g(r)$, with $r$ being the observation distance.
@@
 
-Each of the five islands was divided into regions.
+Each of the five islands included in the survey was divided into regions.
 Initial transects were given random positions and angles, with subsequent transects placed parallel, 2km apart (the ``random-systematic'' method).
 Stations were placed every 150m along the transects.
-Two observers, standing 20 m apart, would wait several minutes for the wildlife to quiet down after their arrival, and then spend 8 minutes noting any birds they could observe either visually or aurally.
+Two observers, standing 20 m apart, would wait several minutes for the wildlife to quiet down after their arrival, and then spend 8 minutes noting any birds they could detect either visually or aurally.
 Distance and species were noted, as well as other covariates such as foliage density, and weather conditions.
@@
 Lastly, note that the detection distances are high between 100--200 m, and recall that the stations were placed 150 m apart.
-Since the stations on a transect were (generally) visited in sequence, this means it is possible a bird could have been observed from more than one station.
+Since the stations on a transect were generally visited in sequence, this means it is possible a bird could have been observed from more than one station.
 How does this affect the independence of our observations?
@@
 
-In discussing line transects, \textcite{barry2001} treat it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
+In discussing line transects, \textcite{barry2001} treats it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
 This would seem to (but does not explicitly) endorse the possibility of overlapping VCPs.
@@
 One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously with some uniform density, $D$, throughout the space \textcite{ramsey1981}.
-(If the data is clustered we have other problems, not addressed here.) If if the uniform distribution is the case, then it would make sense that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
+(If the data is clustered we have other problems, not addressed here.) If the uniform distribution is the case, then it would make sense that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
 This echoes Buckland’s statement a few paragraphs previous.
 
-\textcite[244]{thompson2012} mentions (page 244) that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
+\textcite[244]{thompson2012} mentions that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
 
@@
 
-\subsection{Assumptions}
-\textcite{buckland2001} lists three primary assumptions:
-\begin{enumerate}
-	\item Objects at the line or point are detected with certainty $g(0)=1$
-	\item Objects are detected at their initial location
-	\item Measurements are exact
-\end{enumerate}
-
-There are additional assumptions listed by both Buckland and \textcite{ramsey1979} and summarized by \textcite{quang1993}:
-\begin{enumerate}
-	\setcounter{enumi}{3}
-	\item Sightings of animals are independent events
-	\item at least one animal is detected $n > 0$
-	\item detectability is the same for all animals at distance r and in all directions.
-	\item no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
-	\item the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
-	\item the stations are chosen at random in the study area
-\end{enumerate}
-
-%Ramsey 1979: Uniform distribution, random process distribution, independent distribution,
-
-% Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE
-
-\subsection{Properties}
-\subsubsection{Bias \& Consistency}
+\subsection{Bias \& Consistency}
 \textcite{buckland2001,ramsey1979} Buckland et al. (2001) and Ramsey \& Scott (1970) both stated that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
@@
 As seen in the previous section, the issue of independence of the point transects does not have a consensus in the literature.
-If independence is required for proper estimation, then the random-systematic placement (hereafter, ``transect'') layout should not be used.
+If independence is required for proper estimation, then the random-systematic placement (hereafter, ``transect'') layout may require some adjustment of the estimate to bring it in line with survey layouts where the observations areas do not overlap.
 
-Through simulation, we can explore what happens under three different VCP placement schemes, and compare the resulting density estimates to their known values to evaluate performance.
+Through simulation, we can explore what happens under three different VCP placement schemes, detection densities, and object movement, and compare the resulting density estimates to their known values to evaluate the performance of the estimators.
 
@@
 		\includegraphics[width=\textwidth]{../images/layout_rand-sys-4.pdf}
-		\caption{A Random-Systematic Layout}
+		\caption{A Transect Layout}
 		\label{fig:transect1}
@@
 		\includegraphics[width=\textwidth]{../images/layout_rand-sys-6.pdf}
-		\caption{Another Random-Systematic Layout}
+		\caption{Another Transect Layout}
 		\label{fig:transect2}
@@
 
-As seen in Figure~\ref{fig:by5}, there is very little detection happening beyond 500 m. There are a total of 15 observations beyond 500 m, which is 0.9\% of the data.
+As seen in Figure~\ref{fig:82dist}, there is very little detection happening beyond 500 m. There are a total of 15 observations beyond 500 m, which is 0.9\% of the data.
 
@@
 
-\section{Simulation Setup}
+\subsection{$g(r)$ Options}
 The first simulation compares the three layouts described in section~\ref{sec:layout} with two different options for $g(r)$, a half-normal detection probability function, and a detection probability estimated from the empirical observation distances.
-\subsection{Half Normal}
+\subsubsection{Half Normal}
 For a maximum possible observation distance of 0.500 (500 m), the parameter for the half-normal detection function is $\theta = 8.77$, with a scale parameter $\delta = 0.179$.
@@
 
-Half-normal density functions were done in R with the \texttt{fdrtool} package.
+Half-normal density functions were executed in R with the \texttt{fdrtool} package.
 
-\subsection{Empirical Observations}
+\subsubsection{Empirical Observations}
 \begin{figure}
@@
 
+\subsubsection{Movement}
+In the Figure~\ref{fig:82dist}, the histogram of empirical observation distances, there appears to be evidence of movement away from the observer.
+To see if the movement would have an effect on the independence of our VCPs, depending on placement, a second simulation was run.
+
+Three movement classes were defined:
+\begin{enumerate}
+\item No Movement: Objects stayed at original point.
+\item Temporary Movement: Objects within 100 m had a random chance to move directly away from observer.
+Movement was larger if they started closer to observer.
+They were reset to original location before moving to next VCP
+\item Permanent Movement: As with temporary movement, only not reset before moving on to next VCP. (Movement could compound if object was within the maximum observation radius for more than one VCP.)
+\end{enumerate}
+
+For this simulation, the movement was simulated to happen within the first 100 m from the observer (0.1 on graph), as visually indicated by Figure~\ref{fig:82dist}.
+The simulation was not intended to exactly model the movement observed by \textcite{micronesian}, but to get a general sense of what might happen to $\hat{D}$ in each of the two settings.
+
 \section{Results}
-\subsection{Half-normal vs. Empirical $g(x)$}
+\subsection{Half-normal vs. Empirical $g(r)$}
 \begin{figure}
@@
 	\includegraphics[width=\textwidth]{../images/Emp_Vs_Hnorm.pdf}
-	\caption{1000 Simulations, Empirical vs Half-normal Detection function, for three different layouts. \label{fig:sim1}}
+	\caption{1000 Simulations, Empirical vs Half-normal Detection function, for three different layouts. Dashed line equals true value of 20 objects per km$^2$\label{fig:sim1}}
 
@@
 We also see bias in the estimates using the half-normal function for $g(x)$, but it is much less severe than in the Empirical data.
+\textcite{quang1993} does provide a bias-correction method, but it was not used for this simulation.
 
 Addressing our primary question of interest, does the overlap of the VCP affect the population density estimate due to violation of independence, the answer would appear to be no.
-Looking at the simulations with the Half-normal detection function, the histograms are fairly similar for the three layouts.
+Looking at the simulations with the Half-normal detection function, the histograms are fairly similar across the three layouts.
 
-\begin{table}
-	\caption{Empirical vs. Half-normal detection function, across all three layouts. 1000 simulations.}
+Table~\ref{table:sim1} supports what we observe in the histogram: the means vary between our two $g(r)$ options, but are approximately the same within the each option, regardless of which VCP layout we have chosen.
+\begin{table}[h]
+	\caption{Empirical vs. Half-normal detection function, across all three layouts. 1000 simulations. True density 20/km$^2$}
 	\begin{tabular}{ r r r r r r r r r}
@@
 
-\texttt{TO DO: along with d.hat, get the variance estimate for d.hat, and return T/F if the true value was contained in the interval.}
-
-\texttt{TO DO: look at 3 moved categories for the 3 layouts}
+\texttt{TO DO: return T/F if the true value was contained in the interval. redo graph to match others.}
 
@@
 \end{figure}
-In the Figure~\ref{fig:82dist}, the histogram of empirical observation distances, there appears to be evidence of movement away from the observer.
-To see if the movement would have an effect on the independence of our VCPs, depending on placement, a second simulation was run.
-
-As with the previous simulation, the same three layout types were used. Three movement classes were defined:
-\begin{enumerate}
-\item No Movement: Same as previous simulation, objects stayed at original point
-\item Temporary Movement: Objects within 100 m moved directly away from observer, with more movement if they started closer to observer. They were reset to original location before moving to next VCP
-\item Permanent Movement: As with temporary movement, only not reset before moving on to next VCP.
-\end{enumerate}
-
-Figure~\ref{fig:sim2} and Table~\ref{table:sim2} illustrate the results.
-As with the previous simulation, the results are visually very similar for all three layout types.
-All are biased low, with more bias in the simulations with movement.
-Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
-
-For this simulation, the movement was simulated to happen within the first 100 m from the observer (0.1 on graph), as visually indicated by Figure~\ref{fig:82dist}.
-
 \begin{table}
@@
 
-no small distances, so overestimating area surveyed compared to # ob
+Figure~\ref{fig:sim2} and Table~\ref{table:sim2} illustrate the results of our movement type simulation. For this run, only the half-normal detection function was used.
 
-we will be low if top small, bottom large.
+Within a column, the results are fairly similar in shape and spread.
+We see the most obvious differences here between columns, indicating that the type of movement causes more bias in our estimate than the potential overlap of our observation areas.
+
+Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
+Table~\ref{table:sim2} shows that the 95\% confidence intervals \parencite{quang1993} capture the true mean about 59\% of the time, no matter which movement type, Temporary or Compounded.
+
+The similarity for the Structured and Random layouts can be explained by a general lack of compounded movement.
+For the Structured layout, the animals would only move once, additional movement would not be triggered by an overlapping VCP, so the Compounded Movement setting would have the same end result as the Temporary Movement setting.
+For the Random VCP layout we would expect there to be a larger effect from the Compounded movement, because we will have VCPs whose observation distances overlap.
+
+Looking at Figure~\ref{fig:random}, which is just one possible random layout, there are 8 pairs of VCP where the circles representing the 200 m mark overlap, and 3 where they overlap to any significant degree.
+With most of the movement happening within that first 100 m, we would expect the estimate to be fairly similar to the Structured layout setting, if this random layout is typical of those generated.
+
+\texttt{TODO: Try to xplain the .50 jump in the estimate for transect/compounded. No theory or story atm.}
+
+%NOTE: no small distances, so overestimating area surveyed compared to \# ob
+%NOTE: we will be low if top small, bottom large.
 
@@
 To test the effectiveness of my movement algorithm, I compared the detection distances from the data where the objects moved, to the detection distances from the same data, assuming they stayed still.
-The movement data, (left panel, Figure \#) is squashed more towards 0.1 Km than the still data (right panel).
+The movement data, (left panel, Figure~\ref{fig:moveTest}) is squashed more towards 0.1 Km than the still data (right panel).
 But there's a problem.
@@
 
-\texttt{TO DO: Images}
-
-Figure \ref{fig:simDist} shows the distances from 36 randomly placed VCPs to the objects within the max observation distance.
-As we can see, there are simply fewer objects close to the observer, which increases as the distance moves away.
-(Rephrase) This is an artifact of the VCP being a circle.
-
-\texttt{TO DO: Same graph for Line transect}
-
-\begin{figure}
+\begin{figure}[h]
 	\center
-	\includegraphics[width=0.5\textwidth]{../images/simulated_bird_distances.pdf}
-	\caption{Simulated bird distances from center point of 36 randomly placed VCPs.\label{fig:simDist}}
-
+	\includegraphics[width=\textwidth]{../images/movementTest.pdf}
+	\caption{Simulated bird distances from center point of 36 randomly placed VCPs.\label{fig:moveTest}}
 \end{figure}
 
-With a line transect, as we move away from the line, the distance "surveyed" increases linearly. With a point transect, as we move away from the point, the distance surveyed increases exponentially.
+The problem is because circles are different than rectangles.
 
-\texttt{TO DO: Equations showing what happens when you compare distance at r vs distance at r+x}
+With a line transect, as we move away from the line, the distance ``surveyed" increases linearly.
+With a point transect, as we move away from the point, the distance surveyed increases exponentially.
+Figure~\ref{fig:perfectArea} shows the relationship for a rectangle and circle with the same area.
+
+Figure~\ref{fig:perfectCount} shows the difference in the expected number of observed objects at distance $r$ given a line transect vs.
+a VCP.
+For the VCP, Area was calculated as $2*pi*r$.
+For the line transect, Area was calculated as $(2r)*L$, with $L=0.785$ chosen so that the area of the circle and rectangle were equal when $r=0.5$.
+
+$E[observed] = Area*Density*P(object observed | distance r)$
+
+My conclusion from this is that for a VCP survey, even if everything is perfect, you will not see the nice shouldered layout of your empirical detection distances.
+\textcite{ramsey1981} speaks about this somewhat, when comparing the ``CUM-D'' curve, or the plot of the cumulative number of observations at distance $r$ for both line and point transects.
+However, almost all the illustrations on how your collected data should look, shows data that echoes the detection probability curve, with the half-bell-shape with a nice shoulder.
+Since this shape is impossible, I think some better examples are in order.
+
+Better mathematicians than I have worked on distance sampling, and this is why in analyzing point transect data we take the derivative of $g(r)$, to adjust for this difference in how area is calculated.
+
+%Thought - is this why we use h(0) for the point transects? need to verify, h(r) = f'(r)??  budkland 1987 paper has h(r) = f(r)/r, need to check teh book.  - yes.
 
@@
 	\centering
-	\caption{Squares (Line Transect, solid) vs. Circles (Point Transect, dashed)}
+	\caption{Rectangles (Line Transect, solid) vs. Circles (Point Transect, dashed)}
 	\begin{subfigure}[b]{0.45\textwidth}
@@
 		\caption{Area of rectangle vs circle at same distance from point/line}
+		\label{fig:perfectArea}
 	\end{subfigure}
@@
 		\caption{Estimated detection numbers for rectangle (solid) vs circle (dashed)}
+		\label{fig:perfectCount}
 	\end{subfigure}
-
 \end{figure}
 
-Thought - is this why we use h(0) for the point transects? need to verify, h(r) = f'(r)??  budkland 1987 paper has h(r) = f(r)/r, need to check teh book.
-
-\subsection{Notation}
-This notation is borrowed from Quang (1993) following Buckland (1987)
-
-\begin{tabular}{r l}
-$t$:	&	Number of plots surveyed\\
-$D$:	& Common density of animals in all plots\\
-$m_j$:	& Number of detected animals in the $j$th plot\\
-$\bar m$:	& Average number detected per plot\\
-$\mu$:	& Expected number of animals in a plot, $\mu = E(m_j)$\\
-$n$:	& Aggregate numbe of animals detected in all plots, $n=\sum_{j=1}^t m_j$\\
-$\lambda$:	&	Expected total number of detected animals in all plots, $\lambda=E(n)=t\mu$\\
-$R_i$:	& Distance from station to the $i$th detected animal\\
-$R_1,\cdots,R_n$: & Detected distances pooled from all plots\\
-$s$:	& Sample standard deviation of $R_1,\cdots,R_n$\\
-$f(r)$: & Common probability density function (pdf) of the detected distances, $R_1,\cdots,R_n$\\
-$h$: 	& Bandwidth used in kernel estimation
-\end{tabular}\\
-
-For a typical individual plot:
-
-\begin{tabular}{r l}
-$N$:	&	True number of animals in the plot (unknown)\\
-$m$:	&	Number of detected animals in the plot\\
-$g(r)$:	&	Radial detection function, $g(r)$=Pr(Animal is detected | Animal is at distance $r$)\\
-$w$:	&	Distance beyond which no animals can be detected (= radius of the plot)\\
-$p$:	& Mean detection probability within the plot\\
-$B$:	&   $B = 2/(pw^2)$\\
-
-\end{tabular}
+\section{Conclusions}
+If all else is held equal, VCP layout does not seem to play a large role in density estimate using the kernel method as described by \textcite{quang1993}.
+Violations of the expected detection probability curve, or by movement of the objects away from the observer play more of a role in biasing the resulting estimate.
 

### 2014-05-28 code/ExampleGraphs.R

@@ -166,7 +166,7 @@ th.y <- predict(by5.lo,th.x)
 th.y[1] <- th.y[141] <- 0
 
 
-ggplot()+geom_histogram(data=by5, aes(x, y), stat="identity")+geom_smooth(data=by5, aes(x, y), method="loess", span=0.2, size=3)+geom_line(inherit.aes=FALSE, aes(x=th.x, y=th.y), color="red", linetype=2, size=2)+theme_bw(18)+xlab("Detection Distance in Meters")+ylab('Detection Density')
+ggplot()+geom_histogram(data=by5, aes(x, y), stat="identity", fill=OSUOrange)+geom_smooth(data=by5, aes(x, y), method="loess", span=0.2, size=3, color=OSURed)+geom_line(inherit.aes=FALSE, aes(x=th.x, y=th.y), color=OSUBlue2, linetype=2, size=2)+theme_bw(18)+xlab("Detection Distance in Meters")+ylab('Detection Density')
 
 ggsave("images/loess.pdf", width=8, height=4)
 

### 2014-05-28 fiddling/Vectors.R

@@ -181,6 +181,23 @@ abline(a=.05, b=-.05/.150)
 
 # the hell? the "still" distances also have a drop close to the point
 
+Rj <- c(ddist.mov, ddist.still)
+cat <- c(rep("Move", 4598),
+         rep("Still", 4610)
+  )
+
+Movement <- data.frame(Rj, cat)
+
+ggplot(data=Movement, aes(x=Rj))+geom_histogram(aes(y=..count..), fill=OSUOrange, colour="white")+facet_grid(.~cat)+theme_bw(18)+ylab("Count of Observed")+xlab("Observation Distance 0.1=100 m")
+
+ggsave("images/movementTest.pdf", height=4, width=8)
+
+
+
+
+
+#-------------------------------------------------
+
 xy.vcp <- VCP.transectLayout()
 raw.Rj <- data.frame()
 xy.objects <- get.Coords(20)

### 2014-05-28 report/draft.tex
don't know what i changed.

@@
 This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for (moving) wildlife and watch where they are walking \cite{ramsey1979}.
-VCP have the added benefit of allowing the wildlife to settle.
-Animals may react to an observer moving through the area by hiding or moving away.
-If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal, animal behavior may be more natural and less influenced by the presence of the observer.
+
+With line transects, animals may react to an observer moving through the area by hiding or moving away.
+If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal (possible with point transects), animal behavior may be more natural and less influenced by the presence of the observer.
 
 Observations of objects can take the form of visual sightings or detection of auditory cues, or one confirmed by the other, as long as an accurate distance can be measured.
+
+Wrong Place:  Using halfnormal because it fits the shoulder criteria, and can be easily scaled to our truncation distance through it's parameter.
 
@@
 
-\subsection{Micronesian Bird Data}
+\section{Micronesian Bird Data}
 \begin{figure}
@@
 \begin{enumerate}
-\item[iv] Sightings of animals are independent events
-\item[v] at least one animal is detected $n > 0$
-\item[vi] detectability is the same for all animals at distance r and in all directions.
-\item[vii] no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
-\item[viii] the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
-\item[ix] the stations are chosen at random in the study area
+	\setcounter{enumi}{3}
+	\item Sightings of animals are independent events
+	\item at least one animal is detected $n > 0$
+	\item detectability is the same for all animals at distance r and in all directions.
+	\item no animal can be detected beyond a finite distance $w: g(r) = 0$ for $r > w$
+	\item the function $g(r)$ has at least 4 bounded derivatives on $[0, w)$, and $g'(0)=0$
+	\item the stations are chosen at random in the study area
 \end{enumerate}
 
-Ramsey 1979
-Uniform distribution, random process distribution, independent distribution,
+%Ramsey 1979: Uniform distribution, random process distribution, independent distribution,
 
-Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE
+% Not explicitly stated, is that we are assuming the objects of interest are homogeneously distributed throughout the region being surveyed. REFERENCE
 
@@
 
-Figures 2 and 3 illustrate the first two arrangements.
-Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota.
-The surveyed area of the Palie region totaled 9.41 km$^2$ during the original survey, and contained two transects of 16 and 17 stations each, 33 stations total (Ramsey et al 1982).
+Figures \ref{fig:structured} and \ref{fig:random} illustrate the first two arrangements.
+Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota from.
+The surveyed area of the Palie region totaled 9.41 km$^2$ during the original survey, and contained two transects of 16 and 17 stations each, 33 stations total \parencite{micronesian}.
 
-As seen in Figure 1, there is very little detection happening beyond 500 m. There are 15 observations beyond 500 m, which is 0.9\% of the data.
+As seen in Figure~\ref{fig:by5}, there is very little detection happening beyond 500 m. There are 15 observations beyond 500 m, which is 0.9\% of the data.
 

### 2014-05-29 report/draft3.tex
Updated draft after meeting with Claudio.

--- 2014-05-28 report/draft2.tex
@@
 \usepackage{graphicx}
-\usepackage{caption}
+\usepackage[labelfont={bf},margin=1cm]{caption}
 \usepackage{subcaption}
@@
 
+\AtEveryBibitem{ %
+	\clearfield{url} %
+	\clearfield{ISSN} %
+	\clearfield{ISBN} %
+}
+
+%\usepackage{hyperref}
+
 \begin{document}
-Ignore this page, it's so I can have a reference for the different citation formats and will be removed before turning in the paper.
-
-cite \cite{buckland2001} \cite{komodo2013} \\
-cite* \cite*{buckland2001} \cite*{komodo2013}\\
-parencite \parencite{buckland2001} \parencite{komodo2013}\\
-parencite* \parencite*{buckland2001} \parencite*{komodo2013}\\
-textcite \textcite{buckland2001} \textcite{komodo2013}\\
-smartcite - does the super script/foot note thing.
-\newpage
+%Ignore this page, it's so I can have a reference for the different citation formats and will be removed before turning in the paper.
+%
+%
+%cite \cite{buckland2001} \cite{komodo2013} \\
+%cite* \cite*{buckland2001} \cite*{komodo2013}\\
+%parencite \parencite{buckland2001} \parencite{komodo2013}\\
+%parencite* \parencite*{buckland2001} \parencite*{komodo2013}\\
+%textcite \textcite{buckland2001} \textcite{komodo2013}\\
+%smartcite - does the super script/foot note thing.
+%\newpage
 \begingroup  
   \centering
-  \LARGE Master's Report - Variable Circular Plots: Station Placement and the Independence Assumption\\[1em]
+  \LARGE Masters Report - Variable Circular Plots: Station Placement and the Independence Assumption\\[1em]
   \large Matt Edwards\\
@@
 Distance sampling, in the form of line and point transects (alternatively, Variable Circular Plots (VCP)), is a common estimation method.
-They are useful when individual objects are too small to be seen from the air or may be masked by the canopy (birds, small animals, ground plants.)
+They are useful when individual objects are too small to be seen from the air or may be masked by the canopy (birds, small animals, ground plants, etc.).
 
-Distance sampling has been around since the early 1900's.
-In the 1970's, more rigorous scientific methods were applied, including papers by Ramsey and Scott \parencite*{ramsey1979,ramsey1981} which augmented work done by Emlen in 1971.
+Distance sampling has been around since the early 1900s.
+In the 1970s, more rigorous scientific methods were applied, including papers by Ramsey and Scott \parencite*{ramsey1979,ramsey1981} which augmented work done by Emlen in 1971.
 \cite{burnham1980} is a monograph on the theory and application of line transect sampling.
@@
 Distance sampling continues to be used and modified, combined with mark-recapture methods \parencite{laake2011}, extended for use with underwater acoustics for estimates of krill populations \parencite{krill2011}, and a study by \textcite{camera2011} combines distance sampling with camera traps to estimate populations.
-
-\textcite{buckland2001} has a more in-depth discussion of the history of distance sampling.
+For a more in-depth discussion of the history of distance sampling see \textcite{buckland2001}.
 
@@
 
-\textcite[6]{buckland2001} draws a distinction between point transects, where only the area immediately surrounding the point is surveyed (area of estimated perfect detectability), and Variable Circular Plots, where you census the area $\pi w^2$, where $w$ is the furthest distance at which you can observe the object of interest.
+\textcite{buckland2001} draws a distinction between point transects, where only the area immediately surrounding the point is surveyed (area of estimated perfect detectability), and Variable Circular Plots, where a census is taken of the area $\pi w^2$, where $w$ is the furthest distance at which there is a possibility of observing the object of interest.
 However, in \textcite{buckland2006, quang1993} the terms are used interchangeably.
-They will be used interchangeably here also, with the understanding this paper addresses the second type, where an area of $pi*w^2$ is censused.
+They will be used interchangeably here also, with the understanding this paper addresses the second type, where an area of $\pi w^2$ is censused.
 
@@
 Considering the probability of detecting an object, we can imagine that there is an area of (nearly) perfect detectability near the observer, that drops off the further away from the observer the object is.
-This detectability function is labeled $g(r)$, with $r$ being the observation distance.
+This detectability function, defined as the probability of detecting an object, given that it is at distance $r$ is labeled $g(r)$, with $r$ being the observation distance.
 If no other covariates are tracked, we end up with a count of objects observed, and the distances from the line or point at which they were observed.
@@
 
-An estimate of population density is a count of objects observed, divided by the area in which they were observed, times the probability of observing the object at that distance.
+An estimate of population density is $n$, the count of objects observed, divided by the area in which they were observed, times the probability of observing the object at that distance.
+\cite{buckland2001} %eq 2.12, pg 39
 
-$\hat{D}=E[n]/(Area)*P(observing object|distance r)$
+$\hat{D}=n/(Area)*P(observing object|distance r)$
 
@@
 \end{figure}
-The Micronesian Forest Bird Survey \cite{micronesian} is a work using Variable Circular Plots (VCPs) placed using a random-systematic transect design to monitor the density of several bird species in the Micronesian islands.
+The ``Micronesian Forest Bird Survey'' by \textcite{micronesian} is a work using Variable Circular Plots (VCPs) placed using a random-systematic transect design to monitor the density of several bird species in the Micronesian islands.
 
@@
 First, there are very low counts of birds immediately around the station, a peak at around 100 m, a fairly level section from 100--200 m, and a drop off after 200 m.
-This would seem to indicate that the presence of the observers caused either movement away from the station (the crest around 100m) or alternatively a suppression of vocalizations and/or movement directly around the station.
+This would seem to indicate that the presence of the observers caused either movement away from the station (the crest around 100 m) or alternatively a suppression of vocalizations and/or movement directly around the station.
 
@@
 One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously with some uniform density, $D$, throughout the space \textcite{ramsey1981}.
-(If the data is clustered we have other problems, not addressed here.) If the uniform distribution is the case, then it would make sense that it shouldn't matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
+(If the data is clustered we have other problems, not addressed here.) If the uniform distribution is the case, then it would make sense that it should not matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
 This echoes Buckland’s statement a few paragraphs previous.
@@
 \subsection{Bias \& Consistency}
-\textcite{buckland2001,ramsey1979} Buckland et al. (2001) and Ramsey \& Scott (1970) both stated that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
+\textcite{buckland2001,ramsey1979} Buckland et al. (2001) and Ramsey \& Scott (1970) both claim that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
 
-\textcite{barry2001} Barry and Walsh (2001) compared design-based and  model-based estimates of density and variance.
+\textcite{barry2001} compared design-based and  model-based estimates of density and variance.
 (Their focus was line transects, but they state it is analogous to point transects and the conclusions should carry across.) They concluded that model-based estimates were unbiased if and only if the assumptions of objects being independent and uniformly located held.
-They noted that in practice, this is probably not a feasible assumption.
+They noted that, in practice, this is probably not a feasible assumption.
 For design-based inference, they caution about both the density and variance estimates.
 
-\textcite{roeder1987} Roeder, Dennis, \& Garton (1987) use simulations to demonstrate that no method is completely applicable in 100\% of situations, and will go sideways at some point, depending on the specific conditions.
+\textcite{roeder1987} use simulations to demonstrate that no method is completely applicable in 100\% of situations, and will go wrong at some point, depending on the specific conditions.
 
-The overall conclusion is that, in theory, density estimates using VCPs are unbiased and (reasonably) consistent, but that the assumptions necessary for the theory to hold are untenable.
+The overall conclusion is that, if all assumptions hold, density estimates using VCPs are unbiased and reasonably consistent.
+In practice, however, having the assumptions actually hold true is a precarious proposition.
 The appropriate estimation of $g(r)$ will vary depending on the specific situation, and it is up to the scientist and statistician to understand the data and choose the best approach.
@@
 \section{Simulation}
-As seen in the previous section, the issue of independence of the point transects does not have a consensus in the literature.
+As was discussed in the previous section, the issue of independence of the point transects does not have a consensus in the literature.
 If independence is required for proper estimation, then the random-systematic placement (hereafter, ``transect'') layout may require some adjustment of the estimate to bring it in line with survey layouts where the observations areas do not overlap.
 
-Through simulation, we can explore what happens under three different VCP placement schemes, detection densities, and object movement, and compare the resulting density estimates to their known values to evaluate the performance of the estimators.
+Through simulation, we can explore what happens under different VCP placement schemes, detection densities, and object movement amounts, and compare the resulting density estimates to their known values to evaluate the performance of the estimators.
 
@@
 	\centering
-	\caption{VCP Layout options. On graph, 1 unit = 1 km. Circles represent 200 m mark.}
+
 	\begin{subfigure}[b]{0.45\textwidth}
@@
 	\end{subfigure}
+	\caption{VCP Layout options. On graph, 1 unit = 1 km. Circles represent 200 m mark.}
 	\label{fig:layouts}
@@
 
-\subsection{$g(r)$ Options}
+\subsection{Detection Probability Options}
 The first simulation compares the three layouts described in section~\ref{sec:layout} with two different options for $g(r)$, a half-normal detection probability function, and a detection probability estimated from the empirical observation distances.
@@
 For a half-normal with parameter $\theta=8.7732$, $f(0)=5.5852$ so to scale the value to 1:
-
+\begin{center}
 $\delta_{HN} = 1/5.5852 = 0.1790$
+\end{center}
 
@@
 \section{Results}
+Two simulations were run, one looking at the three layout options crossed with the two detection probability function options and the other examining the effect of the three layout options against the three movement options, holding the detection probability option standard.
 \subsection{Half-normal vs. Empirical $g(r)$}
@@
 
-\texttt{TODO: Try to xplain the .50 jump in the estimate for transect/compounded. No theory or story atm.}
+The combination of compounded movement and the transect layout seems to have an effect on the density estimate.
+The other estimates for either movement type, are all in the same ballpark, around 11.8 with around a 59\% rate for capturing the true population value.
+For the Transect Layout and Compounded movement, this rises 0.50 to a mean population density estimate of 12.42, and a capture percent of 65\%.
+It is unclear if there is something about this combination that is reducing bias (possible, but feels unlikely), or if it is instead biasing the estimate upwards, which appears like bias reduction because the estimate is already low.
+
+%\texttt{TODO: Try to xplain the .50 jump in the estimate for transect/compounded. No theory or story atm.}
 
@@
 \section{The Problem with Circles}
-To test the effectiveness of my movement algorithm, I compared the detection distances from the data where the objects moved, to the detection distances from the same data, assuming they stayed still.
+To test the effectiveness of the movement algorithm, the detection distances from the data where the objects moved was compared to the detection distances from the same data, assuming the objects stayed still.
 The movement data, (left panel, Figure~\ref{fig:moveTest}) is squashed more towards 0.1 Km than the still data (right panel).
-But there's a problem.
-The still data doesn't show the nice shoulder at 0 that the literature instructs us to expect.
+But there is a problem.
+The still data does not show the nice shoulder at 0 that the literature instructs us to expect.
 
@@
 The problem is because circles are different than rectangles.
-
-With a line transect, as we move away from the line, the distance ``surveyed" increases linearly.
+With a line transect, as we move away from the line an equal distance in either direction, the distance surveyed increases linearly.
 With a point transect, as we move away from the point, the distance surveyed increases exponentially.
@@
 a VCP.
-For the VCP, Area was calculated as $2*pi*r$.
-For the line transect, Area was calculated as $(2r)*L$, with $L=0.785$ chosen so that the area of the circle and rectangle were equal when $r=0.5$.
+For the VCP, Area was calculated as $2\pi r$.
+For the line transect, Area was calculated as $(2r)L$, with $L=0.785$ chosen so that the area of the circle and rectangle were equal when $r=0.5$.
 
@@
 
-My conclusion from this is that for a VCP survey, even if everything is perfect, you will not see the nice shouldered layout of your empirical detection distances.
-\textcite{ramsey1981} speaks about this somewhat, when comparing the ``CUM-D'' curve, or the plot of the cumulative number of observations at distance $r$ for both line and point transects.
-However, almost all the illustrations on how your collected data should look, shows data that echoes the detection probability curve, with the half-bell-shape with a nice shoulder.
-Since this shape is impossible, I think some better examples are in order.
+The take-away point is that for a VCP survey, even if everything is perfect, the empirical detection distances will not show the nice shouldered layout of the ideal detection probability curve.
+\textcite{ramsey1981} speak about this, when comparing the ``CUM-D'' curve, or the plot of the cumulative number of observations at distance $r$ for both line and point transects.
+However, almost all the illustrations on how the collected data should look, shows data that echoes the detection probability curve: the half-bell-shape with a nice shoulder.
+Since this shape is impossible for VCP, explicit examples should be shown in the literature so researchers do not assume their is something wrong with their data.
 
-Better mathematicians than I have worked on distance sampling, and this is why in analyzing point transect data we take the derivative of $g(r)$, to adjust for this difference in how area is calculated.
+Many researchers have worked on distance sampling, and this effect is taken into account in the calculation of population density estimates.
+When analyzing point transect data we take the derivative of $g(r)$, an adjustment which is not necessary in line transect surveys.
+It can be shown that this solution adjusts for the difference in how area is calculated, to result in correct density estimates.
 
@@
 	\centering
-	\caption{Rectangles (Line Transect, solid) vs. Circles (Point Transect, dashed)}
 	\begin{subfigure}[b]{0.45\textwidth}
@@
 	\end{subfigure}
+	\caption{Rectangles (Line Transect, solid) vs. Circles (Point Transect, dashed)}
+	\label{fig:linevs}
 \end{figure}
@@
 \section{Conclusions}
-If all else is held equal, VCP layout does not seem to play a large role in density estimate using the kernel method as described by \textcite{quang1993}.
+If all else is held equal, VCP layout does not seem to play a large role in population density estimates using the kernel method as described by \textcite{quang1993}.
 Violations of the expected detection probability curve, or by movement of the objects away from the observer play more of a role in biasing the resulting estimate.
 
-\section{Bibliography}
+All of the population density estimates were biased low using Quang's formula for $\hat{D}$.
+This bias was observed even when the detection function was a half-normal curve and analyzed with a normal kernel, and the VCPs were placed so that the observation area did not overlap.
+The bias was not severe in this case, it estimated a population density around 18 objects per km$^2$, when the true estimate was 20/km$^2$, but it is noticeable.
+Dr.
+Quang's paper does offer a bias-corrected estimate, which was not used in this simulation.
+
+%\section{Bibliography}
 

### 2014-05-31 code/SimulationDriver.R
Final simulation set up

@@ -7,7 +7,7 @@ library(fdrtool) # for halfnormal which you need for project_functions
 # ============ set up environment ============ 
 D.known <- 20
 max.r <- 0.5
-nsim <- 100
+nsim <- 1000
 
 # === half normal parameters for g(r) ===
 hn.params <- VCP.defineHalfnorm(max.r)
@@ -31,7 +31,8 @@ xy.structured <- VCP.structuredLayout()
 D.hat.df <- data.frame()
 
 # http://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/
-pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+#pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+# didn't work on Mac for somereason >_>
 
 ## IN loop -----------------------------------------------------------------
 for(i in 1:nsim){
@@ -46,24 +47,45 @@ for(i in 1:nsim){
   #   for each detection function
   #     get the density estimate.
   
-  D.hnorm.S <- VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm")
-  D.hnorm.R <- VCP.dHat(x.y, xy.random, hn.params, w=max.r, g.type="hnorm")
-  D.hnorm.T <- VCP.dHat(x.y, xy.transect, hn.params, w=max.r, g.type="hnorm")
+  ## Half normal g(r)------------------
+  #D.hnorm.S <- VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm")
+  #D.hnorm.R <- VCP.dHat(x.y, xy.random, hn.params, w=max.r, g.type="hnorm")
+  #D.hnorm.T <- VCP.dHat(x.y, xy.transect, hn.params, w=max.r, g.type="hnorm")
   
-  #D.pass <- cbind(D.hnorm.S, D.hnorm.R, D.hnorm.T) # group variables
+  D.S.hnorm <- c(VCP.dHat(x.y, xy.structured, hn.params, w=max.r, g.type="hnorm",
+                           true.D=D.known),
+                  Layout="structured", GType="HalfNorm")
+  D.R.hnorm <- c(VCP.dHat(x.y, xy.random, hn.params, w=max.r, g.type="hnorm",
+                           true.D=D.known),
+                  Layout="random", GType="HalfNorm")
[…]

### 2014-05-31 code/functions/DensityFunctions.R

@@ -27,6 +27,13 @@ VCP.Dhat.hnormKernel <- function(m, R.i, true.D=NULL) {
 }
 
 
+#xy.objects <- get.Coords(20)
+#xy.vcp <- VCP.randomLayout()
+#params <- hn.params
+#w=0.5
+#g.type="hnorm"
+#transects=F
+#true.D = 20
 
 ## g.type can be "hnorm" or "emp"; it defaults to "hnorm" if an unrecognized value entered
 
@@ -52,15 +59,15 @@ VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transect
     
     # calculate straight-line distance from observer to object, R.j
     
-    candidate.xy <- mutate(candidate.xy, x.dist = x-curr.x, y.dist=y-curr.y, R.j = sqrt(x.dist^2+y.dist^2))
+    candidate.xy <- mutate(candidate.xy, x.dist = x-curr.x, y.dist=y-curr.y, Rj = sqrt(x.dist^2+y.dist^2))
     
     # feed R.j into detection function
     detected=NULL
     if(tolower(g.type)=="emp"){
-      detected <- VCP.loessProb(candidate.xy$R.j, params)
+      detected <- VCP.loessProb(candidate.xy$Rj, params)
       #print("emp")
     } else {
-      detected <- VCP.hnormProb(candidate.xy$R.j, params)
+      detected <- VCP.hnormProb(candidate.xy$Rj, params)
       #print("hnorm")
     }
     
@@ -68,10 +75,12 @@ VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transect
     
     # store list of detected R.j in data frame, with station # and transect # if applicable
     m[i] <- sum(detected)
-    R.j <- rbind(R.j, candidate.xy %.% filter(detected==1) %.% select(x, y, R.j) %.% mutate(s_id = i, t_id=xy.vcp$t_id[i]))
+    if(sum(detected) > 0){
+      R.j <- rbind(R.j, candidate.xy %.% filter(detected==1) %.% select(x, y, Rj) %.% mutate(s_id = i, t_id=xy.vcp$t_id[i]))
+    }
   }
   
-  dhat <- VCP.Dhat.hnormKernel(m, R.j$R.j, true.D)
+  dhat <- VCP.Dhat.hnormKernel(m, R.j$Rj, true.D)
   return(dhat)
 }
 

### 2014-05-31 report/draft3.tex
Last graph and table modifications.
Some extra text discussing the new tables.
corrected old tables with new sim values.

@@
 
-	\includegraphics[width=\textwidth]{../images/Emp_Vs_Hnorm.pdf}
+	\includegraphics[width=\textwidth]{../images/Emp_Vs_Hnorm_1-2.pdf}
 	\caption{1000 Simulations, Empirical vs Half-normal Detection function, for three different layouts. Dashed line equals true value of 20 objects per km$^2$\label{fig:sim1}}
@@
 
-Table~\ref{table:sim1} supports what we observe in the histogram: the means vary between our two $g(r)$ options, but are approximately the same within the each option, regardless of which VCP layout we have chosen.
+Table~\ref{table:sim1} supports what we observe in the histogram: the centers vary between our two $g(r)$ options, but are approximately the same within the each option, regardless of which VCP layout we have chosen.
+Table~\ref{table:sim1capture} shows the percentage of simulations that captured the true object density of 20 per km$^2$ within a 95\% confidence interval.
+\cite{quang1993}
+
+There may be evidence of a drop in accuracy with the Transect layout, but it is difficult to determine if this is a valid observation, or an artifact of this particular simulation.
+The drop, if it exists, is much less severe with the half-normal detection function.
+With the empirical detection function, the structured layout percentage swings much higher than the result from our random layout, unlike for half-normal where they are just over 1\% apart.
+Repeated simulations, could help narrow down if this is a true effect, or simply random noise.
 \begin{table}[h]
-	\caption{Empirical vs. Half-normal detection function, across all three layouts. 1000 simulations. True density 20/km$^2$}
-	\begin{tabular}{ r r r r r r r r r}
-	   $g(x)$	 & Layout		& Mean	& Std Dev	& LB 95 & UB 95	& 25th	& Median	&  75th \\ \hline \hline
-	   Empirical & Random		& 14.52 & 3.35 		& 7.97	& 21.08	& 12.17	& 14.46	& 16.74\\
-	   Empirical & Structured	& 14.50 & 3.22 		& 8.19 	& 20.81 & 12.38 & 14.45 & 16.62\\
-	   Empirical & Transect		& 14.46 & 3.53 		& 7.55 	& 21.37 & 11.85 & 14.36 & 16.88\\ \hline
-	 Half-normal & Random 		& 18.17 & 4.50 		& 9.34 	& 27.00 & 15.24 & 17.94 & 21.00\\
-	 Half-normal & Structured	& 17.94 & 4.29 		& 9.54 	& 26.35 & 15.19 & 17.67 & 20.78\\
-	 Half-normal & Transect		& 18.04 & 4.49 		& 9.23 	& 26.84 & 14.90 & 17.75 & 20.74\\
 
+	\begin{tabular}{ r r |r r| r r r}
+		$g(x)$      & Layout     & Mean  & Std Dev & 25th  & Median & 75th  \\ \hline\hline
+		Empirical   & Random     & 14.47 & 3.37    & 12.08 & 14.38  & 16.73 \\
+		Empirical   & Structured & 14.49 & 3.15    & 12.27 & 14.47  & 16.59 \\
+		Empirical   & Transect   & 14.26 & 3.66    & 11.65 & 14.14  & 16.80 \\ \hline
+		Half-normal & Random     & 17.90 & 4.57    & 14.67 & 17.86  & 20.88 \\
+		Half-normal & Structured & 18.03 & 4.16    & 15.05 & 17.91  & 20.82 \\
+		Half-normal & Transect   & 17.80 & 4.64    & 14.52 & 17.44  & 20.78
 	\end{tabular}
+	\caption{Centers and Spread for Empirical vs. Half-normal detection functions. 1000 simulations. True density 20/km$^2$}
 	\label{table:sim1}
 \end{table}
+\begin{table}[h]
 
-\texttt{TO DO: return T/F if the true value was contained in the interval. redo graph to match others.}
+	\begin{tabular}{ r| r r| r r|}
+		           & \multicolumn{2}{|c|}{Empirical} & \multicolumn{2}{|c|}{Half Normal} \\ \hline\hline
+		Layout     & Mean $\hat{D}$ & \% Capture     & Mean $\hat{D}$ & \% Capture       \\ \hline\hline
+		Random     & 14.47          & 73.6\%         & 17.90          & 94.0\%           \\
+		Structured & 14.49          & 77.5\%         & 18.03          & 95.3\%           \\
+		Transect   & 14.27          & 69.6\%         & 17.80          & 91.2\%           \\ \hline
+	\end{tabular}
+	\caption{True Density Capture, Emperical vs. Half-normal detection function for three layouts. 1000 simulations.}
+	\label{table:sim1capture}
+\end{table}
 
@@
 \end{figure}
+\begin{table}[h]
+
+	\begin{tabular}{ r r| r r| r r r}
+		Movement  & Layout     & Mean  & Std Dev & 25th  & Median & 75th  \\ \hline\hline
+		None      & Random     & 17.86 & 4.28    & 15.03 & 17.64  & 20.65 \\
+		None      & Structured & 18.01 & 4.51    & 14.95 & 17.62  & 20.94 \\
+		None      & Transect   & 17.90 & 4.78    & 14.74 & 17.67  & 20.95 \\ \hline
+		Permanent & Random     & 11.86 & 3.42    & 9.58  & 11.61  & 14.09 \\
+		Permanent & Structured & 11.90 & 3.65    & 9.37  & 11.67  & 14.08 \\
+		Permanent & Transect   & 12.42 & 3.72    & 9.93  & 12.17  & 14.79 \\ \hline
+		Temporary & Random     & 11.76 & 3.57    & 9.22  & 11.75  & 14.02 \\
+		Temporary & Structured & 11.74 & 3.48    & 9.37  & 11.56  & 13.85 \\
+		Temporary & Transect   & 11.75 & 3.57    & 9.24  & 11.72  & 14.20
+	\end{tabular}
+	\caption{Centers and Spread for Movement and Layout combinations. 1000 simulations. True density 20/km$^2$}
+	\label{table:sim2}
+\end{table}
 \begin{table}
-	\caption{Half-normal detection function, three layouts, three movement types. 1000 simulations.}
+
 	\begin{tabular}{ r| r r| r r| r r|}
@@
 	\end{tabular}
-	\label{table:sim2}
+	\caption{Mean and Capture Rate, Half-normal detection function, three layouts, three movement types. 1000 simulations.}
+	\label{table:sim2capture}
 \end{table}
 
-Figure~\ref{fig:sim2} and Table~\ref{table:sim2} illustrate the results of our movement type simulation. For this run, only the half-normal detection function was used.
+Figure~\ref{fig:sim2} and Tables~\ref{table:sim2} and~\ref{table:sim2capture} illustrate the results of our movement type simulation. For this run, only the half-normal detection function was used.
 
+For testing combinations of movement and VCP layout, only the half-normal detection function was utilized.
+Figure~\ref{fig:sim} compares the histograms of the nine combinations.
+On the whole, the estimates are biased low, as was observed in the previous simulation, without movement.
 Within a column, the results are fairly similar in shape and spread.
-We see the most obvious differences here between columns, indicating that the type of movement causes more bias in our estimate than the potential overlap of our observation areas.
+The most obvious differences lie between columns, indicating that the type of movement causes more bias in our estimate than the potential overlap of our observation areas.
+
+Table~\ref{table:sim2} reinforces this conclusion, as the measures of center and spread are reasonably consistent within Movement type, regardless of the VCP Layout.
+Differences are observed between movement types, with No Movement having the estimates with the smallest amount of downward bias, and Compounded and Temporary movement biasing the population density estimate about the same amount.
 
 Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
-Table~\ref{table:sim2} shows that the 95\% confidence intervals \parencite{quang1993} capture the true mean about 59\% of the time, no matter which movement type, Temporary or Compounded.
+Table~\ref{table:sim2capture} shows that the 95\% confidence intervals \parencite{quang1993} capture the true mean about 59\% of the time, no matter which movement type, Temporary or Compounded.
 
@@
 The combination of compounded movement and the transect layout seems to have an effect on the density estimate.
-The other estimates for either movement type, are all in the same ballpark, around 11.8 with around a 59\% rate for capturing the true population value.
+The other estimates for either movement type, are all in the same ballpark, around 11.8 with an approximate 59\% rate for capturing the true population value.
 For the Transect Layout and Compounded movement, this rises 0.50 to a mean population density estimate of 12.42, and a capture percent of 65\%.
@@
 
+There is a possible effect of the combination of compounded movement and the transect layout.
+It is unclear from these results what the particular cause of the increased estimate and accuracy is.
+Future work might involve modification of the simulation functions to return more information, such as the count of observed animals and the observation distances to assist with understanding why this combination is producing an estimate that appears out of line with it's neighbors.
+
 %\section{Bibliography}

### 2014-05-31 code/SimulationAnalysis_05-18.R

@@ -1,5 +1,12 @@
 library(ggplot2)
+library(dplyr)
 
+OSUOrange <- rgb(243/255,115/255,33/255) #http://oregonstate.edu/brand/color-palette
+OSURed <- rgb(192/255,49/225,26/255)
+OSUBlue1 <- rgb(93/255, 135/255, 161/255)
+OSUBlue2 <- rgb(156/255, 197/255, 202/255)
+
+#### Original Sim 1 ------------------------------------------------------------------------
 raw.data <- read.csv("data/Sim1_05-18.csv", stringsAsFactors=F, header=T)
 
 d.hats <- c(raw.data$D.hnorm.S, raw.data$D.hnorm.R, raw.data$D.hnorm.T,
@@ -32,3 +39,60 @@ dhat.df %.% group_by(g.X, layout) %.%
             )
 
 
+#####  Mac Sim 1.2 ----------------------------------------------------------------------------
+mac.sim <- tbl_df(read.csv("data/Sim1_05-31_MAC.csv", stringsAsFactors=F, header=T))
+
+names(mac.sim) <- c("D.hat", "ContainsTrue", "Layout", "GType")
+
+
+ggplot(mac.sim, aes(x=D.hat))+geom_histogram(aes(y=..density..), binwidth=1, colour="white", fill=OSUOrange)+facet_grid(Layout~GType)+theme_bw(18)+geom_vline(xintercept=20, colour=OSURed, linetype="longdash", size=1)+xlab("Density Estimates")+ylab("Frequency")+ggtitle("Mac Sim")
+
+ggsave("images/Emp_Vs_Hnorm.pdf", width=6, height=6)
+
+#hmm. not really the drastic differences I was expecting.
+#after tweaking movement, it looks better
+
+move.summ <- D.hat.df %.% group_by(Layout, MoveType) %.% summarise(avg=mean(Dhat), n.In=sum(inCI))
+
+
+#####  PC Sim 1.2 ----------------------------------------------------------------------------
+PC.sim <- tbl_df(read.csv("data/Sim1_05-31_PC.csv", stringsAsFactors=F, header=T))
+
+names(PC.sim) <- c("D.hat", "ContainsTrue", "Layout", "GType")
+
+
[…]

### 2014-05-31 code/SimulationDriver.R

@@ -31,7 +31,7 @@ xy.structured <- VCP.structuredLayout()
 D.hat.df <- data.frame()
 
 # http://ryouready.wordpress.com/2009/03/16/r-monitor-function-progress-with-a-progress-bar/
-#pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
+pb <- winProgressBar(title="progress bar", min=0, max=nsim, width=300)
 # didn't work on Mac for somereason >_>
 
 ## IN loop -----------------------------------------------------------------
@@ -67,13 +67,13 @@ for(i in 1:nsim){
   #D.emp.T <- VCP.dHat(x.y, xy.transect, emp.params, w=max.r, g.type="emp")
   D.S.emp <- c(VCP.dHat(x.y, xy.structured, emp.params, w=max.r, g.type="emp",
                           true.D=D.known),
-                 Layout="structured", GType="Emperical")
+                 Layout="structured", GType="Empirical")
   D.R.emp <- c(VCP.dHat(x.y, xy.random, emp.params, w=max.r, g.type="emp",
                           true.D=D.known),
-                 Layout="random", GType="Emperical")
+                 Layout="random", GType="Empirical")
   D.T.emp <- c(VCP.dHat(x.y, xy.transect, emp.params, w=max.r, g.type="emp",
                           true.D=D.known),
-                 Layout="transect", GType="Emperical")
+                 Layout="transect", GType="Empirical")
   
   ## Add to data frame --------------
   #D.pass <- cbind(D.hnorm.S, D.hnorm.R, D.hnorm.T, D.emp.S, D.emp.R, D.emp.T)
@@ -87,5 +87,5 @@ for(i in 1:nsim){
 }
 ## END Loop -----------------------------------------------------------------
 
-write.csv(D.hat.df, "data/Sim1_05-31.csv", row.names=FALSE)
+write.csv(D.hat.df, "data/Sim1_05-31_PC.csv", row.names=FALSE)
 #   already done, don't want to overwrite original simulation data.

### 2014-05-31 code/functions/DensityFunctions.R

@@ -78,6 +78,7 @@ VCP.dHat <- function(xy.objects, xy.vcp, params, w=0.5, g.type="hnorm", transect
     if(sum(detected) > 0){
       R.j <- rbind(R.j, candidate.xy %.% filter(detected==1) %.% select(x, y, Rj) %.% mutate(s_id = i, t_id=xy.vcp$t_id[i]))
     }
+      ## had to add if statement on 5/31 because when running it on Mac, was getting instances were no objects were detected, even if several were in the .1-.2 range, and it was throwing errors.  PC had higher max observation
   }
   
   dhat <- VCP.Dhat.hnormKernel(m, R.j$Rj, true.D)

### 2014-06-03 report/draft3.tex
Modified equations

@@
 \usepackage{textcomp}
+\usepackage{amsmath} %larger fractions
 %\usepackage[round,authoryear]{natbib}
@@
   \large Matt Edwards\\
-  26 May 2014\par
+ % 26 May 2014\par
+  \today\par
 \endgroup
@@
 
-$\hat{D}=n/(Area)*P(observing object|distance r)$
+$\hat{D}=n/(\mbox{Area})*P(\mbox{observing object}|\mbox{distance }r)$
 
@@
 
-$\theta = \frac{\sqrt{\pi /2}}{\sigma}$
+$\theta = \dfrac{\sqrt{\pi /2}}{\sigma}$
 
@@
 
-$\sigma = w/3.5 = 0.1429$
+$\sigma = \dfrac{w}{3.5} = 0.1429$
 
-$\theta = \frac{\sqrt{\pi /2}}{0.1429}=8.7732$
+$\theta = \dfrac{\sqrt{\pi /2}}{0.1429}=8.7732$
 
@@
 \begin{center}
-$\delta_{HN} = 1/5.5852 = 0.1790$
+$\delta_{HN} = \dfrac{1}{5.5852} = 0.1790$
 \end{center}
@@
 Half-normal density functions were executed in R with the \texttt{fdrtool} package.
-
-\subsubsection{Empirical Observations}
 \begin{figure}
@@
 \end{figure}
-
-We can see in Figure \ref{fig:by5} that above 100 m, observations were commonly heaped into 10 or 25 m distances, with the latter being more heavily used. Above 200 m, 50 m heaps were common.
+\subsubsection{Empirical Observations}
+Figure \ref{fig:by5} illustrates that above 100 m, observations were commonly heaped into 10 or 25 m distances, with the latter being more heavily used. Above 200 m, 50 m heaps were common.
 
@@
 
-$\delta_{EMP}=\frac{1}{max(predicted)}=40.5780$
+$\delta_{EMP}=\dfrac{1}{max(predicted)}=40.5780$
 
@@
 
-$E[observed] = Area*Density*P(object observed | distance r)$
+$E[observed] = Area*Density*P(\mbox{object observed}|\mbox{distance }r)$
 

### 2014-06-03 report/draft3.tex
Pre-final readthrough

@@
 Distance sampling, in the form of line and point transects (alternatively, Variable Circular Plots (VCP)), is a common estimation method.
-They are useful when individual objects are too small to be seen from the air or may be masked by the canopy (birds, small animals, ground plants, etc.).
+VCPs are useful when individual objects are too small to be seen from the air or may be masked by the canopy (birds, small animals, ground plants, etc.).
 
-Distance sampling has been around since the early 1900s.
+Distance sampling has been around since road-side surveys in the early 1900s.
 In the 1970s, more rigorous scientific methods were applied, including papers by Ramsey and Scott \parencite*{ramsey1979,ramsey1981} which augmented work done by Emlen in 1971.
-\cite{burnham1980} is a monograph on the theory and application of line transect sampling.
+\textcite{burnham1980} wrote a monograph on the theory and application of line transect sampling.
 Non-parametric methods have been examined by \textcite{quang1993,mack1998}, among others.

### 2014-06-04 report/draft3.tex
Final Draft of Master's Project Report.

@@
 This results in data similar to the line transect, but with more safety for the observer as they do not have to watch for moving wildlife at the same time as navigating potentially treacherous terrain.
-\cite{ramsey1979}.
+\parencite{ramsey1979}.
 
 With line transects, animals may react to an observer moving through the area by hiding or moving away.
-If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal--possible with point transects--animal behavior may be more natural and less influenced by the presence of the observer.
+If the observer is standing still, and has allowed a ``cooling'' period for things to return to normal---possible with point transects---animal behavior may be more natural and less influenced by the presence of the observer.
 
@@
 
-An estimate of population density is $n$, the count of objects observed, divided by the area in which they were observed, times the probability of observing the object at that distance.
+An estimate of population density is $n$, the count of objects observed, divided by the area in which they were observed times the probability of observing the object at that distance.
 \cite{buckland2001} %eq 2.12, pg 39
 
-$\hat{D}=n/(\mbox{Area})*P(\mbox{observing object}|\mbox{distance }r)$
+$\hat{D}=\dfrac{n}{\mbox{Area}*P(\mbox{observing object}|\mbox{distance }r)}$
 
@@
 This data represents the aggregated data from two islands, Tinian and Rota, and all four observers.
+Most of the observations were made by sound, with a smaller percent visual, and a smaller percent heard-then-seen.
 We can observe several things in this plot:
@@
 
-Secondly, the further away from the station, the distance observations were heaped into convenient distances. With smaller bin widths, this is also evident in the 100--200 m range as well.
+Secondly, the further away from the station, the distance observations were heaped into convenient distances.
+With smaller bin widths, this is also evident in the 100--200 m range as well.
+See Figure~\ref{fig:by5}.
 
@@
 How does this affect the independence of our observations?
-Will it make a difference in our final population density estimates?
+Will it make a difference in our final population density estimates if a bird is observed from more than one station?
 
@@
 This would seem to implicitly state that their observational ranges might by chance overlap, meaning a single object of interest could be observed from more than one station.
-\textcite[240]{thompson2012} likewise indicates random placement of line transects which would seem to allow for a potential overlap of observation zones.
+\textcite{thompson2012} likewise indicates random placement of line transects which would seem to allow for a potential overlap of observation zones.
 
@@
 
-\textcite[233]{buckland2001} discuss that each line or point must be ``randomly and independently located,'' with ``an equal probability of selection'' for all portions of the study area.
+\textcite{buckland2001} discuss that each line or point must be ``randomly and independently located,'' with ``an equal probability of selection'' for all portions of the study area.
 
-\textcite[235]{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet unsampled lines.''
+\textcite{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet unsampled lines.''
 
@@
 
-In discussing line transects, \textcite{barry2001} treats it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
+In discussing line transects, \textcite{barry2001} treat it as a two-stage sample, with non-unique clusters (each study object can be in more than one cluster) having an equal probability of selection.
 This would seem to (but does not explicitly) endorse the possibility of overlapping VCPs.
 
-One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously with some uniform density, $D$, throughout the space \textcite{ramsey1981}.
-(If the data is clustered we have other problems, not addressed here.) If the uniform distribution is the case, then it would make sense that it should not matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
+One of the assumptions of distance sampling is that the objects being studied are distributed homogeneously with some uniform density, $D$, throughout the space.
+\parencite{ramsey1981} (If the data are clustered we have other problems, not addressed here.) If the uniform distribution is the case, then it would make sense that it should not matter if the observations overlap, as long as the act of observing at one station does not affect the observations at the neighboring station.
 This echoes Buckland’s statement a few paragraphs previous.
 
-\textcite[244]{thompson2012} mentions that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
+\textcite{thompson2012} mentions that the systematic selection of transects (random position of first, with additional transects a fixed distance apart) does not affect the approximate unbiasedness of the density estimator, but will have an effect on the approximate unbiasedness of the variance estimator.
 
-It is clear the literature is divided on whether or not our observation areas should overlap. VCP layouts similar to \textcite{micronesian} ``random-systematic'' layout are common in the literature.
+It is clear the literature is divided on whether or not our observation areas should overlap.
+VCP layouts similar to \textcite{micronesian} ``random-systematic'' layout are common in the literature.
+We as scientists need to be aware of any bias this may introduce into our estimates.
 
 \subsection{Bias \& Consistency}
-\textcite{buckland2001,ramsey1979} Buckland et al. (2001) and Ramsey \& Scott (1970) both claim that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
+\textcite{buckland2001,ramsey1979} both claim that if our detectability estimate $g(z)=$1 out to $r$, then $\hat{D}$ is unbiased.
 
@@
 \end{figure}
-Figures~\ref{fig:structured} and \ref{fig:random} illustrate the first two arrangements. Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota.
+Figures~\ref{fig:structured} and \ref{fig:random} illustrate the first two arrangements.
+Density in these images is based on the 20 birds per km$^2$ of the Palie region on Rota.
+\parencite{micronesian}
 
-As seen in Figure~\ref{fig:82dist}, there is very little detection happening beyond 500 m. There are a total of 15 observations beyond 500 m, which is 0.9\% of the data.
+As seen in Figure~\ref{fig:82dist} or Figure~\ref{fig:by5}, there is very little detection happening beyond 500 m.
+There are a total of 15 observations beyond 500 m, which is 0.9\% of the data.
+For ease of simulation, 500 m was designated as our truncation distance, $w$, the distance beyond which the probability of detection is effectively 0.
+
+The Palie region contained two transects of 16 and 17 stations each, 33 stations total \parencite{micronesian}.
+For the transect layout, an angle was chosen between $0$ and $\pi$.
+A transect of 18 stations 150 m apart was constructed, and a second transect of 18 stations constructed running parallel.
+The transects were then randomly placed with in the graph, no closer to the edge than 500 km (0.5 on graph.) This transect layout is illustrated in Figures~\ref{fig:transect1} and \ref{fig:transect2}.
+The total of 36 VCPs was carried over to the Structured and Random layouts, for an equivalent number of points.
 
@@
 
-The Palie region contained two transects of 16 and 17 stations each, 33 stations total \parencite{micronesian}.
-For the transect layout, an angle was chosen between $0$ and $\pi$.
-A transect of 18 stations 150 m apart was constructed, and a second transect of 18 stations constructed running parallel.
-The transects were then randomly placed with in the graph, no closer to the edge than 500 km (0.5 on graph.) This transect layout is illustrated in Figures~\ref{fig:transect1} and \ref{fig:transect2}.
-The total of 36 VCPs was carried over to the Structured and Random layouts, for an equivalent number of points.
+For the simulations, 0.1 unit = 1 km. All figures except Figure~\ref{fig:82dist} and \ref{fig:by5} follow this convention.
 
@@
 Using the bin midpoints, the density estimate was multiplied by 5 to get the density value for that bin.
-A LOESS regression was run with the midpoint as $x$ and $y=density*5$, using the \texttt{loess()} package in R with a span of 0.2.
+A LOESS regression was run with the midpoint as $x$ and $y=density*5$, using the \texttt{loess} package in R with a span of 0.2.
 This is the solid line in Figure~\ref{fig:by5}.
 
-This gives us an object that can be used with \texttt{predict()} to get density estimates for specific distances.
+This gives us an object that can be used with the \texttt{predict} function to get density estimates for specific distances.
 The values generated by feeding our midpoints into our prediction function are represented by the dashed line in Figure~\ref{fig:by5}.
 
-Following the scaling of the highest point to 1, as was done with the half normal curve, where the highest point was $g(0)$, we also find $\delta$ for this empirical detection probability.
-A sequence of $x$ values was generated covering the range from 0--500 m.
-This was plugged into the \texttt{predict()} function using our LOESS object.
+This empirical detection function needs to be scaled as we did with the half normal detection function.
+A sequence of values was generated covering the range from 0--500 m.
+These were plugged into the \texttt{predict()} function using our LOESS object.
 The highest density value was then scaled to be 1:
@@
 \subsubsection{Movement}
-In the Figure~\ref{fig:82dist}, the histogram of empirical observation distances, there appears to be evidence of movement away from the observer.
+In Figure~\ref{fig:82dist}, the histogram of empirical observation distances, there appears to be evidence of movement away from the observer.
 To see if the movement would have an effect on the independence of our VCPs, depending on placement, a second simulation was run.
@@
 \item No Movement: Objects stayed at original point.
-\item Temporary Movement: Objects within 100 m had a random chance to move directly away from observer.
-Movement was larger if they started closer to observer.
-They were reset to original location before moving to next VCP
-\item Permanent Movement: As with temporary movement, only not reset before moving on to next VCP. (Movement could compound if object was within the maximum observation radius for more than one VCP.)
+\item Temporary Movement: Objects within 100 m of the observer had a random chance to move directly away from the observer.
+Movement was larger if the object started closer to observer.
+Objects were reset to original location before moving to next VCP
+\item Permanent Movement: As with temporary movement, only object location was not reset before moving on to next VCP.
+(Movement could compound if object was within the maximum observation radius for more than one VCP.)
 \end{enumerate}
@@
 Two simulations were run, one looking at the three layout options crossed with the two detection probability function options and the other examining the effect of the three layout options against the three movement options, holding the detection probability option standard.
-\subsection{Half-normal vs. Empirical $g(r)$}
+\subsection{Half-normal vs. Empirical Detection Probability}
 \begin{figure}
@@
 Figure \ref{fig:sim1} shows the results of 1000 simulations, each analyzed with Empirical and Half-normal detection function, $g(x)$, and using each of the three layouts, Structured, Random, and Transect.
-All $\hat{D}$  were done using the kernel method with a normal kernel \parencite{quang1993}.
+All $\hat{D}$ estimations were done using the kernel method with a normal kernel \parencite{quang1993}.
 
@@
 
-As we would expect, the density estimates using $g(x)$ based on the Empirical distance observations is biased low.
-We also see bias in the estimates using the half-normal function for $g(x)$, but it is much less severe than in the Empirical data.
+As we would expect, the density estimates using $g(x)$ based on the Empirical distance observations are biased low.
+We also see bias in the estimates using the half-normal function for $g(x)$, but it is much less severe than those based on the Empirical data.
 \textcite{quang1993} does provide a bias-correction method, but it was not used for this simulation.
@@
 Table~\ref{table:sim1capture} shows the percentage of simulations that captured the true object density of 20 per km$^2$ within a 95\% confidence interval.
-\cite{quang1993}
+\parencite{quang1993} Again, it is fairly consistent within detection probability type, regardless of layout choice.
 
-There may be evidence of a drop in accuracy with the Transect layout, but it is difficult to determine if this is a valid observation, or an artifact of this particular simulation.
+There may be evidence of a drop in accuracy with the Transect layout.
+Both the mean and median is lower for the transect layout (Table~\ref{table:sim1} and \ref{table:sim1capture}), but it is difficult to determine if this is a valid observation, or an artifact of this particular simulation.
 The drop, if it exists, is much less severe with the half-normal detection function.
+
 With the empirical detection function, the structured layout percentage swings much higher than the result from our random layout, unlike for half-normal where they are just over 1\% apart.
@@
 \end{figure}
+
+Figure~\ref{fig:sim2} and Tables~\ref{table:sim2} and~\ref{table:sim2capture} illustrate the results of our movement type simulation. For this run, only the half-normal detection function was used.
+
+For testing combinations of movement and VCP layout, only the half-normal detection function was utilized.
+Figure~\ref{fig:sim2} compares the histograms of the nine combinations.
+On the whole, the estimates are biased low, as was observed in the previous simulation, without movement.
+Within a column, the results are fairly similar in shape and spread.
+The most obvious differences lie between columns, indicating that the type of movement causes more bias in our estimate than the potential overlap of our observation areas.
+
+Table~\ref{table:sim2} reinforces this conclusion, as the measures of center and spread are reasonably consistent within Movement type, regardless of the VCP Layout.
+Differences are observed between movement types, with No Movement having the estimates with the smallest amount of downward bias, and Compounded and Temporary movement biasing the population density estimate about the same amount.
+
+Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
+Table~\ref{table:sim2capture} shows that the 95\% confidence intervals \parencite{quang1993} capture the true mean about 59\% of the time, no matter which movement type, Temporary or Compounded.
+
+The similarity for the Structured and Random layouts can be explained by a general lack of compounded movement.
+For the Structured layout, the animals would only move once, additional movement would not be triggered by an overlapping VCP, so the Compounded Movement setting would have the same end result as the Temporary Movement setting.
+For the Random VCP layout we would expect there to be a larger effect from the Compounded movement, because we will have VCPs whose observation distances overlap.
+
+Looking at Figure~\ref{fig:random}, which is just one possible random layout, there are 8 pairs of VCP where the circles representing the 200 m mark overlap, and 3 where they overlap to any significant degree.
+With most of the movement happening within that first 100 m, we would expect the estimate to be fairly similar to the Structured layout setting, if this random layout is typical of those generated.
+
+The combination of compounded movement and the transect layout seems to have an effect on the density estimate.
+The other estimates for either movement type, are all in the same ballpark, around 11.8 with an approximate 59\% rate for capturing the true population value.
+For the Transect Layout and Compounded movement, this increases by 0.50 to a mean population density estimate of 12.42, and a capture percent of 65\%.
+It is unclear if there is something about this combination that is reducing bias (possible, but feels unlikely), or if it is instead biasing the estimate upwards, which appears like bias reduction because the estimate is already low.
+
 \begin{table}[h]
@@
 
-Figure~\ref{fig:sim2} and Tables~\ref{table:sim2} and~\ref{table:sim2capture} illustrate the results of our movement type simulation. For this run, only the half-normal detection function was used.
-
-For testing combinations of movement and VCP layout, only the half-normal detection function was utilized.
-Figure~\ref{fig:sim} compares the histograms of the nine combinations.
-On the whole, the estimates are biased low, as was observed in the previous simulation, without movement.
-Within a column, the results are fairly similar in shape and spread.
-The most obvious differences lie between columns, indicating that the type of movement causes more bias in our estimate than the potential overlap of our observation areas.
-
-Table~\ref{table:sim2} reinforces this conclusion, as the measures of center and spread are reasonably consistent within Movement type, regardless of the VCP Layout.
-Differences are observed between movement types, with No Movement having the estimates with the smallest amount of downward bias, and Compounded and Temporary movement biasing the population density estimate about the same amount.
-
-Unexpectedly, the type of movement, temporary or permanent, does not seem to play a large part in changing the bias.
-Table~\ref{table:sim2capture} shows that the 95\% confidence intervals \parencite{quang1993} capture the true mean about 59\% of the time, no matter which movement type, Temporary or Compounded.
-
-The similarity for the Structured and Random layouts can be explained by a general lack of compounded movement.
-For the Structured layout, the animals would only move once, additional movement would not be triggered by an overlapping VCP, so the Compounded Movement setting would have the same end result as the Temporary Movement setting.
-For the Random VCP layout we would expect there to be a larger effect from the Compounded movement, because we will have VCPs whose observation distances overlap.
-
-Looking at Figure~\ref{fig:random}, which is just one possible random layout, there are 8 pairs of VCP where the circles representing the 200 m mark overlap, and 3 where they overlap to any significant degree.
-With most of the movement happening within that first 100 m, we would expect the estimate to be fairly similar to the Structured layout setting, if this random layout is typical of those generated.
-
-The combination of compounded movement and the transect layout seems to have an effect on the density estimate.
-The other estimates for either movement type, are all in the same ballpark, around 11.8 with an approximate 59\% rate for capturing the true population value.
-For the Transect Layout and Compounded movement, this rises 0.50 to a mean population density estimate of 12.42, and a capture percent of 65\%.
-It is unclear if there is something about this combination that is reducing bias (possible, but feels unlikely), or if it is instead biasing the estimate upwards, which appears like bias reduction because the estimate is already low.
-
 %\texttt{TODO: Try to xplain the .50 jump in the estimate for transect/compounded. No theory or story atm.}
@@
 The movement data, (left panel, Figure~\ref{fig:moveTest}) is squashed more towards 0.1 Km than the still data (right panel).
-But there is a problem.
-The still data does not show the nice shoulder at 0 that the literature instructs us to expect.
+But there is a problem: The still data does not show the nice shoulder at 0 that the literature instructs us to expect.
 
@@
 \textcite{ramsey1981} speak about this, when comparing the ``CUM-D'' curve, or the plot of the cumulative number of observations at distance $r$ for both line and point transects.
-However, almost all the illustrations on how the collected data should look, shows data that echoes the detection probability curve: the half-bell-shape with a nice shoulder.
-Since this shape is impossible for VCP, explicit examples should be shown in the literature so researchers do not assume their is something wrong with their data.
+However, in the other references examined, almost all the illustrations on how the collected data should look, shows data that echoes the detection probability curve: the half-bell-shape with a nice shoulder.
+Since this shape is impossible for point transects, explicit examples should be shown in the literature so researchers do not assume there is something wrong with their data.
 
@@
 It is unclear from these results what the particular cause of the increased estimate and accuracy is.
-Future work might involve modification of the simulation functions to return more information, such as the count of observed animals and the observation distances to assist with understanding why this combination is producing an estimate that appears out of line with it's neighbors.
+Future work might involve modification of the simulation functions to return more information, such as the count of observed objects and the observation distances to assist with understanding why this combination is producing an estimate that appears out of line with it's neighbors.
 
@@
 %\bibliography{VCPbibliography}
+
+\subsection*{Code}
+Simulation code is available at \url{https://github.com/mhedwards/wombat-VCP}
 \end{document}

### 2014-06-07 report/draft3.tex
final, submitted version of report.

@@
 
+Does this realization invalidate the original simulation?
+Yes and no.
+In order to model the true detection probability curve from the Micronesian data, we would need to transform the empirical distances from the VCP in a fashion that would make them more equitable to the empirical distances we would get from a line transect.
+In other words, we would need to apply a transform to change the dotted line in Figure~\ref{fig:perfectCount} to the solid line in the same figure.
+So, our original simulation missed the ``letter'' of the intent.
+
+However, the spirit of the intent for the first simulation was to compare a ``perfect'' detectability curve to an imperfect one, and to that end, the simulation proceeded as intended.
+The detectability curve based on the empirical distances violated one of the fundamental assumptions of distance sampling, that $g(0)=1$, and we were able to observe the effect of that violation in the increased bias of our estimate.
+Future work should absolutely consider this transformation, and an evaluation of the actual detection probability curve present in the Collared Kingfisher data.
+
 %Thought - is this why we use h(0) for the point transects? need to verify, h(r) = f'(r)??  budkland 1987 paper has h(r) = f(r)/r, need to check teh book.  - yes.

### 2014-06-08 presentation/VCPSlides.tex
Presentation start

\documentclass{beamer}

\definecolor{OSUdkbrn}{RGB}{104,80,64}
\definecolor{OSUmedbrn}{RGB}{176,96,16}
\definecolor{OSUltbrn}{RGB}{224,195,152}
% from: http://oregonstate.edu/brand/color-palette

\mode<presentation>
{
	\usetheme{CambridgeUS}
	\setbeamercovered{transparent}
	\setbeamertemplate{footline}[page number]{}
	\setbeamercolor{palette tertiary}{fg=white, bg=OSUdkbrn}
	\setbeamercolor{title}{fg=OSUdkbrn}
	\setbeamercolor{frametitle}{fg=OSUdkbrn}
}

\title{Variable Circular Plots: Station Placement and the Independence Assumption}
\subtitle{Master's Research Project}
\author{Matt Edwards\\ Advisor: Claudio Fuentes}
\date{June 11, 2014}

\begin{document}
\begin{frame}
	\titlepage
\end{frame}

%\begin{frame}
%	\frametitle{Outline}
%	\tableofcontents
%\end{frame}

% -------------------------------- INTRODUCTION --------------------------------------------------
\section{Introduction}

\begin{frame}{Distance Sampling}
\begin{itemize}
\item Population Density Estimation
\item[]
\item Used often in Ecological Sciences
\item[]
\item Began with roadside surveys in early 1900s
\item[]
\item Theory developed through mid-1900s

\end{itemize}
\end{frame}

\begin{frame}{Distance Sampling: History}
\begin{itemize}
\item Emlen 1971
\item Ramsey, Scott 1979, 1980 Papers
\item Burnham, Anderson, and Jeffrey L. Laake 1980 Monograph
\item Buckland et. al. 1993, 2001
\item Continued Development into 2000s
\begin{itemize}
\item Combined with Mark-Recapture Methods
\item Extended for use with underwater acoustics for estimates of krill populations 
\item Combined with camera traps to estimate populations
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Distance Sampling}

\end{frame}

\end{document}

### 2014-06-08 presentation/VCPSlides.tex
Continued working on slides.

@@
 \documentclass{beamer}
+
+%\usepackage[utf8x]{inputenc}
+\usepackage{textcomp}
 
@@
 % from: http://oregonstate.edu/brand/color-palette
+
+\usepackage[backend=biber,citestyle=authoryear,bibstyle=authoryear,maxbibnames=10]{biblatex}
+\addbibresource{../report/VCPbibliography.bib}
 
@@
 	\setbeamercolor{frametitle}{fg=OSUdkbrn}
+	\setbeamertemplate{itemize items}[square]
+	\setbeamercolor*{item}{fg=OSUdkbrn}
 }
@@
 	\begin{itemize}
-\item Emlen 1971
-\item Ramsey, Scott 1979, 1980 Papers
-\item Burnham, Anderson, and Jeffrey L. Laake 1980 Monograph
-\item Buckland et. al. 1993, 2001
+	\item Emlen---1971
+	\item Ramsey, Scott---1979, 1980 Papers
+	\item Burnham, Anderson, and Jeffrey L. Laake---1980 Monograph
+	\item Buckland et. al. 1993, 2001---text
 	\item Continued Development into 2000s
@@
 
-\begin{frame}{Distance Sampling}
+\begin{frame}{Line Vs. Point Transects}
+	\begin{itemize}
+	\item Line transects are walked
+	\begin{itemize}
+		\item Distances perpendicular to transect are recorded, as projected on ground
+	\end{itemize}
+	\item[]
+	\item Point transects: observer(s) stand at station, observe everything within 360\textdegree 
+		\begin{itemize}
+		\item Alternately: Variable Circular Plots (VCP)
+		\item Allow ``cooling'' period
+		\item Safer for observer
+		\item Straight line distance from Observer is recorded, as projected on ground
+		\end{itemize}
+	\item[]		
+	\item Observations can be from visual sightings, auditory clues, or a combination
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Detection Curve: $g(r)$}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio=true]{../images/detectionCurve.pdf}
+	\end{figure}
+	\note{
+		\begin{itemize}
+		\item Perfect Detectability Near to Observer, dropping off. The ``shoulder''
+		\item $g(r)$ = probability of detecting an object, given it is at distance $r$
+		\item scaled so $g(0) = 1$
+		\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Density Estimation}
+	\begin{center}
+	$\hat{D}=\dfrac{n}{\mbox{Area}*P(\mbox{observing object}|\mbox{distance }r)}$\\
+	\vspace{0.5cm}
+	$\hat{D}=\dfrac{n}{\mbox{Area}*g(r)}$
+	\end{center}
+
+	\note{
+		\begin{itemize}
+		\item Much of the difficulty in distance sampling comes from estimating g(r)
+		\item For this presentation, the kernel method, a non-parametric method, was used.
+		\item in very broad terms, it looks at the average count per VCP and estimates area and probability by averaging over the observed observation distances inside a normal kernel. 
+		\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Micronesian Forest Bird Survey: 1982}
+\begin{itemize}
+\item Engbring, Ramsey \& Wildman (1986)
+\item[]
+\item Used VCP to survey several bird species in the Micronesian islands
+\item[]
+\item Each of 5 Islands divided into regions
+\item[]
+\item Transect randomly placed within region (angle \& starting position selected randomly)
+\item[]
+\item Stations placed every 150 m along transect
+\item[]
+\item Additional transects placed 2 km parallel
+\end{itemize}
+
+	\note{
+	\begin{itemize}
+	\item Observers would wait 2 minutes after arrival, then spend 8 minutes observing all they could
+	\item Distance and species were noted, as well as foliage and other variables.
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Collared Kingfisher Observation Data}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{../images/histogram_dist_20m.pdf}
+	\label{fig:82dist}
+\end{figure}
+
+	\note{
+		\begin{itemize}
+		\item Aggregated over two islands, Rota \& Tinan
+		\item Aggregated over all observers, 4
+		\item Primarily auditory cues, some visual, and some heard-then-seen
+		\item Data tell us 3 main things
+		\begin{itemize}
+			\item Very low bird counts immediately around the station. Some evidence of movement.
+			\item Evidence of heaping, the further away from station. More prominent with smaller bins.
+			\item Stations places 150 Meters Apart.
+		\end{itemize}
+		\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Research Question}
+Do overlapping observation areas violate any underlying independence assumptions?\\
+\vspace{1cm}Will it make a difference in our final population density estimates if a bird is observed from more than one station?
+\end{frame}
+
+\begin{frame}{VCP Indepencence}
+	\begin{itemize}
+	\item \textcite{ramsey1979,buckland1987,thompson2012} discuss assumption that VCPs are randomly placed.
+	\begin{itemize}
+	\item Implicit, but not explicit, possibility of overlap
+	\end{itemize}
+	\item[]
+	\item \textcite{reynolds1980} state the possibility of observing the same bird from more than one station should be avoided.
+	\item[]
+	\item \textcite{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet un-sampled lines.''
+	\end{itemize}
+
+		\note{
+			\begin{itemize}
+			\item Some Selected Quotes
+			\item Literature is divided
+			\item Transect layouts, similar to micronesian study, are common
+			\item We as scientists need to understand how this might bias our results.
+			\end{itemize}
+		}
+\end{frame}
+
+\begin{frame}{Final Words}
+\begin{quote}
+``All models are wrong, but some are useful.'' --George E. P. Box
+\end{quote}
 

### 2014-06-08 fiddling/DataExploration.R

@@ -43,7 +43,7 @@ ggplot(coki.82, aes(x=Distance))+geom_histogram(aes(fill = ..count..), binwidth=
 ggsave("images/histogram_distance5.pdf", width=8, height=4)
 
 ggplot(coki.82, aes(x=Distance))+geom_histogram(binwidth=20, fill=OSUOrange, color="white")+
-  theme_bw(18)+
+  theme_bw(18)+geom_vline(xintercept=c(150,300,450,600), color=OSUBlue2, size=2, linetype="dashed")+
   scale_x_continuous(breaks=seq(0,700,100))+xlab("Observation Distance (m)")+ylab("# of Birds")
 
 ggsave("images/histogram_dist_20m.pdf", width=8, height=4)

### 2014-06-08 presentation/VCPSlides.tex
Updated Slides, updated some graphics, created some new graphics.

@@
 \usepackage{textcomp}
+%\usepackage{graphicx}
+%\usepackage{subcaption}
+%\usepackage{floatrow}
+\usepackage{subfig}
+\usepackage{array}
 
@@
 	\setbeamertemplate{footline}[page number]{}
+	\setbeamertemplate{navigation symbols}{} % remove navigation symbols
 	\setbeamercolor{palette tertiary}{fg=white, bg=OSUdkbrn}
@@
 	\setbeamercolor*{item}{fg=OSUdkbrn}
+%	
+%	\setbeameroption{show only notes} % ----------------------- NOTES TOGGLE ---------------------
+%
 }
@@
 
+% -------------------------------- Micronesian Study --------------------------------------------------
+\section{Micronesian Study}
+
 \begin{frame}{Micronesian Forest Bird Survey: 1982}
@@
 	}
+\end{frame}
+
+\begin{frame}{Micronesia}
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/islands.jpg}
+	\end{figure}
+	\tiny
+	\begin{center}\color{OSUdkbrn}
+	Mariana Islands, modified from work by M. Minderhoud, (public domain) via  Wikimedia Commons
+	\end{center}
+
 \end{frame}
@@
 			\item Literature is divided
-			\item Transect layouts, similar to micronesian study, are common
+			\item Transect layouts, similar to Micronesian study, are common
 			\item We as scientists need to understand how this might bias our results.
+			\item if layout is going to bias our results, then we need to be aware of that, so adjustments can be made to survey design or analysis. 
 			\end{itemize}
 		}
+\end{frame}
+
+% -------------------------------- SIMULATION --------------------------------------------------
+\section{Simulation}
+
+\begin{frame}{Simulation Set Up}
+	Modeled after Palie Region on Rota:
+
+	\begin{itemize}
+	\item[]
+	\item 20 birds per km$^2$
+	\item[]
+	\item 9.41 km$^2$ Study Area
+	\item[]
+	\item Two transects, 16 \& 17 stations each
+	\item[]
+	\item Truncation distance, $w$, taken to be 500 m
+	\end{itemize}
+	\note{
+	\begin{itemize}
+	\item 	Mainly because a density of 20/km$^2$ was a nice round number.
+	\item There is only 0.9 \% of data beyond 500 m
+	\item let me deal with nicer numbers, while still adhering to the spirit of the original data.
+	\end{itemize}
+
+	}
+\end{frame}
+
+\begin{frame}{Transect Layouts}
+1 unit on graph = 1000 km
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutT2.pdf}
+		\end{column}
+		\begin{column}{0.48\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutT1.pdf}
+		\end{column}
+	\end{columns}
+
+	\note{
+	\begin{itemize}
+	\item 2 transects, 18 stations each
+	\item angle randomly chosen
+	\item starting position randomly chosen
+	\end{itemize}	
+	}
+\end{frame}
+
+\begin{frame}{Structured and Random Layouts}
+1 unit on graph = 1000 km
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutS.pdf}
+		\end{column}
+		\begin{column}{0.48\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutR.pdf}
+		\end{column}
+	\end{columns}
+
+	\note{
+		\begin{itemize}
+		\item Structured layout is why we're at 6km$^2$. needed to be able to have 500 m of direction in either area, without overlap
+		\item this gave 36 stations, which was carried over to the other layout types.
+		\item For Random Method, center point of VCP was randomly placed, within the buffer zone.
+		\end{itemize}	
+	}
+\end{frame}
+
+\begin{frame}{Detection Curve: Ideal}
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/detectionCurveScaled.pdf}
+	\end{figure}
+		\note{
+			\begin{itemize}
+			\item A halfnormal curve, scaled so that $g(0.5)~0$ and $g(0)=1$
+			\item used FDRTOOL package in R.
+			\end{itemize}	
+		}
+\end{frame}
+
+\begin{frame}{Detection Curve: Empirical}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\textwidth]{../images/loess.pdf}
+	\end{figure}
+
+		\note{
+			\begin{itemize}
+			\item data heaped into 5 m bins, midpoint of bin as x, density of bin * 5 as y
+			\item Loess regression performed with a span of 2. This is solid line.
+			\item Gives us object that can be used with predict function. Dashed line is estimate run with Predict to ensure.
+			\item As with the half-normal, the density was scaled so that the highest point gave a probability of 1.
+			\item with Predict, distances that extrapolated beyond the original dataset returned ``na'', these were set to a probability of detection of 0.
+			\end{itemize}	
+		}
+\end{frame}
+
+\begin{frame}{Movement}
+	Three Movement Options:
+	\begin{itemize}
+	\item No Movement: Objects stayed at origin point
+	\item[]
+	\item Temporary Movement: Objects would shift away from observer, but shift back to origin after observer ``left''
+	\item[]
+	\item Compounded Movement: Objects would shift away from observer, would remain at new point after observer moved. (Movement could compound if object was within movement radius for more than one VCP)
+	\end{itemize}
+
+	\note{
+		\begin{itemize}
+		\item Based on the empirical observations, movement was simulated within the first 100 m from observer
+		\item Higher chance of movement if closer to observer
+		\item Moved further if originated closer to observer.
+		\end{itemize}	
+	}
+\end{frame}
+% -------------------------------- RESULTS --------------------------------------------------
+\section{Results}
+
+\begin{frame}{Empirical vs. Half-Normal}
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/Emp_Vs_Hnorm_1-2.pdf}
+	\end{figure}
+	\note{
+		\begin{itemize}
+		\item 1000 Simulations. Each object layout was analyzed with the 6 combinations represented in the graph
+		\item Density estimates done using the kernel method with a normal kernel \parencite{quang1993}
+		\item Estimates for Empirical detection function biased low.
+		\item Estimates using half-normal detection function biased low, but not as bad.
+		\end{itemize}	
+	}
+
+\end{frame}
+
+\begin{frame}{Empirical vs. Half-Normal}{Centers and Spread}
+\begin{table}
+	\begin{tabular}{ r r |r r| r r r}
+		$g(x)$      & Layout     & Mean  & Std Dev & 25th  & Median & 75th  \\ \hline\hline
+		Empirical   & Random     & 14.47 & 3.37    & 12.08 & 14.38  & 16.73 \\
+		Empirical   & Structured & 14.49 & 3.15    & 12.27 & 14.47  & 16.59 \\
+		Empirical   & Transect   & 14.26 & 3.66    & 11.65 & 14.14  & 16.80 \\ \hline
+		Half-normal & Random     & 17.90 & 4.57    & 14.67 & 17.86  & 20.88 \\
+		Half-normal & Structured & 18.03 & 4.16    & 15.05 & 17.91  & 20.82 \\
+		Half-normal & Transect   & 17.80 & 4.64    & 14.52 & 17.44  & 20.78
+	\end{tabular}
+\end{table}
+
+	\note{
+	\begin{itemize}
+	\item supports graphic: Centers vary between detectability options, but not much between layouts.
+	\item May be some evidence of a downward shift with the Transect Layout. Can observe in Empirical mean \& median, and Half-normal Median. 
+	\item Hard to say if this is true effect or something else.
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Empirical vs. Half-Normal}{True Density Capture}
+\begin{table}
+
+	\begin{tabular}{ r| r r| r r|}
+		           & \multicolumn{2}{|c|}{Empirical} & \multicolumn{2}{|c|}{Half Normal} \\ \hline\hline
+		Layout     & Mean $\hat{D}$ & \% Capture     & Mean $\hat{D}$ & \% Capture       \\ \hline\hline
+		Random     & 14.47          & 73.6\%         & 17.90          & 94.0\%           \\
+		Structured & 14.49          & 77.5\%         & 18.03          & 95.3\%           \\
+		Transect   & 14.27          & 69.6\%         & 17.80          & 91.2\%           \\ \hline
+	\end{tabular}
+
+\end{table}
+
+	\note{
+	\begin{itemize}
+	\item Again, Transect layout may have some downward trend, as the capture percentages are lower here
+	\item Everything is relative. We have a larger up-tick in the Empirical/Structured cell than for Half-Normal and Structured.
+	\item Repeated runs of the simulation, and higher simulation counts could help determine if this is a significant pattern.
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Movement Results}
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/MovementSim2.pdf}
+	\end{figure}
+	\note{
+		\begin{itemize}
+		\item 1000 Simulations. Again, each object layout was analyzed with the 9 combinations represented in the graph
+		\item Density estimates done using the kernel method with a normal kernel \parencite{quang1993}
+		\item Again, layout does not seem to play as large of a part as the other factor.
+		\item All were run with half-normal detectability function
+		\end{itemize}	
+	}
+\end{frame}
+
+\begin{frame}{Movement Results}{Centers \& Spread}
+	\begin{table}
+
+		\begin{tabular}{ r r| r r| r r r}
+			Movement  & Layout     & Mean  & Std Dev & 25th  & Median & 75th  \\ \hline\hline
+			None      & Random     & 17.86 & 4.28    & 15.03 & 17.64  & 20.65 \\
+			None      & Structured & 18.01 & 4.51    & 14.95 & 17.62  & 20.94 \\
+			None      & Transect   & 17.90 & 4.78    & 14.74 & 17.67  & 20.95 \\ \hline
+			Compounded & Random     & 11.86 & 3.42    & 9.58  & 11.61  & 14.09 \\
+			Compounded & Structured & 11.90 & 3.65    & 9.37  & 11.67  & 14.08 \\
+			Compounded & Transect   & 12.42 & 3.72    & 9.93  & 12.17  & 14.79 \\ \hline
+			Temporary & Random     & 11.76 & 3.57    & 9.22  & 11.75  & 14.02 \\
+			Temporary & Structured & 11.74 & 3.48    & 9.37  & 11.56  & 13.85 \\
+			Temporary & Transect   & 11.75 & 3.57    & 9.24  & 11.72  & 14.20
+		\end{tabular}
+
+	\end{table}
+	\note{
+	\begin{itemize}
+	\item Reinforced histograms: results consistent within movement type, regardless of layout type
+	\item No Movement is consistent with what was observed in initial simulation
+	\item 
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Movement Results}{True Density Capture}
+	\begin{table}
+
+		\begin{tabular}{ r| c c| c c| c c|}
+
+			& \multicolumn{2}{|c|}{No Movement}	& \multicolumn{2}{|c|}{Compounded}	& \multicolumn{2}{|c|}{Temporary}\\ 
+	 \hline \hline
+
+	 Layout		&\footnotesize Mean $\hat{D}$	&\footnotesize \%  &\footnotesize Mean $\hat{D}$ &\footnotesize \%  &\footnotesize Mean $\hat{D}$ &\footnotesize \% 	\\ \hline \hline
+	 Random		& 17.86 			& 94.2\% 		& 11.86	& 59.6\%	& 11.76	& 59.3\% \\
+	 Structured	& 18.01 			& 94.3\% 		& 11.90 & 59.9\% 	& 11.74 & 59.5\% \\
+	 Transect	& 17.90 			& 90.8\% 		& 12.42 & 65.0\% 	& 11.75 & 58.0\% \\ \hline
+
+		\end{tabular}
+
+	\end{table}
+
+	\note{
+	\begin{itemize}
+	\item Movement type does not appear to play a large part in changing the bias.
+	\item We capture the true density about 59\% of the time, regardless of movement type.
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Movement And Layouts}
+
+	\begin{columns}
+		\begin{column}{0.3\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutS.pdf}
+		\end{column}
+		\begin{column}{0.3\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutR.pdf}
+		\end{column}
+		\begin{column}{0.3\textwidth}
+			\includegraphics[width=\textwidth]{../images/slides-layoutT2.pdf}
+		\end{column}
+	\end{columns}
+	\vspace{1cm}
+	1 unit = 1000 km, circles represent 200 m out from station
+	\note{
+		\begin{itemize}
+		\item Makes sense for structured, if the objects are only moving within the first 100 m, they would not move twice.
+		\item For the Randomized Layout, there are only 3, in this example, where there is a significant amount of overlap.
+If we take this as representative of the population of layouts, we would not be seeing many layouts with a significant amount of double or triple movement.
+An estimate from this layout should be similar to the Structured Layout, which we see.
+		\item It's with the Transect layout where the most compounded movement occurs.
+		\item (Go Back 1 Slide) There may be an effect of the combination of The Compounded movement and the Transect layout.
+		\item not sure if this is bias reduction, or an upward bias that looks like bias reduction because we're already low.
+		\end{itemize}
+	}
+\end{frame}
+
+% -------------------------------- Lines vs. Circles --------------------------------------------------
+\section{Problem with Circles}
+
+% -------------------------------- CONCLUSION --------------------------------------------------
+\section{Conclusion}
+
+\begin{frame}{Conclusion}
+	If all else is held equal, VCP layout does not seem to play a large role in the bias of population density estimates using the kernel method described by \textcite{quang1993}.\par
+		\vspace{1cm}
+	Violations of the expected detection probability curve, or by movement of the objects away from the observer play a greater role in biasing the resulting estimate.
+
+	\note{
+	\begin{itemize}
+	\item All of the estimates were biased low using Quang's formula
+	\item this was observed even with an ``ideal'' half-normal detection curve
+	\item Dr. Quang's paper does offer a bias-corrected estimate, but it was not used for this project.
+	\end{itemize}
+	}
+\end{frame}
+
+\begin{frame}{Future Work}
+	\begin{itemize}
+	\item Explore the possible interaction effect from Compounded Movement and the Transect Layout
+	\item[]
+	\item Modify simulation functions to return more information to assist diagnostics
+	\begin{itemize}
+	\item Count of Observed Objects
+	\item Observation Distances
+	\end{itemize}
+	\item[]
+	\item Modify using the Empirical data as a detectability function, to account for difference between line and point transects.
+	\item[]
+	\item Add the option to bias-correct the estimates, to see how well Quang's kernel method adjusts for these variations.
+	\end{itemize}
 \end{frame}
@@
 
+\begin{frame}{Questions? }
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/kingfisher.jpg}
+	\end{figure}
+	\tiny
+	\begin{center}\color{OSUdkbrn}
+	``Collared Kingfisher (Todiramphus chloris)'' by JJ Harrison, (CC-BY-SA 3.0) via  Wikimedia Commons
+	\end{center}
+
+\end{frame}
+
+\section{Additional Data}
+\begin{frame}{Object placement Simulation}
+
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/slides-pointGeneration.pdf}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Half-Normal Scaling}
+	The half-normal parameter $\theta$ is related to the standard deviation of a standard normal distribution by the relationship: 
+	\begin{center}
+	$\theta = \dfrac{\sqrt{\pi /2}}{\sigma}$
+	\end{center}
+
+	If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where $P(X > 3.5) < 0.001$) then $\theta$ is:
+	\begin{center}
+	$\sigma = \dfrac{w}{3.5} = 0.1429$\\
+	\vspace{0.5cm}
+	$\theta = \dfrac{\sqrt{\pi /2}}{0.1429}=8.7732$
+	\end{center}
+\end{frame}
+
+\begin{frame}{Half-Normal Scaling}
+	For a half-normal with parameter $\theta=8.7732$, $f(0)=5.5852$ so to scale the value to 1:
+	\begin{center}
+	$\delta_{HN} = \dfrac{1}{5.5852} = 0.1790$
+	\end{center}
+
+	Half-normal density functions were executed in R with the \texttt{fdrtool} package.
+\end{frame}
+
 \end{document}

### 2014-06-08 code/AdditionalGraphics.R

@@ -1,0 +1,81 @@
+library(fdrtool) #half normal
+library(ggplot2)
+
+OSUOrange <- rgb(243/255,115/255,33/255) #http://oregonstate.edu/brand/color-palette
+OSURed <- rgb(192/255,49/225,26/255)
+OSUBlue1 <- rgb(93/255, 135/255, 161/255)
+OSUBlue2 <- rgb(156/255, 197/255, 202/255)
+
+hn <- VCP.defineHalfnorm(500)
+
+x <- seq(0, 500, length=100)
+y <- dhalfnorm(x, hn$theta)*hn$delta
+
+
+
+
+
+ggplot()+geom_line(aes(x=x, y=y), size=2, color=OSUOrange)+theme_bw(18)+xlab("Distance from transect: r")+ylab("g(r)")
+
+ggsave("images/detectionCurve.pdf", width=8, height=6)
+
+
+
+hn.5 <- VCP.defineHalfnorm(.5)
+
+x <- seq(0, .5, length=100)
+y <- dhalfnorm(x, hn.5$theta)*hn.5$delta
+
+ggplot()+geom_line(aes(x=x, y=y), size=2, color=OSUOrange)+theme_bw(18)+xlab("Distance from transect: r")+ylab("g(r)")
+
+ggsave("images/detectionCurveScaled.pdf", width=8, height=6)
+
+
+
+#### circles vs squares
+OSUOrange <- rgb(243/255,115/255,33/255) #http://oregonstate.edu/brand/color-palette
+OSURed <- rgb(192/255,49/225,26/255)
+OSUBlue1 <- rgb(93/255, 135/255, 161/255)
+OSUBlue2 <- rgb(156/255, 197/255, 202/255)
+OSUMedBrn <- rgb(176/255,96/255,16/255)
+OSUDkGray <- rgb(171/255,175/255,166/255)
+# requires dplyr
+# requires get.Coords from VCP_Utilities.R
+
+library(dplyr)
+library(ggplot2)
+
+# get coords, using a density of 20 birds/km2
+x.y <- get.Coords(20)
+
+
+
+ggplot()+geom_point(data=x.y, aes(x,y), color=OSUDkGray)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)+
[…]

### 2014-06-08 code/ExampleGraphs.R

@@ -4,6 +4,7 @@ OSURed <- rgb(192/255,49/225,26/255)
 OSUBlue1 <- rgb(93/255, 135/255, 161/255)
 OSUBlue2 <- rgb(156/255, 197/255, 202/255)
 OSUMedBrn <- rgb(176/255,96/255,16/255)
+OSUDkGray <- rgb(171/255,175/255,166/255)
 # requires dplyr
 # requires get.Coords from VCP_Utilities.R
 
@@ -27,9 +28,13 @@ for(i in 1:36) {
 
 # generate and save plot
 # you will get warnings when running these, because it is trimming out points to get to the 0,6 bounds
-ggplot()+geom_point(data=x.y, aes(x,y))+geom_point(aes(x=stat.X, y=stat.Y), color=OSUOrange, inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour=OSUOrange, size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+ggplot()+geom_point(data=x.y, aes(x,y), color=OSUDkGray)+geom_point(aes(x=stat.X, y=stat.Y), color=OSUOrange, inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour=OSUOrange, size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
 
 ggsave("images/layout_structured.pdf", width=8, height=8)
+
+
+ggplot()+geom_point(data=x.y, aes(x,y), color=OSUDkGray)+geom_point(aes(x=stat.X, y=stat.Y), color=OSUMedBrn, inherit.aes=FALSE)+geom_path(data=ell, aes(a,b,group=gp), colour=OSUMedBrn, size=.75)+coord_fixed(xlim=c(-0.1,6.1), ylim=c(-0.1,6.1))+theme_bw(18)
+ggsave("images/slides-layoutS.pdf", width=8, height=8)
 
 
 ## -------------- Random VCP layout -------------- 
@@ -45,10 +50,17 @@ for(i in 1:36) {
   ell <- rbind(ell, df)
 }
 
[…]

### 2014-06-08 fiddling/DataExploration.R

@@ -10,6 +10,7 @@ library(lubridate) # date/time manipulation
 
 raw.data <- tbl_df(read.csv("data/COKI.csv", stringsAsFactors=FALSE))
 
+
 ## ===== CODING ====
 
 Labels.Type <- c("Heard", "Seen", "", "H&S")

### 2014-06-09 code/AdditionalGraphics.R
FInal Slides update for tonight
modified a copule graphics to fit presentation color scheme.

@@ -39,6 +39,7 @@ OSUBlue1 <- rgb(93/255, 135/255, 161/255)
 OSUBlue2 <- rgb(156/255, 197/255, 202/255)
 OSUMedBrn <- rgb(176/255,96/255,16/255)
 OSUDkGray <- rgb(171/255,175/255,166/255)
+OSUDkBrn <- rgb(104/255,80/255,64/255)
 # requires dplyr
 # requires get.Coords from VCP_Utilities.R
 

### 2014-06-09 code/VCPvsLine.R

@@ -6,7 +6,7 @@
 
 D <- 20
 hn.paras <- VCP.defineHalfnorm(.5)
-L <- 36*.15 # length of single transect of 36 VCP separated by .15 km
+#L <- 36*.15 # length of single transect of 36 VCP separated by .15 km
 r <- seq(0,.5, by=0.01)
 
 pi*(.5^2) #- area of circle with radius= 0.5 = 0.7853982
@@ -30,7 +30,7 @@ plot(r, A.lt)
 plot(r, A.vcp)
 
 library(ggplot2)
-ggplot()+geom_line(aes(x=r, y=A.lt), size=2)+geom_line(aes(x=r, y=A.vcp), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("Area covered")
+ggplot()+geom_line(aes(x=r, y=A.lt), size=2, colour=OSUDkBrn)+geom_line(aes(x=r, y=A.vcp), linetype="dashed", colour=OSUOrange, size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("Area covered")
 
 ggsave("images/rect-circ-area.pdf", width=6, height=6)
 
@@ -50,7 +50,7 @@ p.dpts <- VCP.dhnorm(d.pts, hn.paras)
 LT <- D * A.lt.diff[2:51]*p.dpts
 VCP <- D * A.vcp.diff[2:51] * p.dpts
 
-ggplot()+geom_line(aes(x=r[2:51], y=LT), size=2)+geom_line(aes(x=r[2:51], y=VCP), linetype="dashed", colour="red", size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("E[N]")
+ggplot()+geom_line(aes(x=r[2:51], y=LT), size=2, colour=OSUDkBrn)+geom_line(aes(x=r[2:51], y=VCP), linetype="dashed", colour=OSUOrange, size=2)+theme_bw(18)+xlab("Distance from transect")+ylab("E[N]")
 
 ggsave("images/rect-circ-detection.pdf", width=6, height=6)
 

### 2014-06-09 presentation/VCPSlides.tex
FInal Slides update for tonight
modified a copule graphics to fit presentation color scheme.
updated slides and notes.

@@
 %	
-%	\setbeameroption{show only notes} % ----------------------- NOTES TOGGLE ---------------------
+	\setbeameroption{show only notes} % ----------------------- NOTES TOGGLE ---------------------
 %
@@
 	\item[]
-	\item \textcite{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet un-sampled lines.''
+	\item \textcite{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transects, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet un-sampled lines.''
 	\end{itemize}
@@
 \begin{frame}{Transect Layouts}
-1 unit on graph = 1000 km
+1 unit on graph = 1 km
 	\begin{columns}
@@
 \begin{frame}{Structured and Random Layouts}
-1 unit on graph = 1000 km
+1 unit on graph = 1 km
 	\begin{columns}
@@
 	\vspace{1cm}
-	1 unit = 1000 km, circles represent 200 m out from station
+	1 unit = 1 km, circles represent 200 m out from station
 	\note{
@@
 % -------------------------------- Lines vs. Circles --------------------------------------------------
-\section{Problem with Circles}
+\section{Problems}
+
+\begin{frame}{Movement Testing}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\textwidth]{../images/movementTest.pdf}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Change in Area}
+
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{../images/slides-LTr.pdf}
+		\end{column}
+		\begin{column}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{../images/slides-VCPr.pdf}
+		\end{column}
+	\end{columns}
+
+\end{frame}
+
+\begin{frame}{Circles vs. Rectangles}{Area and Expected Observation Counts}
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{../images/rect-circ-area.pdf}
+		\end{column}
+		\begin{column}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{../images/rect-circ-detection.pdf}
+		\end{column}
+	\end{columns}
+	\begin{center}
+	\par Solid line is Line Transect, Dashed line is VCP
+	\end{center}
+
+\end{frame}
 

### 2014-06-10 presentation/ExtraSlides.tex
added Text to slides at Claudio's suggestion.

--- 2014-06-08 presentation/VCPSlides.tex
@@
 \subtitle{Master's Research Project}
-\author{Matt Edwards\\ Advisor: Claudio Fuentes}
+\author{Matt Edwards}
 \date{June 11, 2014}
@@
 \begin{document}
-\begin{frame}
-	\titlepage
+\section{Additional Data}
+\begin{frame}{Object placement Simulation}
+
+	\begin{figure}
+		\centering
+		\includegraphics[height=.80\textheight]{../images/slides-pointGeneration.pdf}
+	\end{figure}
 \end{frame}
 
-%\begin{frame}
-%	\frametitle{Outline}
-%	\tableofcontents
-%\end{frame}
+\begin{frame}{Half-Normal Scaling}
+	The half-normal parameter $\theta$ is related to the standard deviation of a standard normal distribution by the relationship: 
+	\begin{center}
+	$\theta = \dfrac{\sqrt{\pi /2}}{\sigma}$
+	\end{center}
 
-% -------------------------------- INTRODUCTION --------------------------------------------------
-\section{Introduction}
-
-\begin{frame}{Distance Sampling}
-	\begin{itemize}
-	\item Population Density Estimation
-	\item[]
-	\item Used often in Ecological Sciences
-	\item[]
-	\item Began with roadside surveys in early 1900s
-	\item[]
-	\item Theory developed through mid-1900s
-
-	\end{itemize}
+	If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where $P(X > 3.5) < 0.001$) then $\theta$ is:
+	\begin{center}
+	$\sigma = \dfrac{w}{3.5} = 0.1429$\\
+	\vspace{0.5cm}
+	$\theta = \dfrac{\sqrt{\pi /2}}{0.1429}=8.7732$
+	\end{center}
 \end{frame}
 
-\begin{frame}{Distance Sampling: History}
-	\begin{itemize}
-	\item Emlen---1971
-	\item Ramsey, Scott---1979, 1980 Papers
-	\item Burnham, Anderson, and Jeffrey L. Laake---1980 Monograph
-	\item Buckland et. al. 1993, 2001---text
-	\item Continued Development into 2000s
-	\begin{itemize}
-	\item Combined with Mark-Recapture Methods
-	\item Extended for use with underwater acoustics for estimates of krill populations 
-	\item Combined with camera traps to estimate populations
-	\end{itemize}
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Line Vs. Point Transects}
-	\begin{itemize}
-	\item Line transects are walked
-	\begin{itemize}
-		\item Distances perpendicular to transect are recorded, as projected on ground
-	\end{itemize}
-	\item[]
-	\item Point transects: observer(s) stand at station, observe everything within 360\textdegree 
-		\begin{itemize}
-		\item Alternately: Variable Circular Plots (VCP)
-		\item Allow ``cooling'' period
-		\item Safer for observer
-		\item Straight line distance from Observer is recorded, as projected on ground
-		\end{itemize}
-	\item[]		
-	\item Observations can be from visual sightings, auditory clues, or a combination
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Detection Curve: $g(r)$}
-	\begin{figure}
-		\centering
-		\includegraphics[width=\textwidth,height=0.80\textheight,keepaspectratio=true]{../images/detectionCurve.pdf}
-	\end{figure}
-	\note{
-		\begin{itemize}
-		\item Perfect Detectability Near to Observer, dropping off. The ``shoulder''
-		\item $g(r)$ = probability of detecting an object, given it is at distance $r$
-		\item scaled so $g(0) = 1$
-		\end{itemize}
-	}
-\end{frame}
-
-\begin{frame}{Density Estimation}
+\begin{frame}{Half-Normal Scaling}
+	For a half-normal with parameter $\theta=8.7732$, $f(0)=5.5852$ so to scale the value to 1:
 	\begin{center}
-	$\hat{D}=\dfrac{n}{\mbox{Area}*P(\mbox{observing object}|\mbox{distance }r)}$\\
-	\vspace{0.5cm}
-	$\hat{D}=\dfrac{n}{\mbox{Area}*g(r)}$
+	$\delta_{HN} = \dfrac{1}{5.5852} = 0.1790$
 	\end{center}
 
-	\note{
-		\begin{itemize}
-		\item Much of the difficulty in distance sampling comes from estimating g(r)
-		\item For this presentation, the kernel method, a non-parametric method, was used.
-		\item in very broad terms, it looks at the average count per VCP and estimates area and probability by averaging over the observed observation distances inside a normal kernel. 
-		\end{itemize}
-	}
-\end{frame}
-
-\begin{frame}{Micronesian Forest Bird Survey: 1982}
-\begin{itemize}
-\item Engbring, Ramsey \& Wildman (1986)
-\item[]
-\item Used VCP to survey several bird species in the Micronesian islands
-\item[]
-\item Each of 5 Islands divided into regions
-\item[]
-\item Transect randomly placed within region (angle \& starting position selected randomly)
-\item[]
-\item Stations placed every 150 m along transect
-\item[]
-\item Additional transects placed 2 km parallel
-\end{itemize}
-
-	\note{
-	\begin{itemize}
-	\item Observers would wait 2 minutes after arrival, then spend 8 minutes observing all they could
-	\item Distance and species were noted, as well as foliage and other variables.
-	\end{itemize}
-	}
-\end{frame}
-
-\begin{frame}{Collared Kingfisher Observation Data}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{../images/histogram_dist_20m.pdf}
-	\label{fig:82dist}
-\end{figure}
-
-	\note{
-		\begin{itemize}
-		\item Aggregated over two islands, Rota \& Tinan
-		\item Aggregated over all observers, 4
-		\item Primarily auditory cues, some visual, and some heard-then-seen
-		\item Data tell us 3 main things
-		\begin{itemize}
-			\item Very low bird counts immediately around the station. Some evidence of movement.
-			\item Evidence of heaping, the further away from station. More prominent with smaller bins.
-			\item Stations places 150 Meters Apart.
-		\end{itemize}
-		\end{itemize}
-	}
-\end{frame}
-
-\begin{frame}{Research Question}
-Do overlapping observation areas violate any underlying independence assumptions?\\
-\vspace{1cm}Will it make a difference in our final population density estimates if a bird is observed from more than one station?
-\end{frame}
-
-\begin{frame}{VCP Indepencence}
-	\begin{itemize}
-	\item \textcite{ramsey1979,buckland1987,thompson2012} discuss assumption that VCPs are randomly placed.
-	\begin{itemize}
-	\item Implicit, but not explicit, possibility of overlap
-	\end{itemize}
-	\item[]
-	\item \textcite{reynolds1980} state the possibility of observing the same bird from more than one station should be avoided.
-	\item[]
-	\item \textcite{buckland2001} states ``Transects are normally spaced at a sufficient distance to avoid detecting an object from two neighboring transacts, although this is not usually critical unless sampling a line changes the animal distribution at neighboring, as yet un-sampled lines.''
-	\end{itemize}
-
-		\note{
-			\begin{itemize}
-			\item Some Selected Quotes
-			\item Literature is divided
-			\item Transect layouts, similar to micronesian study, are common
-			\item We as scientists need to understand how this might bias our results.
-			\end{itemize}
-		}
-\end{frame}
-
-\begin{frame}{Final Words}
-\begin{quote}
-``All models are wrong, but some are useful.'' --George E. P. Box
-\end{quote}
-
+	Half-normal density functions were executed in R with the \texttt{fdrtool} package.
 \end{frame}

### 2014-06-10 presentation/VCPSlides.tex
corrected a few things.
added Text to slides at Claudio's suggestion.

@@
 %	
-	\setbeameroption{show only notes} % ----------------------- NOTES TOGGLE ---------------------
+%	\setbeameroption{show only notes} % ----------------------- NOTES TOGGLE ---------------------
 %
@@
 \begin{frame}{Distance Sampling}
+Distance sampling is a method of population density estimation that considers the probably of observing the object of interest at the distance at which it was seen.
+
 	\begin{itemize}
-	\item Population Density Estimation
-	\item[]
 	\item Used often in Ecological Sciences
@@
 \begin{frame}{Distance Sampling: History}
+A selection of relevant papers:
 	\begin{itemize}
@@
 
-\begin{frame}{Line Vs. Point Transects}
+\begin{frame}{Line vs Point Transects}
 	\begin{itemize}
@@
 	\item[]		
-	\item Observations can be from visual sightings, auditory clues, or a combination
+	\item Both: Observations can be from visual sightings, auditory clues, or a combination
 	\end{itemize}
@@
 
+\begin{frame}{Transect Examples}
+
+	\begin{columns}
+		\begin{column}{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{../images/slides-LTr.pdf}\\
+		Line Transect
+		\end{column}
+		\begin{column}{0.48\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{../images/slides-VCPr.pdf}\\
+		VCP/Point Transect
+		\end{column}
+	\end{columns}
+
+\end{frame}
+
 \begin{frame}{Detection Curve: $g(r)$}
+The probability of detecting an object, give it is at distance $r$.
 	\begin{figure}
@@
 	\end{center}
+
+	\begin{itemize}
+	\item $\hat{D}$: Estimated Population Density
+	\item $n$: number of objects observed
+	\item $Area$: The area surveyed
+	\item $g(r)$: $P(\mbox{observing object}|\mbox{distance }r)$
+	\end{itemize}
 
@@
 \begin{frame}{Collared Kingfisher Observation Data}
-
+Aggregated empirical observation distances from 1982 study.
 \begin{figure}
@@
 \begin{frame}{Research Question}
-Do overlapping observation areas violate any underlying independence assumptions?\\
-\vspace{1cm}Will it make a difference in our final population density estimates if a bird is observed from more than one station?
+\Large
+{\bfseries Do overlapping observation areas violate any underlying independence assumptions?}\\
+\vspace{1cm}
+{\bfseries Will it make a difference in our final population density estimates if a bird is observed from more than one station?}
 \end{frame}
@@
 \begin{frame}{VCP Indepencence}
+	VCP Analysis assumes sightings of animals are independent events.
 	\begin{itemize}
+	\item[]
 	\item \textcite{ramsey1979,buckland1987,thompson2012} discuss assumption that VCPs are randomly placed.
@@
 \begin{frame}{Transect Layouts}
+Examples of two random Transect layouts.\\
 1 unit on graph = 1 km
@@
 \begin{frame}{Structured and Random Layouts}
+Examples of Structured and Random layouts.\\
 1 unit on graph = 1 km
@@
 \begin{frame}{Detection Curve: Ideal}
+``Ideal'' detection probability curve, maximum observation distance set to 0.5 unit (500 m) distance, and scaled so $g(0)=1$
 	\begin{figure}
 		\centering
-		\includegraphics[height=.80\textheight]{../images/detectionCurveScaled.pdf}
+		\includegraphics[height=.75\textheight]{../images/detectionCurveScaled.pdf}
 	\end{figure}
@@
 \begin{frame}{Detection Curve: Empirical}
+Detection probability curve from empirical data, shown unscaled.
 	\begin{figure}
@@
 \begin{frame}{Movement}
+	The histogram of observed distances shows evidence of movement, so it was considered as an option for simulation.\\
+	\vspace{0.33cm}
 	Three Movement Options:
@@
 \begin{frame}{Empirical vs. Half-Normal}
+	\small
+	1000 simulations, comparing Empirical vs Half-Normal detection probability, no movement, 
 	\begin{figure}
@@
 \begin{frame}{Empirical vs. Half-Normal}{Centers and Spread}
+
 \begin{table}
@@
 \end{table}
+\small
+Comparing measures of centers and spread, 1000 simulations, no movement.
 
@@
 
+\small
+Comparing rates of capture for true object density, 1000 simulations, no movement.
+
 	\note{
@@
 \begin{frame}{Movement Results}
+\small
+	1000 simulations, comparing effect of movement types. Half-normal detection probability used for all.
 	\begin{figure}
@@
 \begin{frame}{Movement Results}{Centers \& Spread}
+
 	\begin{table}
@@
 		\end{tabular}
-
+	\small
+	Comparing measures of center and spread for different movement types. 1000 simulations, half-normal detection probability used for all.
 	\end{table}
@@
 \begin{frame}{Movement Results}{True Density Capture}
+
 	\begin{table}
@@
 	\end{table}
-
+	\small
+	Comparing capture of true object density for different movement types. 1000 simulations, half-normal detection probability used for all.
 	\note{
@@
 \begin{frame}{Movement Testing}
+Results of movement code test (left) vs no movement ``baseline'' (right).
 	\begin{figure}
@@
 \begin{frame}{Change in Area}
-
+	Area increase from 1 unit expansion in Line Transect does not equal the area from the same increase in a VCP.
 	\begin{columns}
@@
 		\begin{column}{0.48\textwidth}
-		\includegraphics[width=\textwidth]{../images/rect-circ-area.pdf}
+		\centering
+		\includegraphics[width=\textwidth]{../images/rect-circ-area.pdf}\\
+		\small
+		Increase in Area
 		\end{column}
 		\begin{column}{0.48\textwidth}
-		\includegraphics[width=\textwidth]{../images/rect-circ-detection.pdf}
+		\centering
+		\includegraphics[width=\textwidth]{../images/rect-circ-detection.pdf}\\
+		\small
+		Expected number of observations, given distance.
 		\end{column}
@@
 \begin{frame}{Conclusion}
-	If all else is held equal, VCP layout does not seem to play a large role in the bias of population density estimates using the kernel method described by \textcite{quang1993}.\par
-		\vspace{1cm}
-	Violations of the expected detection probability curve, or by movement of the objects away from the observer play a greater role in biasing the resulting estimate.
+There are two primary conclusions:\\
+\vspace{0.5cm}
+	\begin{itemize}
+	\item If all else is held equal, VCP layout does not seem to play a large role in the bias of population density estimates using the kernel method described by \textcite{quang1993}.
+	\item[]
+	\item Violations of the expected detection probability curve, or by movement of the objects away from the observer play a greater role in biasing the resulting estimate.
+	\end{itemize}
 
@@
 \begin{frame}{Future Work}
+Some possible areas for future research:
+
 	\begin{itemize}
@@
 	\item Modify using the Empirical data as a detectability function, to account for difference between line and point transects.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Future Work}
+	\begin{itemize}
+	\item Add the option to bias-correct the estimates, to see how well Quang's kernel method adjusts for these variations.
 	\item[]
-	\item Add the option to bias-correct the estimates, to see how well Quang's kernel method adjusts for these variations.
+	\item Address analyzing transects as clusters, since observations on a transect are likely to be related.
+
 	\end{itemize}
@@
 	\end{quote}
+\end{frame}
 
+	\begin{frame}{Acknowledgements}
+	Micronesian Survey Data courtesy of Fred Ramsey\\
+	\vspace{1cm}
+	Masters Committee:
+	\begin{itemize}
+	\item[] Claudio Fuentes, Advisor
+	\item[] Alix Gitelman, Committee Member
+	\item[] Charlotte Wickham, Committee Member
+	\end{itemize}
 \end{frame}
@@
 
-\section{Additional Data}
-\begin{frame}{Object placement Simulation}
-
-	\begin{figure}
-		\centering
-		\includegraphics[height=.80\textheight]{../images/slides-pointGeneration.pdf}
-	\end{figure}
-\end{frame}
-
-\begin{frame}{Half-Normal Scaling}
-	The half-normal parameter $\theta$ is related to the standard deviation of a standard normal distribution by the relationship: 
-	\begin{center}
-	$\theta = \dfrac{\sqrt{\pi /2}}{\sigma}$
-	\end{center}
-
-	If we estimate $\sigma$ by $w/3.5$ (with $w$ being the maximum distance at which we might observe an object, and 3.5 being the standard deviation point where $P(X > 3.5) < 0.001$) then $\theta$ is:
-	\begin{center}
-	$\sigma = \dfrac{w}{3.5} = 0.1429$\\
-	\vspace{0.5cm}
-	$\theta = \dfrac{\sqrt{\pi /2}}{0.1429}=8.7732$
-	\end{center}
-\end{frame}
-
-\begin{frame}{Half-Normal Scaling}
-	For a half-normal with parameter $\theta=8.7732$, $f(0)=5.5852$ so to scale the value to 1:
-	\begin{center}
-	$\delta_{HN} = \dfrac{1}{5.5852} = 0.1790$
-	\end{center}
-
-	Half-normal density functions were executed in R with the \texttt{fdrtool} package.
-\end{frame}
-
 \end{document}
